### 2017-03-04 A New Hope/Catherine_Analysis.Rmd
Create Catherine_Analysis.Rmd

---
title: "Mines Accident Prediction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mine Accidents
Data sources

```{r Packages}
require(dplyr)
library(RcppRoll)
library(zoo)
```

```{r loading raw data}
setwd("C:/Users/CATHY/OneDrive/Documents/2016-2017 Junior/15 Mines Research/violation-data-analysis/")

if(!exists("accidents")) {
  accidents <- load("./Analysis/data/mshs_accidents.RData")
}
if(!exists("violations")) {
  violations <- load("./Analysis/data/mshs_violations.RData")
}
if(!exists("assessed_violations")) {
  assessed_violations <- load("./Analysis/data/mshs_assssd_violation.RData")
}
```

```{r accident cleaning}
actual.accidents <- accidents %>% filter(is.element(inj_degr_desc, c("DAYS AWAY FROM WORK ONLY", 
                                                                     "DYS AWY FRM WRK & RESTRCTD ACT", 
                                                                     "DAYS RESTRICTED ACTIVITY ONLY", 
                                                                     "FATALITY", 
                                                                     "PERM TOT OR PERM PRTL DISABLTY", 
                                                                     "NO VALUE FOUND")))
na.and.no.value.index <- is.element(actual.accidents$inj_degr_desc, "NO VALUE FOUND") & is.na(actual.accidents$days_lost)
actual.accidents[na.and.no.value.index, "days_restrict"] <- 0
actual.accidents[na.and.no.value.index, "days_lost"] <- 0
actual.accidents[which(actual.accidents$inj_degr_desc == "FATALITY"), "days_lost"] <- 300
actual.accidents$ai_dt_actual_date <- strftime(actual.accidents$ai_dt, "%F")

day.level.accidents <- actual.accidents %>% group_by(mine_id, ai_dt) %>% summarize(day_level_days_lost = sum(days_lost, na.rm = T))
quarter.level.accidents <- actual.accidents %>% group_by(mine_id, cal_qtr, cal_yr) %>% summarize(quarter_level_days_lost = sum(days_lost, na.rm = T))
```

```{r accident calculation}

longest.period <- 3
actual.start.year <- 2000
start.year <- 2000 - longest.period
mines.with.accidents.post.start.year <- (actual.accidents %>% filter(cal_yr >= actual.start.year))$mine_id %>% unique()
end.year <- 2015
number.years <- end.year - start.year + 1

roll_over <- function(longest.period, actual.start.year, start.year, )
## Challenge: fill in nonpresent values for mines with 0 days lost
## For example, mine could have just not had a day lost, but that won't show up in `quarter.level.accidents`
## 1. First, initialize all blank data frame by figuring out final structure (we know exactly what pattern mine_id, quarter, and year should follow
##    a particular set of repetitions)
## 2. Anti-join will only keep data from the set of keys not found in the actual data
## 3. Then we can just join the two sets of rows!
blank.full.quarter.level.days_lost <- data.frame(mine_id = rep(mines.with.accidents.post.start.year, each = 4 * number.years),
                                                 quarter = rep(1:4, times = length(mines.with.accidents.post.start.year) * number.years),
                                                 year = rep(rep(start.year:end.year, times = rep(4, number.years)),
                                                            times = length(mines.with.accidents.post.start.year)),
                                                 num.days.lost = 0)
full.quarter.level.days_lost <- anti_join(blank.full.quarter.level.days_lost, quarter.level.accidents,
                                          by = c("mine_id" = "mine_id", "quarter" = "cal_qtr", "year" = "cal_yr"))
full.quarter.level.days_lost <- rbind(full.quarter.level.days_lost,
                                      quarter.level.accidents %>% 
                                        select(mine_id, quarter = cal_qtr,
                                               year = cal_yr, num.days.lost = quarter_level_days_lost) %>% 
                                        data.frame() %>% filter(year >= start.year) %>% 
                                        filter(year <= end.year)) %>% arrange(mine_id, year, quarter)

## Challenge: Calculating past period statistics
## lag will put value of index n - 1 at index n
## roll_sum will do a sum from the specified number of previous rows and put NA's at the beginning
## We filter with actual.start.year so there are no NA's. 
##   (The year that data is ready and the year that data needs to start being used are different.)
full.quarter.level.days_lost <- full.quarter.level.days_lost %>% 
                                group_by(mine_id) %>%
                                mutate(last.quarter = lag(num.days.lost),
                                       last.year = roll_sum(lag(num.days.lost), 4, align = "right", fill = NA),
                                       last.three.years = roll_sum(lag(num.days.lost), 12, align = "right", fill = NA)) %>%
                                filter(year >= actual.start.year)

full.quarter.level.days_lost
#664128
```
```{r violation cleaning}
violations <- mshs_violations
quarter.level.violations <- violations %>% group_by(mine_id, cal_qtr, cal_yr) %>% 
                                            summarize(quarter_level_violation = n())

longest.period <- 3
actual.start.year <- 2000
start.year <- 2000 - longest.period
mines.with.violation.post.start.year <- (violations %>% filter(cal_yr >= actual.start.year))$mine_id %>% unique()
end.year <- 2015
number.years <- end.year - start.year + 1

## Challenge: fill in nonpresent values for mines with 0 days lost
## For example, mine could have just not had a day lost, but that won't show up in `quarter.level.accidents`
## 1. First, initialize all blank data frame by figuring out final structure (we know exactly what pattern mine_id, quarter, and year should follow
##    a particular set of repetitions)
## 2. Anti-join will only keep data from the set of keys not found in the actual data
## 3. Then we can just join the two sets of rows!
blank.full.quarter.level.violations <- data.frame(mine_id = rep(mines.with.violation.post.start.year, each = 4 * number.years),
                                                 quarter = rep(1:4, times = length(mines.with.violation.post.start.year) * number.years),
                                                 year = rep(rep(start.year:end.year, times = rep(4, number.years)),
                                                            times = length(mines.with.violation.post.start.year)),
                                                 viol.quantity = 0)
full.quarter.level.violations <- anti_join(blank.full.quarter.level.violations, quarter.level.violations,
                                          by = c("mine_id" = "mine_id", "quarter" = "cal_qtr", "year" = "cal_yr"))
full.quarter.level.violations <- rbind(full.quarter.level.violations,
                                       quarter.level.violations %>% 
                                        select(mine_id, quarter = cal_qtr,
                                               year = cal_yr, viol.quantity = quarter_level_violation) %>% 
                                        data.frame() %>% filter(year >= start.year) %>% 
                                        filter(year <= end.year)) %>% arrange(mine_id, year, quarter)

## Challenge: Calculating past period statistics
## lag will put value of index n - 1 at index n
## roll_sum will do a sum from the specified number of previous rows and put NA's at the beginning
## We filter with actual.start.year so there are no NA's. 
##   (The year that data is ready and the year that data needs to start being used are different.)
full.quarter.level.violations <- full.quarter.level.violations %>% 
  group_by(mine_id) %>%
  mutate(last.quarter = lag(viol.quantity),
         last.year = roll_sum(lag(viol.quantity), 4, align = "right", fill = NA),
         last.three.years = roll_sum(lag(viol.quantity), 12, align = "right", fill = NA)) %>%
  filter(year >= actual.start.year)

full.quarter.level.violations

```
```{r Compiling data}
quarter.days_lost.violation <- merge(full.quarter.level.days_lost, full.quarter.level.violations , by = c("mine_id","year","quarter"),all=TRUE)
quarter.days_lost.violation[is.na(test)] <- 0
quarter.days_lost.violation
```

```{r Saving Data into RData files}
write.csv(full.quarter.level.days_lost, "./A New Hope/quarter_level_days_lost.csv")
write.csv(full.quarter.level.violations, "./A New Hope/quarter_level_violations.csv")
write.csv(quarter.days_lost.violation, "./A New Hope/quarter_days_lost_violation.csv")

save(full.quarter.level.days_lost,file = "./A New Hope/quarter_level_days_lost.RData")
save(full.quarter.level.violations, file = "./A New Hope/quarter_level_violations.RData")
save(quarter.days_lost.violation, file = "./A New Hope/quarter_days_lost_violation.RData")
```

```{r}
test <- quarter.days_lost.violation %>% select(num.days.lost,last.quarter.x,last.year.x,last.three.years.x,last.quarter.y,last.year.y, last.three.years.y)

pairs(test)
```
removing rows containing more than 50% zeros 
```{r}

reduced_both <- lm(num.days.lost~last.quarter.x
                 +last.year.x
                 +last.three.years.x
                 +last.quarter.y
                 +last.year.y
                 +last.three.years.y,
                 data = quarter.days_lost.violation)
summary(reduced_both)
```
```{r}
reduce_days_lost <- lm(num.days.lost~last.quarter.x
                 +last.year.x
                 +last.three.years.x,
                 data = quarter.days_lost.violation)
summary(reduce_days_lost)
```

```{r}
reduce_days_lost <- lm(num.days.lost~factor(last.year.x),
                 data = quarter.days_lost.violation)
summary(reduce_days_lost)

plot(num.days.lost~last.quarter.x,data=quarter.days_lost.violation)
```

```{r}
temp <- quarter.days_lost.violation %>% filter(last.quarter.x != 0, last.quarter.y != 0,
                                                 last.year.x != 0, last.year.y != 0,
                                                 last.three.years.x != 0, last.three.years.y != 0)
plot(num.days.lost~last.quarter.x,data=temp)

reduced_both_remove_0 <- lm(num.days.lost~last.quarter.x
                 +last.year.x
                 +last.three.years.x
                 +last.quarter.y
                 +last.year.y
                 +last.three.years.y,
                 data = temp)
summary(reduced_both_remove_0)

```

```{r}
over_200 <- quarter.days_lost.violation %>% filter(last.quarter.x > 200)
under_200 <- quarter.days_lost.violation %>% filter(last.quarter.x < 200)

#historgram of over 200 days lost from last year

qplot(num.days.lost, data=over_200, geom="histogram")
plot(num.days.lost ~ last.quarter.x,data = over_200)
plot(num.days.lost ~ last.quarter.x,data = under_200)

```

```{r roll_over}
roll_over <- function(data, quarter.level, longest.period, actual.start.year,end.year){
  start.year <- actual.start.year-longest.period

  number.years <-  end.year - start.year + 1

  mines.with.accidents.post.start.year <- (data %>% filter(cal_yr >= actual.start.year))$mine_id %>% unique()

  blank.full.quarter.level.days_lost <- data.frame(mine_id = rep(mines.with.accidents.post.start.year, each = 4 * number.years),
                                                 quarter = rep(1:4, times = length(mines.with.accidents.post.start.year) * number.years),
                                                 year = rep(rep(start.year:end.year, times = rep(4, number.years)),
                                                            times = length(mines.with.accidents.post.start.year)),
                                                 num.days.lost = 0)
  full.quarter.level.days_lost <- anti_join(blank.full.quarter.level.days_lost, quarter.level,
                                            by = c("mine_id" = "mine_id", "quarter" = "cal_qtr", "year" = "cal_yr"))
  full.quarter.level.days_lost <- rbind(full.quarter.level.days_lost,
                                        quarter.level %>% 
                                          select(mine_id, quarter = cal_qtr,
                                                 year = cal_yr, num.days.lost = base) %>% 
                                          data.frame() %>% filter(year >= start.year) %>% 
                                          filter(year <= end.year)) %>% arrange(mine_id, year, quarter)

  ## Challenge: Calculating past period statistics
  ## lag will put value of index n - 1 at index n
  ## roll_sum will do a sum from the specified number of previous rows and put NA's at the beginning
  ## We filter with actual.start.year so there are no NA's. 
  ##   (The year that data is ready and the year that data needs to start being used are different.)
  full.quarter.level.days_lost <- full.quarter.level.days_lost %>% 
                                  group_by(mine_id) %>%
                                  mutate(last.quarter = lag(num.days.lost),
                                         last.year = roll_sum(lag(num.days.lost), 4, align = "right", fill = NA),
                                         last.three.years = roll_sum(lag(num.days.lost), 12, align = "right", fill = NA)) %>%
                                  filter(year >= actual.start.year)

  return(full.quarter.level.days_lost)
}
```

```{r full quarter}

longest.period <- 3
actual.start.year <- 2000
end.year <- 2015

quarter.level.accidents <- actual.accidents %>% group_by(mine_id, cal_qtr, cal_yr)%>% 
                                            summarize(base = sum(days_lost, na.rm = T))
accident_rollover <- roll_over(actual.accidents,
                               quarter.level.accidents,
                               longest.period,
                               actual.start.year,
                               end.year)
accident_rollover

quarter.level.violations <- violations %>% group_by(mine_id, cal_qtr, cal_yr) %>% 
                                            summarize(base = n())
violation_rollover <- roll_over(violations,
                                quarter.level.violations,
                                longest.period,
                                actual.start.year,
                                end.year)
violation_rollover

all.quarter.level.accidents <- accidents %>% group_by(mine_id, cal_qtr, cal_yr)%>% 
                                            summarize(base = sum(days_lost, na.rm = T) + sum(days_restrict, na.rm = T))
all_accident_rollover <- roll_over(accidents,
                                   all.quarter.level.accidents,
                                   longest.period,
                                   1984,
                                   2016)
all_accident_rollover

all_violation_rollover <- roll_over(accidents,
                                   quarter.level.violations,
                                   longest.period,
                                   1984,
                                   2016)
all_violation_rollover

```

```{r}
quaarter.days_lost.violation <- merge(accident_rollover, violation_rollover, by = c("mine_id","year","quarter"),all=TRUE)
quarter.days_lost.violation[is.na(quarter.days_lost.violation)] <- 0
quarter.days_lost.violation

all_accident_violation_rollover <- merge(all_accident_rollover,all_violation_rollover, by = c("mine_id","year","quarter"),all=TRUE)
all_accident_violation_rollover[is.na(all_accident_violation_rollover)]<-0
all_accident_violation_rollover

#remove start year
for (x in unique(all_accident_rollover$mine_id)){
  temp <- all_accident_violation_rollover %>% filter(mine_id==x)
  i <- 1
  while( sum(temp[i,3:11]) == 0){
    i = i+1
  }
  if(i != 1){
    print(x)
    print(i)
    temp <- temp[-1:i,]
  }
}

  #remove end year
for (mine_id in unique(all_accident_rollover$mine_id)){
  temp <- all_accident_violation_rollover %>% filter(mine_id==x)
  i <- nrow(temp)
  while( sum(temp[i,3:11]) == 0){
    i = i-1
  }
  if(i != nrow(temp)){
    print(x)
    print(i)
    temp <- temp[-i:nrow(temp),]
  }
}

```

### 2017-03-04 A New Hope/Catherine_Analysis.Rmd
Update Catherine_Analysis.Rmd

@@
 
+accidents[which(accidents$inj_degr_desc == "FATALITY"), "days_lost"] <- 300
 all.quarter.level.accidents <- accidents %>% group_by(mine_id, cal_qtr, cal_yr)%>% 
@@
                                    longest.period,
-                                   1984,
-                                   2016)
+                                   2000,
+                                   2015)
 all_accident_rollover
@@
                                    longest.period,
-                                   1984,
-                                   2016)
+                                   2000,
+                                   2015)
 all_violation_rollover
@@
 
-#remove start year
+all_accident_violation_cleaned <- data.frame()
+counter <- 1
 for (x in unique(all_accident_rollover$mine_id)){
+  print(counter)
+  counter = counter +1
   temp <- all_accident_violation_rollover %>% filter(mine_id==x)
   i <- 1
-  while( sum(temp[i,3:11]) == 0){
+  while( sum(temp[i,4:11],na.rm = TRUE) == 0){
     i = i+1
@@
   if(i != 1){
-    print(x)
-    print(i)
-    temp <- temp[-1:i,]
-  }
+    temp <- temp[-c(1:15),]
   }
 
-  #remove end year
-for (mine_id in unique(all_accident_rollover$mine_id)){
-  temp <- all_accident_violation_rollover %>% filter(mine_id==x)
   i <- nrow(temp)
-  while( sum(temp[i,3:11]) == 0){
+  while( sum(temp[i,4:11]) == 0){
     i = i-1
@@
   if(i != nrow(temp)){
-    print(x)
-    print(i)
-    temp <- temp[-i:nrow(temp),]
+    temp <- temp[-c(nrow(temp)-i:nrow(temp)),]
   }
+  all_accident_violation_cleaned <- rbind(all_accident_violation_cleaned,temp)
 }
-
+all_accident_violation_cleaned
 ```

### 2017-03-04 A New Hope/Catherine_Analysis.Rmd
Update Catherine_Analysis.Rmd

@@
 counter <- 1
-for (x in unique(all_accident_rollover$mine_id)){
+for (x in unique(all_accident_rollover$mine_id)[599:11830]){
   print(counter)
   counter = counter +1
+
   temp <- all_accident_violation_rollover %>% filter(mine_id==x)
+  for ( i in 4:11){
+    temp[,i] = as.numeric(temp[,i])
+  }
+
+  if(sum(temp[,4:11]) == 0) next
+
+  #beginning
   i <- 1
@@
   if(i != 1){
-    temp <- temp[-c(1:15),]
+    temp <- temp[-c(1:i),]
   }
 
+  #ending
   i <- nrow(temp)
@@
   }
+
   all_accident_violation_cleaned <- rbind(all_accident_violation_cleaned,temp)
+
 }
+
 all_accident_violation_cleaned
+temp
+
 ```

### 2017-03-08 A New Hope/separate_fatality.R
separate num.days.lost and fatality

@@ -1,0 +1,85 @@
+rm(list=ls())
+# Upper Big Branch: 4608436
+
+# setup
+require(dplyr)
+library(RcppRoll)
+setwd("~/Git/violation-data-analysis")
+
+# load accidents, violations, and function roll_over
+if(!exists("accidents") | !exists("violations") | !exists("roll_over")) {
+  load("./Analysis/data/msha_accidents_violations.RData")
+  print("raw RData loaded")
+}
+
+# constant parameters
+longest.period <- 3
+actual.start.year <- 2000
+end.year <- 2015
+
+# generate accident_roll_over and death_roll_over
+actual.accidents <- accidents %>% filter(is.element(inj_degr_desc, c("DAYS AWAY FROM WORK ONLY", 
+                                                                     "DYS AWY FRM WRK & RESTRCTD ACT", 
+                                                                     "DAYS RESTRICTED ACTIVITY ONLY", 
+                                                                     "FATALITY", 
+                                                                     "PERM TOT OR PERM PRTL DISABLTY", 
+                                                                     "NO VALUE FOUND")))
+actual.accidents["deaths"] <- 0
+na.and.no.value.index <- is.element(actual.accidents$inj_degr_desc, "NO VALUE FOUND") & is.na(actual.accidents$days_lost)
+actual.accidents[na.and.no.value.index, "days_restrict"] <- 0
+actual.accidents[na.and.no.value.index, "days_lost"] <- 0
+actual.accidents[which(actual.accidents$inj_degr_desc == "FATALITY"), "deaths"] <- 1
+actual.accidents$ai_dt_actual_date <- strftime(actual.accidents$ai_dt, "%F")
+rm(na.and.no.value.index)
+
+quarter.level.num.days.lost <- actual.accidents %>% group_by(mine_id, cal_qtr, cal_yr)%>% 
+  summarize(base = sum(days_lost, na.rm = T))
+days_lost_rollover <- roll_over(actual.accidents,
+                                quarter.level.num.days.lost,
+                                longest.period,
+                                actual.start.year,
+                                end.year)
+rm(quarter.level.num.days.lost)
+
+quarter.level.num.days.restrict <- actual.accidents %>% group_by(mine_id, cal_qtr, cal_yr)%>% 
+  summarize(base = sum(days_restrict, na.rm = T))
+days_restrict_rollover <- roll_over(actual.accidents,
+                                quarter.level.num.days.restrict,
+                                longest.period,
+                                actual.start.year,
[…]

### 2017-03-08 A New Hope/separate_fatality.R
cleaned and separated data
with num.days.lost, num.days.restrict, num.deaths, and viol.quantity

@@ -1,85 +1,0 @@
-rm(list=ls())
-# Upper Big Branch: 4608436
-
-# setup
-require(dplyr)
-library(RcppRoll)
-setwd("~/Git/violation-data-analysis")
-
-# load accidents, violations, and function roll_over
-if(!exists("accidents") | !exists("violations") | !exists("roll_over")) {
-  load("./Analysis/data/msha_accidents_violations.RData")
-  print("raw RData loaded")
-}
-
-# constant parameters
-longest.period <- 3
-actual.start.year <- 2000
-end.year <- 2015
-
-# generate accident_roll_over and death_roll_over
-actual.accidents <- accidents %>% filter(is.element(inj_degr_desc, c("DAYS AWAY FROM WORK ONLY", 
-                                                                     "DYS AWY FRM WRK & RESTRCTD ACT", 
-                                                                     "DAYS RESTRICTED ACTIVITY ONLY", 
-                                                                     "FATALITY", 
-                                                                     "PERM TOT OR PERM PRTL DISABLTY", 
-                                                                     "NO VALUE FOUND")))
-actual.accidents["deaths"] <- 0
-na.and.no.value.index <- is.element(actual.accidents$inj_degr_desc, "NO VALUE FOUND") & is.na(actual.accidents$days_lost)
-actual.accidents[na.and.no.value.index, "days_restrict"] <- 0
-actual.accidents[na.and.no.value.index, "days_lost"] <- 0
-actual.accidents[which(actual.accidents$inj_degr_desc == "FATALITY"), "deaths"] <- 1
-actual.accidents$ai_dt_actual_date <- strftime(actual.accidents$ai_dt, "%F")
-rm(na.and.no.value.index)
-
-quarter.level.num.days.lost <- actual.accidents %>% group_by(mine_id, cal_qtr, cal_yr)%>% 
-  summarize(base = sum(days_lost, na.rm = T))
-days_lost_rollover <- roll_over(actual.accidents,
-                                quarter.level.num.days.lost,
-                                longest.period,
-                                actual.start.year,
-                                end.year)
-rm(quarter.level.num.days.lost)
-
-quarter.level.num.days.restrict <- actual.accidents %>% group_by(mine_id, cal_qtr, cal_yr)%>% 
-  summarize(base = sum(days_restrict, na.rm = T))
-days_restrict_rollover <- roll_over(actual.accidents,
-                                quarter.level.num.days.restrict,
-                                longest.period,
-                                actual.start.year,
[…]

### 2017-03-08 San Antonio/separate_fatality.R

@@ -1,0 +1,83 @@
+rm(list=ls())
+# Upper Big Branch: 4608436
+# This is a temporary fix
+# Inputs: raw accidents data and final data with only active mine-quarters
+# Output: days lost, days restrict, deaths, and violations with only active mine-quarters
+
+# setup
+require(dplyr)
+library(RcppRoll)
+setwd("~/Git/violation-data-analysis")
+
+# load accidents, violations, and function roll_over
+if(!exists("accidents") | !exists("full.days_lost.accidents.date") | !exists("roll_over")) {
+  load("./San Antonio/raw_accidents_fine_viol.RData")
+  print("raw RData loaded")
+}
+
+# constant parameters
+longest.period <- 3
+actual.start.year <- 2000
+end.year <- 2015
+
+# generate accident_roll_over and death_roll_over
+actual.accidents <- accidents %>% filter(is.element(inj_degr_desc, c("DAYS AWAY FROM WORK ONLY", 
+                                                                     "DYS AWY FRM WRK & RESTRCTD ACT", 
+                                                                     "DAYS RESTRICTED ACTIVITY ONLY", 
+                                                                     "FATALITY", 
+                                                                     "PERM TOT OR PERM PRTL DISABLTY", 
+                                                                     "NO VALUE FOUND")))
+actual.accidents["deaths"] <- 0
+na.and.no.value.index <- is.element(actual.accidents$inj_degr_desc, "NO VALUE FOUND") & is.na(actual.accidents$days_lost)
+actual.accidents[na.and.no.value.index, "days_restrict"] <- 0
+actual.accidents[na.and.no.value.index, "days_lost"] <- 0
+actual.accidents[which(actual.accidents$inj_degr_desc == "FATALITY"), "deaths"] <- 1
+actual.accidents$ai_dt_actual_date <- strftime(actual.accidents$ai_dt, "%F")
+rm(na.and.no.value.index)
+
+quarter.level.num.days.lost <- actual.accidents %>% group_by(mine_id, cal_qtr, cal_yr)%>% 
+  summarize(base = sum(days_lost, na.rm = T))
+days_lost_rollover <- roll_over(actual.accidents,
+                                quarter.level.num.days.lost,
+                                longest.period,
+                                actual.start.year,
+                                end.year)
+rm(quarter.level.num.days.lost)
+
[…]

### 2017-03-08 A New Hope/yl2750.R
none

@@ -1,0 +1,108 @@
+library(ggplot2)
+library(dplyr)
+library(plm)
+
+setwd("~/Git/violation-data-analysis/A New Hope")
+load("days_lost.accident.Rdata")
+
+test.year.quarter <- full.days_lost.accidents.date %>% rowwise() %>%
+  mutate(year = year + quarter / 5)
+result <- plm(num.days.lost ~ last.year.x, index = c("mine_id", "year"), model = "within", data = test.year.quarter)
+summary(result)
+
+result <- lm(num.days.lost~last.quarter.x + last.year.x + last.three.years.x + last.quarter.y + last.year.y + last.three.years.y, data = full.days_lost.accidents.date)
+summary(result)
+
+result <- lm(num.days.lost~last.year.x+mine_id, data=full.days_lost.accidents.date)
+summary(result)
+
+# http://stats.stackexchange.com/questions/79723/interpretation-of-r2-in-fixed-effects-panel-regression
+reg.fe <- plm(num.days.lost~last.year.x, data = full.days_lost.accidents.date, index="mine_id",model = "within")
+summary(reg.fe)
+length(unique(test$mine_id))
+
+test.nonezero <- test %>% filter(num.days.lost >0 & last.year.x >0)
+reg = lm(num.days.lost~last.quarter.x+last.quarter.y+last.year.x+last.year.y+last.three.years.x+last.three.years.y,data=test)
+summary(reg)
+
+new.data.high <- test %>% 
+  filter(num.days.lost >= 250) %>%
+  select(mine_id, year, quarter, num.days.lost, last.year.x, last.year.y)
+
+new.data.low <- test %>% 
+  filter(num.days.lost < 250 & num.days.lost > 0) %>%
+  select(mine_id, year, quarter, num.days.lost, last.year.x, last.year.y)
+
+new.data <- test %>% 
+  filter(num.days.lost >0) %>%
+  select(mine_id, year, quarter, num.days.lost, last.year.x, last.year.y)
+
+
+qplot(last.year.x, num.days.lost, data=new.data.low, color='green')
+
+qplot(new.data.high$last.year.y, geom="histogram",binwidth=10,xlim=c(0,1000))
+
+new.data.high$sr <- 'high'
+new.data.low$sr <- 'low'
+both = rbind(new.data.low,new.data.high)
[…]

### 2017-03-08 San Antonio/separate_fatality.R
reordered and renamed columns

@@ -33,7 +33,7 @@ actual.accidents[na.and.no.value.index, "days_restrict"] <- 0
 actual.accidents[na.and.no.value.index, "days_lost"] <- 0
 actual.accidents[which(actual.accidents$inj_degr_desc == "FATALITY"), "deaths"] <- 1
 actual.accidents$ai_dt_actual_date <- strftime(actual.accidents$ai_dt, "%F")
-rm(na.and.no.value.index)
+rm(na.and.no.value.index, accidents)
 
 quarter.level.num.days.lost <- actual.accidents %>% group_by(mine_id, cal_qtr, cal_yr)%>% 
   summarize(base = sum(days_lost, na.rm = T))
@@ -67,17 +67,27 @@ rm(quarter.level.num.deaths)
 # cleanup violations
 active_violation_rollover = full.days_lost.accidents.date %>%
   select(mine_id, year, quarter, viol.quantity, last.quarter.y, last.year.y, last.three.years.y)
+active_violation_rollover$active <- TRUE
+
 
 # join accidents and violation
 temp <- merge(days_lost_rollover, days_restrict_rollover, by = c("mine_id","year","quarter"),all=TRUE)
 temp[is.na(temp)] <- 0
 temp <- merge(temp, death_rollover, by = c("mine_id","year","quarter"),all=TRUE)
 temp[is.na(temp)] <- 0
-temp <- merge(temp, active_violation_rollover, by = c("mine_id","year","quarter"),all.y=TRUE)
+
+temp <- merge(temp, active_violation_rollover, by = c("mine_id","year","quarter"),all.x=TRUE)
+temp[is.na(temp$active),]$active <- FALSE
 temp[is.na(temp)] <- 0
 
 # result
 complete.active.quarters <- temp
+colnames(complete.active.quarters) <- 
+  c("mine_id", "year", "quarter", 
+    "num.days.lost", "last.quarter.lost", "last.year.lost", "last.three.years.lost",
+    "num.days.restrict", "last.quarter.restrict", "last.year.restrict", "last.three.years.restrict",
+    "num.days.death", "last.quarter.death", "last.year.death", "last.three.years.death",
+    "viol.quantity", "last.quarter.viol", "last.year.viol", "last.three.years.viol",
+    "active")
+complete.active.quarters <- complete.active.quarters[, c(1:3, 20, 4:19)]
 save(complete.active.quarters, file="./San Antonio/result.RData")
-
-

### 2017-03-09 San Antonio/top_tens.Rmd
nothing specific

---
title: "Top Tens"
author: "Yu"
date: "3/8/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
rm(list=ls())
require(dplyr)
knitr::opts_knit$set(root.dir = "~/Git/violation-data-analysis")
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r top 10 based on num.death}
load("./San Antonio/output/result.RData")
complete.active.quarters %>% arrange(desc(num.death)) %>% select(mine_id, mine.name, year, quarter, num.death) %>% head(10)
```

```{r top 10 based on num.days.lost}
complete.active.quarters %>% filter(mine_id == 101322) %>% select(mine.name, year, quarter, last.year.lost) %>% head(10)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2017-03-09 San Antonio/separate_fatality.R

@@ -10,8 +10,11 @@ library(RcppRoll)
 setwd("~/Git/violation-data-analysis")
 
 # load accidents, violations, and function roll_over
-if(!exists("accidents") | !exists("full.days_lost.accidents.date") | !exists("roll_over")) {
-  load("./San Antonio/raw_accidents_fine_viol.RData")
+if(!exists("accidents") | 
+   !exists("full.days_lost.accidents.date") | 
+   !exists("roll_over") | 
+   !exists("mines")) {
+  load("./San Antonio/data/secret_ingredients.RData")
   print("raw RData loaded")
 }
 
@@ -80,14 +83,24 @@ temp <- merge(temp, active_violation_rollover, by = c("mine_id","year","quarter"
 temp[is.na(temp$active),]$active <- FALSE
 temp[is.na(temp)] <- 0
 
+temp <- merge(temp, mines, by = "mine_id", all.x=TRUE)
+
 # result
 complete.active.quarters <- temp
 colnames(complete.active.quarters) <- 
   c("mine_id", "year", "quarter", 
     "num.days.lost", "last.quarter.lost", "last.year.lost", "last.three.years.lost",
     "num.days.restrict", "last.quarter.restrict", "last.year.restrict", "last.three.years.restrict",
-    "num.days.death", "last.quarter.death", "last.year.death", "last.three.years.death",
+    "num.death", "last.quarter.death", "last.year.death", "last.three.years.death",
     "viol.quantity", "last.quarter.viol", "last.year.viol", "last.three.years.viol",
-    "active")
-complete.active.quarters <- complete.active.quarters[, c(1:3, 20, 4:19)]
-save(complete.active.quarters, file="./San Antonio/result.RData")
+    "active", "mine.name")
+complete.active.quarters <- complete.active.quarters[, c(1, 21, 2:3, 20, 4:19)]
+
+save(complete.active.quarters, file="./San Antonio/output/result.RData")
+
+# http://stats.stackexchange.com/questions/59250/how-to-interpret-the-output-of-the-summary-method-for-an-lm-object-in-r
+summary(lm(25*num.death+1*num.days.lost+0.7*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost
+           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
+           +last.quarter.viol+last.year.viol+last.three.years.viol
+           +last.quarter.death+last.year.death+last.three.years.death, 
+           data=complete.active.quarters %>% filter(active)))

### 2017-03-09 San Antonio/separate_fatality.R
nothing specific either

@@ -98,6 +98,7 @@ complete.active.quarters <- complete.active.quarters[, c(1, 21, 2:3, 20, 4:19)]
 
 save(complete.active.quarters, file="./San Antonio/output/result.RData")
 
+# http://blog.minitab.com/blog/adventures-in-statistics-2/how-to-interpret-regression-analysis-results-p-values-and-coefficients
 # http://stats.stackexchange.com/questions/59250/how-to-interpret-the-output-of-the-summary-method-for-an-lm-object-in-r
 summary(lm(25*num.death+1*num.days.lost+0.7*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost
            +last.quarter.restrict+last.year.restrict+last.three.years.restrict

### 2017-03-10 San Antonio/simple_lm.Rmd
simple linear model
y = num.days.lost or num.death

---
title: "Simple Linear Models"
author: "Yu"
date: "3/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(dplyr)
knitr::opts_knit$set(root.dir = "~/Git/violation-data-analysis")
```

## Simple Linear Model

Read these two links first: [How to interpret regression p-values](http://blog.minitab.com/blog/adventures-in-statistics-2/how-to-interpret-regression-analysis-results-p-values-and-coefficients) and
[How to interpret the output of the summary() method for an lm object](http://stats.stackexchange.com/questions/59250/how-to-interpret-the-output-of-the-summary-method-for-an-lm-object-in-r)

- lm on num of days lost or restricted:
```{r echo=FALSE}
load("./San Antonio/output/result.RData")
num.days.lm <- lm(0*num.death+1*num.days.lost+1*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost
           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
           +last.quarter.viol+last.year.viol+last.three.years.viol
           +last.quarter.death+last.year.death+last.three.years.death, 
           data=complete.active.quarters %>% filter(active))
summary(num.days.lm)
```
- lm on num of deaths:
```{r echo=FALSE}
summary(lm(num.death~last.three.years.lost+last.quarter.viol+last.year.viol, data=complete.active.quarters %>% filter(active)))
```

### 2017-03-10 San Antonio/top_tens.Rmd
simple linear model
y = num.days.lost or num.death

@@
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
 ```{r top 10 based on num.death}
@@
 
-```{r top 10 based on num.days.lost}
+```{r top 10 based on certain mine}
 complete.active.quarters %>% filter(mine_id == 101322) %>% select(mine.name, year, quarter, last.year.lost) %>% head(10)
+```
+
+```{r simple lm}
+load("./San Antonio/output/result.RData")
+summary(lm(1*num.death+0*num.days.lost+0*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost
+           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
+           +last.quarter.viol+last.year.viol+last.three.years.viol
+           +last.quarter.death+last.year.death+last.three.years.death, 
+           data=complete.active.quarters %>% filter(active)))
+
+summary(lm(num.death~last.three.years.lost+last.quarter.viol+last.year.viol, data=complete.active.quarters %>% filter(active)))
 ```

### 2017-03-10 San Antonio/separate_fatality.R

@@ -98,10 +98,4 @@ complete.active.quarters <- complete.active.quarters[, c(1, 21, 2:3, 20, 4:19)]
 
 save(complete.active.quarters, file="./San Antonio/output/result.RData")
 
-# http://blog.minitab.com/blog/adventures-in-statistics-2/how-to-interpret-regression-analysis-results-p-values-and-coefficients
-# http://stats.stackexchange.com/questions/59250/how-to-interpret-the-output-of-the-summary-method-for-an-lm-object-in-r
-summary(lm(25*num.death+1*num.days.lost+0.7*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost
-           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
-           +last.quarter.viol+last.year.viol+last.three.years.viol
-           +last.quarter.death+last.year.death+last.three.years.death, 
-           data=complete.active.quarters %>% filter(active)))
+# see simple_lm.R for simple lm

### 2017-03-10 San Antonio/simple_lm.R

@@ -1,0 +1,15 @@
+# some resources:
+# http://blog.minitab.com/blog/adventures-in-statistics-2/how-to-interpret-regression-analysis-results-p-values-and-coefficients
+# http://stats.stackexchange.com/questions/59250/how-to-interpret-the-output-of-the-summary-method-for-an-lm-object-in-r
+
+# setup
+require(dplyr)
+setwd("~/Git/violation-data-analysis")
+load("./San Antonio/output/result.RData")
+
+# lm on composite num.days.lost
+summary(lm(30*num.death+1*num.days.lost+1*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost
+           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
+           +last.quarter.viol+last.year.viol+last.three.years.viol
+           +last.quarter.death+last.year.death+last.three.years.death, 
+           data=complete.active.quarters %>% filter(active)))

### 2017-03-12 San Antonio/presentation_materials.Rmd
presentation materials

---
title: "Presentation Materials"
author: "Yu"
date: "3/12/2017"
output: 
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(dplyr)
knitr::opts_knit$set(root.dir = "~/Git/violation-data-analysis")
```

## Upper Big Branch and Other Deadliest Mine

Below is a list of the top 10 mine-quarters based on number of deaths:

```{r echo=FALSE}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
top.deaths <- complete.active.quarters %>% filter(active) %>% arrange(desc(num.death)) %>% select(mine.name, mine_id, year, quarter, num.death)
head(top.deaths, 10)
```
Mine name: [The Upper Big Branch Mine (4608436)](https://en.wikipedia.org/wiki/Upper_Big_Branch_Mine_disaster)

```{r echo=FALSE}
library(ggplot2)
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
upper.big.branch <- complete.active.quarters %>% filter(mine_id == 4608436 & (year < 2010 | (year == 2010 & quarter < 2))) %>%
  select(year, quarter, num.days.lost, viol.quantity)
# ggplot() +
#   geom_line(data = upper.big.branch, aes(x = year + quarter / 4, y = viol.quantity), color = "red") +
#   geom_line(data = upper.big.branch, aes(x = year + quarter / 4, y = num.days.lost), color = "green")
qplot(x=year + quarter / 4, y=viol.quantity, data=upper.big.branch, geom = "line",
     main = "Upper Big Branch: Number of Quarterly Violations",
     xlab = "Year", ylab = "Count")
```

Mine name: [No. 5 Mine (101322)](http://www.chicagotribune.com/news/chi-0209220285sep22-story.html)

```{r echo=FALSE}
library(ggplot2)
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
no.5.mine <- complete.active.quarters %>% filter(mine_id == 101322 & (year < 2001 | (year == 2001 & quarter < 3))) %>%
  select(year, quarter, num.days.lost, viol.quantity)
qplot(x=year + quarter / 4, y=viol.quantity, data=no.5.mine, geom = "line",
     main = "No. 5 Mine: Number of Quarterly Violations",
     xlab = "Year", ylab = "Count")
```

Mine name: [Darby Mine No. 1 (1518185)](http://www.chicagotribune.com/news/chi-0209220285sep22-story.html)

```{r echo=FALSE}
library(ggplot2)
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
darby.mine <- complete.active.quarters %>% filter(mine_id == 1518185 & (year < 2006 | (year == 2006 & quarter < 2))) %>%
  select(year, quarter, num.days.lost, viol.quantity)
qplot(x=year + quarter / 4, y=viol.quantity, data=darby.mine, geom = "line",
     main = "Darby Mine No. 1: Number of Quarterly Violations",
     xlab = "Year", ylab = "Count")
```

Mine name: [Gibson Mine	(1202215)](http://coaldiver.org/msha-id/1202215)

```{r echo=FALSE}
library(ggplot2)
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
gibson.mine <- complete.active.quarters %>% filter(mine_id == 1202215 & (year < 2007 | (year == 2007 & quarter < 3))) %>%
  select(year, quarter, num.days.lost, viol.quantity)
qplot(x=year + quarter / 4, y=viol.quantity, data=gibson.mine, geom = "line",
     main = "Gibson Mine: Number of Quarterly Violations",
     xlab = "Year", ylab = "Count")
```

## Number of Days Lost not Implying Severity (e.g., New Era Mine)
Link: https://arlweb.msha.gov/drs/ASP/MineAction.asp

Mine name: [The American Coal Company New Era Mine (1102752)](http://www.stltoday.com/business/local/murray-energy-to-close-galatia-mine-next-year/article_aecf6963-c5ed-5edb-abd1-db6c5537a540.html)

Among the top 10 mine-quarters based on number of days lost,
```{r echo=FALSE}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
top.days.lost <- complete.active.quarters %>% filter(active) %>% arrange(desc(num.days.lost)) %>% select(mine.name, year, quarter, num.days.lost)
head(top.days.lost, 10)
```
the "American Coal Company New Era Mine" take 5 spots.
Let's take a look at its second quarter of 2005:
```{r echo=FALSE}
if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
new.era <- accidents %>% 
  filter(mine_id == 1102752 & cal_yr == 2005 & cal_qtr == 2) %>% 
  arrange(desc(days_lost)) %>% 
  select(
    ai_dt, 
    days_lost, 
    ai_class_desc
    #ai_occ_desc 
    #ai_narr
    ) 
head(new.era)
```
In particular, the top accident was:
```{r echo=FALSE}
if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
new.era.top.accidents <- accidents %>% 
  filter(mine_id == 1102752 & cal_yr == 2005 & cal_qtr == 2) %>% 
  arrange(desc(days_lost)) %>% 
  select(
    ai_dt, 
    days_lost, 
    ai_class_desc,
    ai_occ_desc, 
    ai_narr
    ) 
print(as.character(new.era.top.accidents$ai_narr[1]))
```
and the second to top accident was:
```{r echo=FALSE}
if(!exists("new.era.top.accidents")){
 if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
new.era.top.accidents <- accidents %>% 
  filter(mine_id == 1102752 & cal_yr == 2005 & cal_qtr == 2) %>% 
  arrange(desc(days_lost)) %>% 
  select(
    ai_dt, 
    days_lost, 
    ai_class_desc,
    ai_occ_desc, 
    ai_narr
    )  
}
print(as.character(new.era.top.accidents$ai_narr[2]))
```

## Simple Linear Model

The following linear model tests a few predictors and their ability to predict number of days lost or restircted:
```{r echo=FALSE}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
num.days.lm <- lm(0*num.death+1*num.days.lost+1*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost
           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
           +last.quarter.viol+last.year.viol+last.three.years.viol
           +last.quarter.death+last.year.death+last.three.years.death, 
           data=complete.active.quarters %>% filter(active))
summary(num.days.lm)
```
The following linear model tests a few predictors and their ability to predict number of deaths:
```{r echo=FALSE}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
num.deaths.lm <- lm(1*num.death+0*num.days.lost+0*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost
           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
           +last.quarter.viol+last.year.viol+last.three.years.viol
           +last.quarter.death+last.year.death+last.three.years.death, 
           data=complete.active.quarters %>% filter(active))
summary(num.deaths.lm)
```

### 2017-03-12 San Antonio/simple_lm.R

@@ -4,8 +4,39 @@
 
 # setup
 require(dplyr)
+require(survival)
+require(ggplot2)
 setwd("~/Git/violation-data-analysis")
 load("./San Antonio/output/result.RData")
+
+train <- complete.active.quarters
+# train <- train %>% rowwise() %>% mutate(high_severity = num.death > 0 | num.days.lost > 100)
+# train <- train[, c(1:5, 22, 6:21)]
+# train %>% count(high_severity)
+
+View(train %>% filter(mine_id == 1102752))
+new_era <- accidents %>% filter(mine_id == 1102752 & cal_qtr == 2 & cal_yr == 2005) %>% arrange(
+  desc(days_lost)
+)
+View(new_era)
+
+summary(lm(num.days.lost + num.days.restrict ~ last.quarter.viol+last.year.viol+last.three.years.viol, data=train %>% filter(active)))
+
+bad <- train %>% filter(active & num.days.restrict >= 30 & last.quarter.lost > 0)
+bad$mark <- 'bad'
+good <- train %>% filter(active & num.days.restrict < 30 & last.quarter.lost > 0)
+good$mark <- 'good'
+goodbad <- rbind(good, bad)
+goodbad <- goodbad %>% rowwise() %>% mutate(ratio = last.quarter.lost / (last.year.lost / 4))
+ggplot(goodbad, aes(num.days.lost)) + geom_density(alpha = 0.2)
+plot(ecdf(x = train$num.days.lost))
+
+# msha_glm <- clogit(high_severity ~ num.days.lost + viol.quantity + strata(mine_id),
+#                    data = train%>% top_n(100) %>% filter(active) %>% select(mine_id, num.days.lost, high_severity, viol.quantity))
+# 
+# print(msha_glm)
+
+sapply(train, function(x) sum(is.na(x)))
 
 # lm on composite num.days.lost
 summary(lm(30*num.death+1*num.days.lost+1*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost

### 2017-03-13 San Antonio/presentation_materials.Rmd
added perm. dis. San Antonio result

@@
 ```
-Mine name: [The Upper Big Branch Mine (4608436)](https://en.wikipedia.org/wiki/Upper_Big_Branch_Mine_disaster)
+[The Upper Big Branch Mine (4608436)](https://en.wikipedia.org/wiki/Upper_Big_Branch_Mine_disaster)
 
@@
 
-Mine name: [No. 5 Mine (101322)](http://www.chicagotribune.com/news/chi-0209220285sep22-story.html)
+[No. 5 Mine (101322)](http://www.chicagotribune.com/news/chi-0209220285sep22-story.html)
 
@@
 
-Mine name: [Darby Mine No. 1 (1518185)](http://www.chicagotribune.com/news/chi-0209220285sep22-story.html)
+[Darby Mine No. 1 (1518185)](http://www.chicagotribune.com/news/chi-0209220285sep22-story.html)
 
@@
 
-Mine name: [Gibson Mine	(1202215)](http://coaldiver.org/msha-id/1202215)
+[Gibson Mine	(1202215)](http://coaldiver.org/msha-id/1202215)
 
@@
 ## Number of Days Lost not Implying Severity (e.g., New Era Mine)
-Link: https://arlweb.msha.gov/drs/ASP/MineAction.asp
-
-Mine name: [The American Coal Company New Era Mine (1102752)](http://www.stltoday.com/business/local/murray-energy-to-close-galatia-mine-next-year/article_aecf6963-c5ed-5edb-abd1-db6c5537a540.html)
+[The American Coal Company New Era Mine (1102752)](http://www.stltoday.com/business/local/murray-energy-to-close-galatia-mine-next-year/article_aecf6963-c5ed-5edb-abd1-db6c5537a540.html)
 

### 2017-03-13 San Antonio/separate_fatality.R

@@ -31,10 +31,12 @@ actual.accidents <- accidents %>% filter(is.element(inj_degr_desc, c("DAYS AWAY 
                                                                      "PERM TOT OR PERM PRTL DISABLTY", 
                                                                      "NO VALUE FOUND")))
 actual.accidents["deaths"] <- 0
+actual.accidents["perm_dis"] <- 0
 na.and.no.value.index <- is.element(actual.accidents$inj_degr_desc, "NO VALUE FOUND") & is.na(actual.accidents$days_lost)
 actual.accidents[na.and.no.value.index, "days_restrict"] <- 0
 actual.accidents[na.and.no.value.index, "days_lost"] <- 0
 actual.accidents[which(actual.accidents$inj_degr_desc == "FATALITY"), "deaths"] <- 1
+actual.accidents[which(actual.accidents$inj_degr_desc == "PERM TOT OR PERM PRTL DISABLTY"), "perm_dis"] <- 1
 actual.accidents$ai_dt_actual_date <- strftime(actual.accidents$ai_dt, "%F")
 rm(na.and.no.value.index, accidents)
 
@@ -45,6 +47,10 @@ days_lost_rollover <- roll_over(actual.accidents,
                                 longest.period,
                                 actual.start.year,
                                 end.year)
+colnames(days_lost_rollover) <- c(
+  "mine_id", "year", "quarter", "num.days.lost",
+  "last.quarter.lost", "last.year.lost", "last.three.years.lost"
+  )
 rm(quarter.level.num.days.lost)
 
 quarter.level.num.days.restrict <- actual.accidents %>% group_by(mine_id, cal_qtr, cal_yr)%>% 
@@ -54,7 +60,10 @@ days_restrict_rollover <- roll_over(actual.accidents,
                                 longest.period,
                                 actual.start.year,
                                 end.year)
-colnames(days_restrict_rollover)[4] <- "num.days.restrict"
+colnames(days_restrict_rollover) <- c(
+  "mine_id", "year", "quarter", 
+  "num.days.restrict", "last.quarter.restrict", "last.year.restrict", "last.three.years.restrict"
+)
 rm(quarter.level.num.days.restrict)
 
 quarter.level.num.deaths <- actual.accidents %>% group_by(mine_id, cal_qtr, cal_yr)%>% 
@@ -64,13 +73,34 @@ death_rollover <- roll_over(actual.accidents,
[…]

### 2017-03-13 San Antonio/separate_fatality.R
same as the last commit, with bug fixed

@@ -48,8 +48,8 @@ days_lost_rollover <- roll_over(actual.accidents,
                                 actual.start.year,
                                 end.year)
 colnames(days_lost_rollover) <- c(
-  "mine_id", "year", "quarter", "num.days.lost",
-  "last.quarter.lost", "last.year.lost", "last.three.years.lost"
+  "mine_id", "quarter", "year",
+  "num.days.lost", "last.quarter.lost", "last.year.lost", "last.three.years.lost"
   )
 rm(quarter.level.num.days.lost)
 
@@ -61,7 +61,7 @@ days_restrict_rollover <- roll_over(actual.accidents,
                                 actual.start.year,
                                 end.year)
 colnames(days_restrict_rollover) <- c(
-  "mine_id", "year", "quarter", 
+  "mine_id", "quarter", "year",
   "num.days.restrict", "last.quarter.restrict", "last.year.restrict", "last.three.years.restrict"
 )
 rm(quarter.level.num.days.restrict)
@@ -74,7 +74,7 @@ death_rollover <- roll_over(actual.accidents,
                                actual.start.year,
                                end.year)
 colnames(death_rollover) <- c(
-  "mine_id", "year", "quarter", 
+  "mine_id", "quarter", "year",
   "num.death", "last.quarter.death", "last.year.death", "last.three.years.death"
 )
 rm(quarter.level.num.deaths)
@@ -87,7 +87,7 @@ dis_rollover <- roll_over(actual.accidents,
                             actual.start.year,
                             end.year)
 colnames(dis_rollover) <- c(
-  "mine_id", "year", "quarter", 
+  "mine_id", "quarter", "year", 
   "num.dis", "last.quarter.dis", "last.year.dis", "last.three.years.dis"
 )
 rm(quarter.level.num.perm_dis)

### 2017-03-13 San Antonio/simple_lm.R

@@ -36,7 +36,7 @@ plot(ecdf(x = train$num.days.lost))
 # 
 # print(msha_glm)
 
-sapply(train, function(x) sum(is.na(x)))
+sapply(complete.active.quarters, function(x) sum(is.na(x)))
 
 # lm on composite num.days.lost
 summary(lm(30*num.death+1*num.days.lost+1*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost

### 2017-03-13 San Antonio/separate_fatality.R
minor bug fixed

@@ -32,13 +32,13 @@ actual.accidents <- accidents %>% filter(is.element(inj_degr_desc, c("DAYS AWAY 
                                                                      "NO VALUE FOUND")))
 actual.accidents["deaths"] <- 0
 actual.accidents["perm_dis"] <- 0
-na.and.no.value.index <- is.element(actual.accidents$inj_degr_desc, "NO VALUE FOUND") & is.na(actual.accidents$days_lost)
-actual.accidents[na.and.no.value.index, "days_restrict"] <- 0
-actual.accidents[na.and.no.value.index, "days_lost"] <- 0
+# na.and.no.value.index <- is.element(actual.accidents$inj_degr_desc, "NO VALUE FOUND") & is.na(actual.accidents$days_lost)
+actual.accidents[is.na(actual.accidents$days_lost), "days_lost"] <- 0
+actual.accidents[is.na(actual.accidents$days_restrict), "days_restrict"] <- 0
 actual.accidents[which(actual.accidents$inj_degr_desc == "FATALITY"), "deaths"] <- 1
 actual.accidents[which(actual.accidents$inj_degr_desc == "PERM TOT OR PERM PRTL DISABLTY"), "perm_dis"] <- 1
 actual.accidents$ai_dt_actual_date <- strftime(actual.accidents$ai_dt, "%F")
-rm(na.and.no.value.index, accidents)
+# rm(na.and.no.value.index, accidents)
 
 quarter.level.num.days.lost <- actual.accidents %>% group_by(mine_id, cal_qtr, cal_yr)%>% 
   summarize(base = sum(days_lost, na.rm = T))

### 2017-03-13 San Antonio/presentation_materials.Rmd
Major bug fix and simplified active label
Fixed a bug that resulted Sago Mine Disaster being removed from active
mines. Simplify the active label process by assigning TRUE for
none-zero row.

@@
 
-## Upper Big Branch and Other Deadliest Mine
+## Fatal Accidents
 
-Below is a list of the top 10 mine-quarters based on number of deaths:
+### Top 10 Fatal Quarters
 
-```{r echo=FALSE}
+```{r top 10 fatality, echo=FALSE}
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
@@
 ```
-[The Upper Big Branch Mine (4608436)](https://en.wikipedia.org/wiki/Upper_Big_Branch_Mine_disaster)
+### Top 10 Causes of Death
+```{r top 10 causes of death, echo=FALSE}
+if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
+fatal.accidents <- accidents %>% 
+  filter(inj_degr_desc == "FATALITY") %>%
+  group_by(ai_class_desc) %>%
+  summarise(count = n()
+            ) %>%
+  arrange(desc(count))
+head(fatal.accidents, 10)
+```
 
-```{r echo=FALSE}
+### Upper Big Branch and Other Deadliest Mine
+
+#### [Upper Big Branch Mine (4608436)](https://en.wikipedia.org/wiki/Upper_Big_Branch_Mine_disaster)
+
+```{r upper big branch, echo=FALSE}
+# {r upper big branch, echo=FALSE, fig.show = 'hold', out.width = '49%'}
 library(ggplot2)
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-upper.big.branch <- complete.active.quarters %>% filter(mine_id == 4608436 & (year < 2010 | (year == 2010 & quarter < 2))) %>%
-  select(year, quarter, num.days.lost, viol.quantity)
-# ggplot() +
-#   geom_line(data = upper.big.branch, aes(x = year + quarter / 4, y = viol.quantity), color = "red") +
-#   geom_line(data = upper.big.branch, aes(x = year + quarter / 4, y = num.days.lost), color = "green")
-qplot(x=year + quarter / 4, y=viol.quantity, data=upper.big.branch, geom = "line",
-     main = "Upper Big Branch: Number of Quarterly Violations",
-     xlab = "Year", ylab = "Count")
+# upper.big.branch <- complete.active.quarters %>% filter(mine_id == 4608436 & (year < 2010 | (year == 2010 & quarter < 2))) %>%
+#   select(year, quarter, num.days.lost, viol.quantity)
+# # ggplot() +
+# #   geom_line(data = upper.big.branch, aes(x = year + quarter / 4, y = viol.quantity), color = "red") +
+# #   geom_line(data = upper.big.branch, aes(x = year + quarter / 4, y = num.days.lost), color = "green")
+# qplot(x=year + quarter / 4, y=viol.quantity, data=upper.big.branch, geom = "line",
+#      main = "Upper Big Branch: Number of Quarterly Violations",
+#      xlab = "Year", ylab = "Count")
+
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 4608436 & year < 2010) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = "line",
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
 ```
 
-[No. 5 Mine (101322)](http://www.chicagotribune.com/news/chi-0209220285sep22-story.html)
+#### [Sago Mine (4608791, clearly wrong)](https://en.wikipedia.org/wiki/Sago_Mine_disaster)
 
-```{r echo=FALSE}
+```{r sago, echo=FALSE}
 library(ggplot2)
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-no.5.mine <- complete.active.quarters %>% filter(mine_id == 101322 & (year < 2001 | (year == 2001 & quarter < 3))) %>%
-  select(year, quarter, num.days.lost, viol.quantity)
-qplot(x=year + quarter / 4, y=viol.quantity, data=no.5.mine, geom = "line",
-     main = "No. 5 Mine: Number of Quarterly Violations",
-     xlab = "Year", ylab = "Count")
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 4608791 & year < 2006) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = "line",
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
 ```
 
-[Darby Mine No. 1 (1518185)](http://www.chicagotribune.com/news/chi-0209220285sep22-story.html)
+#### [Darby Mine No. 1 (1518185)](http://www.chicagotribune.com/news/chi-0209220285sep22-story.html)
 
-```{r echo=FALSE}
+```{r darby, echo=FALSE}
 library(ggplot2)
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-darby.mine <- complete.active.quarters %>% filter(mine_id == 1518185 & (year < 2006 | (year == 2006 & quarter < 2))) %>%
-  select(year, quarter, num.days.lost, viol.quantity)
-qplot(x=year + quarter / 4, y=viol.quantity, data=darby.mine, geom = "line",
-     main = "Darby Mine No. 1: Number of Quarterly Violations",
-     xlab = "Year", ylab = "Count")
+# darby.mine <- complete.active.quarters %>% filter(mine_id == 1518185 & (year < 2006 | (year == 2006 & quarter < 2))) %>%
+#   select(year, quarter, num.days.lost, viol.quantity)
+# qplot(x=year + quarter / 4, y=viol.quantity, data=darby.mine, geom = "line",
+#      main = "Darby Mine No. 1: Number of Quarterly Violations",
+#      xlab = "Year", ylab = "Count")
+
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 1518185 & year < 2006) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = "line",
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
 ```
 
-[Gibson Mine	(1202215)](http://coaldiver.org/msha-id/1202215)
+#### [Gibson Mine	(1202215)](http://coaldiver.org/msha-id/1202215)
 
-```{r echo=FALSE}
+```{r gibson, echo=FALSE}
 library(ggplot2)
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-gibson.mine <- complete.active.quarters %>% filter(mine_id == 1202215 & (year < 2007 | (year == 2007 & quarter < 3))) %>%
-  select(year, quarter, num.days.lost, viol.quantity)
-qplot(x=year + quarter / 4, y=viol.quantity, data=gibson.mine, geom = "line",
-     main = "Gibson Mine: Number of Quarterly Violations",
-     xlab = "Year", ylab = "Count")
+# gibson.mine <- complete.active.quarters %>% filter(mine_id == 1202215 & (year < 2007 | (year == 2007 & quarter < 3))) %>%
+#   select(year, quarter, num.days.lost, viol.quantity)
+# qplot(x=year + quarter / 4, y=viol.quantity, data=gibson.mine, geom = "line",
+#      main = "Gibson Mine: Number of Quarterly Violations",
+#      xlab = "Year", ylab = "Count")
+
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 1202215 & year < 2007) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = "line",
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
+```
+
+#### Dotiki Mine
+
+```{r Dotiki Mine, echo=FALSE}
+library(ggplot2)
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 1502132 & year < 2010) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = "line",
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
 ```
@@
 Among the top 10 mine-quarters based on number of days lost,
-```{r echo=FALSE}
+```{r top 10 days lost, echo=FALSE}
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
@@
 Let's take a look at its second quarter of 2005:
-```{r echo=FALSE}
+```{r new era, echo=FALSE}
 if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
@@
 In particular, the top accident was:
-```{r echo=FALSE}
+```{r new era accident 1, echo=FALSE}
 if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
@@
 and the second to top accident was:
-```{r echo=FALSE}
+```{r new era, accident 2, echo=FALSE}
 if(!exists("new.era.top.accidents")){
@@
 The following linear model tests a few predictors and their ability to predict number of days lost or restircted:
-```{r echo=FALSE}
+```{r simple lm of days lost/restricted, echo=FALSE}
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
@@
 The following linear model tests a few predictors and their ability to predict number of deaths:
-```{r echo=FALSE}
+```{r simple lm of death, echo=FALSE}
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}

### 2017-03-13 A New Hope/fixed_effects.R

@@ -1,0 +1,63 @@
+setwd("~/jy/Regulatory Analysis Research - Prof. Venkat Venkatasubramanian/")
+
+library(dplyr)
+library(survival)
+library(RcppRoll)
+
+if(!exists("accidents")) {
+  accidents <- read.csv("msha_accident_counts2.csv", stringsAsFactors = F)
+  accidents <- accidents %>% filter(year >= 2000) %>% filter(year <= 2015)
+}
+if(!exists("violations")) {
+  violations <- read.table("msha_violation_counts2.csv", header = T, sep=",", quote='"', stringsAsFactors=FALSE, fill=TRUE)  
+  violations <- violations %>% filter(year >= 2000) %>% filter(year <= 2015)
+  violations$total_penalties <- ifelse(is.na(as.numeric(violations$total_penalties)), 0, as.numeric(violations$total_penalties))
+}
+
+all.mine.ids <- union(unique(accidents$mine_id), unique(violations$mine_id))
+
+yearly.accidents <- accidents %>% group_by(mine_id, year) %>% summarize(total.accidents = sum(perm_disability + fatality))
+yearly.violations <- violations %>% group_by(mine_id, year) %>% summarize(total.violations = sum(num_violations), yearly.total.penalties = sum(total_penalties))
+
+
+window.size <- 3
+start.year <- 2000 + window.size
+end.year <- 2015
+
+all.zero.yearly.accidents <- data.frame(mine_id = rep(unique(all.mine.ids), each = length(start.year:end.year)),
+                                        year = rep(start.year:end.year, times = length(unique(all.mine.ids))),
+                                        total.accidents = rep(0, times = length(unique(all.mine.ids)) * length(start.year:end.year)))
+all.zero.yearly.violations <- data.frame(mine_id = rep(unique(all.mine.ids), each = length(start.year:end.year)),
+                                         year = rep(start.year:end.year, times = length(unique(all.mine.ids))),
+                                         total.violations = rep(0, times = length(unique(all.mine.ids)) * length(start.year:end.year)),
+                                         yearly.total.penalties = rep(0, times = length(unique(all.mine.ids)) * length(start.year:end.year)))
+
+full.yearly.accidents <- all.zero.yearly.accidents %>% anti_join(yearly.accidents, by = c("mine_id", "year")) %>% 
+                                                       bind_rows(yearly.accidents) %>% arrange(mine_id, year) %>%
[…]

### 2017-03-13 San Antonio/fatal_or_explosive.R

@@ -1,0 +1,145 @@
+rm(list=ls())
+# Upper Big Branch: 4608436
+# This is an updated data generation to separate_fatality.R
+# Only days lost caused by "IGNITION OR EXPLOSION OF GAS OR DUST" are recorded
+# Inputs: raw accidents data and final data with only active mine-quarters
+# Output: days lost, days restrict, deaths, and violations
+
+# setup
+require(dplyr)
+library(RcppRoll)
+setwd("~/Git/violation-data-analysis")
+
+# load accidents, violations, and function roll_over
+if(!exists("accidents") | 
+   !exists("full.days_lost.accidents.date") | 
+   !exists("roll_over") | 
+   !exists("mines")) {
+  load("./San Antonio/data/secret_ingredients.RData")
+  print("raw RData loaded")
+}
+
+# constant parameters
+longest.period <- 3
+actual.start.year <- 2000
+end.year <- 2015
+
+# generate accident_roll_over and death_roll_over
+actual.accidents <- accidents %>% filter(
+  is.element(inj_degr_desc, c("DAYS AWAY FROM WORK ONLY",
+                              "DYS AWY FRM WRK & RESTRCTD ACT",
+                              "DAYS RESTRICTED ACTIVITY ONLY",
+                              "FATALITY",
+                              "PERM TOT OR PERM PRTL DISABLTY",
+                              "NO VALUE FOUND"
+                              )
+             ) 
+  )
+
+# initialize new columns: "deaths" and "perm_dis"
+actual.accidents["deaths"] <- 0
+actual.accidents["perm_dis"] <- 0
+actual.accidents[which(actual.accidents$inj_degr_desc == "FATALITY"), "deaths"] <- 1
+actual.accidents[which(actual.accidents$inj_degr_desc == "PERM TOT OR PERM PRTL DISABLTY"), "perm_dis"] <- 1
+
+
+# retain days_lost only to rows with "IGNITION OR EXPLOSION OF GAS OR DUST"
+actual.accidents[is.na(actual.accidents$days_lost), "days_lost"] <- 0
+actual.accidents[is.na(actual.accidents$days_restrict), "days_restrict"] <- 0
+actual.accidents[which(actual.accidents$ai_class_desc !=  "IGNITION OR EXPLOSION OF GAS OR DUST"), "days_lost"] <- 0
+actual.accidents[which(actual.accidents$ai_class_desc !=  "IGNITION OR EXPLOSION OF GAS OR DUST"), "days_restrict"] <- 0
+
[…]

### 2017-03-13 San Antonio/prepare_violations.R

@@ -1,0 +1,21 @@
+rm(list=ls())
+require(dplyr)
+setwd("~/Git/violation-data-analysis")
+violations <- read.csv("./San Antonio/data/AssessedViolations.csv", stringsAsFactors = FALSE)
+colnames(violations)[1] <- "mine_id"
+
+violations <- violations %>% mutate(
+  cal_yr = as.numeric(format(as.Date(OCCURRENCE_DT, "%m/%d/%Y"), "%Y")),
+  cal_mm = as.numeric(format(as.Date(OCCURRENCE_DT, "%m/%d/%Y"), "%m")),
+  cal_qtr = ifelse(
+    cal_mm >9, 4, ifelse(
+      cal_mm > 6, 3, ifelse(
+        cal_mm > 3, 2, 1
+      )
+    )
+  )
+)
+
+violations$violation <- 1
+violations.lite <- violations %>% select(mine_id, cal_qtr, cal_yr, violation)
+save(violations.lite, file="./San Antonio/data/albert_violation.RData")

### 2017-03-13 San Antonio/separate_fatality.R

@@ -11,10 +11,11 @@ setwd("~/Git/violation-data-analysis")
 
 # load accidents, violations, and function roll_over
 if(!exists("accidents") | 
-   !exists("full.days_lost.accidents.date") | 
+   !exists("violations.lite") | 
    !exists("roll_over") | 
    !exists("mines")) {
   load("./San Antonio/data/secret_ingredients.RData")
+  load("./San Antonio/data/albert_violation.RData")
   print("raw RData loaded")
 }
 
@@ -32,13 +33,11 @@ actual.accidents <- accidents %>% filter(is.element(inj_degr_desc, c("DAYS AWAY 
                                                                      "NO VALUE FOUND")))
 actual.accidents["deaths"] <- 0
 actual.accidents["perm_dis"] <- 0
-# na.and.no.value.index <- is.element(actual.accidents$inj_degr_desc, "NO VALUE FOUND") & is.na(actual.accidents$days_lost)
 actual.accidents[is.na(actual.accidents$days_lost), "days_lost"] <- 0
 actual.accidents[is.na(actual.accidents$days_restrict), "days_restrict"] <- 0
 actual.accidents[which(actual.accidents$inj_degr_desc == "FATALITY"), "deaths"] <- 1
 actual.accidents[which(actual.accidents$inj_degr_desc == "PERM TOT OR PERM PRTL DISABLTY"), "perm_dis"] <- 1
 actual.accidents$ai_dt_actual_date <- strftime(actual.accidents$ai_dt, "%F")
-# rm(na.and.no.value.index, accidents)
 
 quarter.level.num.days.lost <- actual.accidents %>% group_by(mine_id, cal_qtr, cal_yr)%>% 
   summarize(base = sum(days_lost, na.rm = T))
@@ -92,15 +91,28 @@ colnames(dis_rollover) <- c(
 )
 rm(quarter.level.num.perm_dis)
 
-# cleanup violations
-active_violation_rollover = full.days_lost.accidents.date %>%
-  select(mine_id, year, quarter, viol.quantity, last.quarter.y, last.year.y, last.three.years.y)
-active_violation_rollover$active <- TRUE
-colnames(active_violation_rollover) <- c(
-  "mine_id", "year", "quarter", 
-  "viol.quantity", "last.quarter.viol", "last.year.viol", "last.three.years.viol",
-  "active"
[…]

### 2017-03-13 San Antonio/simple_lm.R

@@ -1,3 +1,4 @@
+rm(list=ls())
 # some resources:
 # http://blog.minitab.com/blog/adventures-in-statistics-2/how-to-interpret-regression-analysis-results-p-values-and-coefficients
 # http://stats.stackexchange.com/questions/59250/how-to-interpret-the-output-of-the-summary-method-for-an-lm-object-in-r
@@ -7,29 +8,53 @@ require(dplyr)
 require(survival)
 require(ggplot2)
 setwd("~/Git/violation-data-analysis")
-load("./San Antonio/output/result.RData")
+load("./San Antonio/output/result_select_accidents.RData")
+train <- complete.active.quarters %>% filter(active)
 
-train <- complete.active.quarters
+train.non.zero <- train %>% filter(num.days.lost > 0)
+plot(ecdf(train.non.zero$num.days.lost))
+
+qqnorm((train %>% filter(num.days.lost > 0))$num.days.lost)
+
+top.days.lost <- complete.active.quarters %>% filter(active) %>% arrange(desc(num.days.lost)) %>% select(mine_id, mine.name, year, quarter, num.days.lost)
+head(top.days.lost, 10)
+
+# Lightning strike at Shoal Creek mine injures 6
+# http://blog.al.com/spotnews/2007/06/lightning_strike_at_mine_injur.html
+View(accidents %>% filter(mine_id == 102901 & ai_dt == "2007-06-27") %>% arrange(desc(days_lost)))
+
+
 # train <- train %>% rowwise() %>% mutate(high_severity = num.death > 0 | num.days.lost > 100)
 # train <- train[, c(1:5, 22, 6:21)]
 # train %>% count(high_severity)
 
-View(train %>% filter(mine_id == 1102752))
-new_era <- accidents %>% filter(mine_id == 1102752 & cal_qtr == 2 & cal_yr == 2005) %>% arrange(
-  desc(days_lost)
-)
-View(new_era)
+# View(train %>% filter(mine_id == 1102752))
+# new_era <- accidents %>% filter(mine_id == 1102752 & cal_qtr == 2 & cal_yr == 2005) %>% arrange(
+#   desc(days_lost)
+# )
+# View(new_era)
+# 
+# summary(lm(num.days.lost + num.days.restrict ~ last.quarter.viol+last.year.viol+last.three.years.viol, data=train %>% filter(active)))
 
[…]

### 2017-03-13 San Antonio/presentation_materials.Rmd
updated markdown

@@
 author: "Yu"
-date: "3/12/2017"
+date: "3/13/2017"
 output: 
@@
 
-#### [Sago Mine (4608791, clearly wrong)](https://en.wikipedia.org/wiki/Sago_Mine_disaster)
+#### [No. 5 Mine (101322)](http://www.chicagotribune.com/news/chi-0209220285sep22-story.html)
+
+Usable violation data only date back to 2000.
+
+#### [Sago Mine (4608791)](https://en.wikipedia.org/wiki/Sago_Mine_disaster)
 
@@
   filter(mine_id == 4608791 & year < 2006) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = "line",
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
+```
+
+#### [Crandall Canyon Mine (4201715)](https://www.google.com/url?sa=t&rct=j&q=&esrc=s&source=web&cd=2&cad=rja&uact=8&ved=0ahUKEwjOyeGnvtTSAhVN42MKHdtyD7UQFggnMAE&url=http%3A%2F%2Fwww.deseretnews.com%2Farticle%2F695206530%2FVictims-of-Crandall-Canyon-Mine-disaster.html&usg=AFQjCNHqcZNs-qEa9SipaDkfKfYy5okkGQ&sig2=WMuhLw2C0D1jAxkiBM1uiw)
+
+```{r crandall, echo=FALSE}
+library(ggplot2)
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 4201715 & year < 2007) %>%
   group_by(year) %>%

### 2017-03-14 San Antonio/presentation_materials.Rmd
only rank top 10 fatalities after 2005

--- 2017-03-13 San Antonio/presentation_materials.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, message=FALSE)
 require(dplyr)
@@
 
-## Upper Big Branch and Other Deadliest Mine
+## Fatal Accidents
 
-Below is a list of the top 10 mine-quarters based on number of deaths:
+### Top 10 Quarterly Fatalities Since 2005
 
-```{r echo=FALSE}
+```{r top 10 fatality}
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-top.deaths <- complete.active.quarters %>% filter(active) %>% arrange(desc(num.death)) %>% select(mine.name, mine_id, year, quarter, num.death)
+top.deaths <- complete.active.quarters %>% filter(active & year >= 2005) %>% arrange(desc(num.death)) %>% select(mine.name, mine_id, year, quarter, num.death)
 head(top.deaths, 10)
 ```
-[The Upper Big Branch Mine (4608436)](https://en.wikipedia.org/wiki/Upper_Big_Branch_Mine_disaster)
+### Top 10 Causes of Death
+```{r top 10 causes of death}
+if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
+fatal.accidents <- accidents %>% 
+  filter(inj_degr_desc == "FATALITY") %>%
+  group_by(ai_class_desc) %>%
+  summarise(count = n()
+            ) %>%
+  arrange(desc(count))
+head(fatal.accidents, 10)
+```
 
-```{r echo=FALSE}
+### Upper Big Branch and Other Deadliest Mine
+
+#### [Upper Big Branch Mine (4608436)](https://en.wikipedia.org/wiki/Upper_Big_Branch_Mine_disaster)
+
+```{r upper big branch}
+# {r upper big branch, echo=FALSE, fig.show = 'hold', out.width = '49%'}
 library(ggplot2)
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-upper.big.branch <- complete.active.quarters %>% filter(mine_id == 4608436 & (year < 2010 | (year == 2010 & quarter < 2))) %>%
-  select(year, quarter, num.days.lost, viol.quantity)
-# ggplot() +
-#   geom_line(data = upper.big.branch, aes(x = year + quarter / 4, y = viol.quantity), color = "red") +
-#   geom_line(data = upper.big.branch, aes(x = year + quarter / 4, y = num.days.lost), color = "green")
-qplot(x=year + quarter / 4, y=viol.quantity, data=upper.big.branch, geom = "line",
-     main = "Upper Big Branch: Number of Quarterly Violations",
-     xlab = "Year", ylab = "Count")
+# upper.big.branch <- complete.active.quarters %>% filter(mine_id == 4608436 & (year < 2010 | (year == 2010 & quarter < 2))) %>%
+#   select(year, quarter, num.days.lost, viol.quantity)
+# # ggplot() +
+# #   geom_line(data = upper.big.branch, aes(x = year + quarter / 4, y = viol.quantity), color = "red") +
+# #   geom_line(data = upper.big.branch, aes(x = year + quarter / 4, y = num.days.lost), color = "green")
+# qplot(x=year + quarter / 4, y=viol.quantity, data=upper.big.branch, geom = "line",
+#      main = "Upper Big Branch: Number of Quarterly Violations",
+#      xlab = "Year", ylab = "Count")
+
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 4608436 & year < 2010) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
 ```
 
-[No. 5 Mine (101322)](http://www.chicagotribune.com/news/chi-0209220285sep22-story.html)
+<!--
+#### [No. 5 Mine (101322)](http://www.chicagotribune.com/news/chi-0209220285sep22-story.html)
 
-```{r echo=FALSE}
+Usable violation data only date back to 2000.
+-->
+
+#### [Sago Mine (4608791)](https://en.wikipedia.org/wiki/Sago_Mine_disaster)
+
+```{r sago}
 library(ggplot2)
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-no.5.mine <- complete.active.quarters %>% filter(mine_id == 101322 & (year < 2001 | (year == 2001 & quarter < 3))) %>%
-  select(year, quarter, num.days.lost, viol.quantity)
-qplot(x=year + quarter / 4, y=viol.quantity, data=no.5.mine, geom = "line",
-     main = "No. 5 Mine: Number of Quarterly Violations",
-     xlab = "Year", ylab = "Count")
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 4608791 & year < 2006) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
 ```
 
-[Darby Mine No. 1 (1518185)](http://www.chicagotribune.com/news/chi-0209220285sep22-story.html)
+#### [Crandall Canyon Mine (4201715)](https://www.google.com/url?sa=t&rct=j&q=&esrc=s&source=web&cd=2&cad=rja&uact=8&ved=0ahUKEwjOyeGnvtTSAhVN42MKHdtyD7UQFggnMAE&url=http%3A%2F%2Fwww.deseretnews.com%2Farticle%2F695206530%2FVictims-of-Crandall-Canyon-Mine-disaster.html&usg=AFQjCNHqcZNs-qEa9SipaDkfKfYy5okkGQ&sig2=WMuhLw2C0D1jAxkiBM1uiw)
 
-```{r echo=FALSE}
+```{r crandall}
 library(ggplot2)
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-darby.mine <- complete.active.quarters %>% filter(mine_id == 1518185 & (year < 2006 | (year == 2006 & quarter < 2))) %>%
-  select(year, quarter, num.days.lost, viol.quantity)
-qplot(x=year + quarter / 4, y=viol.quantity, data=darby.mine, geom = "line",
-     main = "Darby Mine No. 1: Number of Quarterly Violations",
-     xlab = "Year", ylab = "Count")
+
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 4201715 & year < 2007) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
 ```
 
-[Gibson Mine	(1202215)](http://coaldiver.org/msha-id/1202215)
+#### [Darby Mine No. 1 (1518185)](http://www.chicagotribune.com/news/chi-0209220285sep22-story.html)
 
-```{r echo=FALSE}
+```{r darby}
 library(ggplot2)
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-gibson.mine <- complete.active.quarters %>% filter(mine_id == 1202215 & (year < 2007 | (year == 2007 & quarter < 3))) %>%
-  select(year, quarter, num.days.lost, viol.quantity)
-qplot(x=year + quarter / 4, y=viol.quantity, data=gibson.mine, geom = "line",
-     main = "Gibson Mine: Number of Quarterly Violations",
-     xlab = "Year", ylab = "Count")
+# darby.mine <- complete.active.quarters %>% filter(mine_id == 1518185 & (year < 2006 | (year == 2006 & quarter < 2))) %>%
+#   select(year, quarter, num.days.lost, viol.quantity)
+# qplot(x=year + quarter / 4, y=viol.quantity, data=darby.mine, geom = "line",
+#      main = "Darby Mine No. 1: Number of Quarterly Violations",
+#      xlab = "Year", ylab = "Count")
+
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 1518185 & year < 2006) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
+```
+
+#### [Gibson Mine	(1202215)](http://coaldiver.org/msha-id/1202215)
+
+```{r gibson}
+library(ggplot2)
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+# gibson.mine <- complete.active.quarters %>% filter(mine_id == 1202215 & (year < 2007 | (year == 2007 & quarter < 3))) %>%
+#   select(year, quarter, num.days.lost, viol.quantity)
+# qplot(x=year + quarter / 4, y=viol.quantity, data=gibson.mine, geom = "line",
+#      main = "Gibson Mine: Number of Quarterly Violations",
+#      xlab = "Year", ylab = "Count")
+
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 1202215 & year < 2007) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
+```
+
+#### [Revenue Mine (503528)](https://www.dol.gov/opa/media/press/msha/MSHA20141984fs.pdf)
+
+```{r Revenue Mine}
+library(ggplot2)
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 503528 & year < 2013) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
+```
+
+#### [Dotiki Mine (1502132)](http://www.huffingtonpost.com/2010/04/29/kentucky-mine-accident-2-_n_556851.html)
+
+```{r Dotiki Mine}
+library(ggplot2)
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 1502132 & year < 2010) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
+```
+
+#### [P-3 Hignite (1510212)](https://arlweb.msha.gov/FATALS/2005/FTL05c1415.asp)
+
+```{r p-3}
+library(ggplot2)
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 1510212 & year < 2005) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
+```
+
+#### [D-14 Stillhouse (1517165)](https://arlweb.msha.gov/FATALS/2005/FTL05c1112.asp)
+
+```{r D-14}
+library(ggplot2)
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 1517165 & year < 2005) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
+```
+
+#### [Equality (1519344)](https://arlweb.msha.gov/FATALS/2011/FTL11c1819.asp)
+
+```{r equality}
+library(ggplot2)
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 1519344 & year < 2011) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "count")
 ```
@@
 Among the top 10 mine-quarters based on number of days lost,
-```{r echo=FALSE}
+```{r top 10 days lost}
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
@@
 Let's take a look at its second quarter of 2005:
-```{r echo=FALSE}
+```{r new era}
 if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
@@
 In particular, the top accident was:
-```{r echo=FALSE}
+```{r new era accident 1}
 if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
@@
 and the second to top accident was:
-```{r echo=FALSE}
+```{r new era, accident 2}
 if(!exists("new.era.top.accidents")){
@@
 
-## Simple Linear Model
+## Statistical Models
+
+### Simple Linear Model
 
 The following linear model tests a few predictors and their ability to predict number of days lost or restircted:
-```{r echo=FALSE}
+```{r simple lm of days lost/restricted}
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
@@
 The following linear model tests a few predictors and their ability to predict number of deaths:
-```{r echo=FALSE}
+```{r simple lm of death}
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
@@
 ```
+
+### Lognormal Distribution
+
+```{r lognormal, echo=FALSE}
+# require(fitdistrplus)
+# if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+# if(!exists("f")){
+#   training.nonzero <- complete.active.quarters %>% filter(num.days.lost >0 & active)
+#   f <- fitdist(log(training.nonzero$num.days.lost), "norm")
+# }
+# par("mar")
+# par(mar=c(1,1,1,1))
+# plotdist(log(training.nonzero$num.days.lost),"norm",para=list(mean=f$estimate[1], sd=f$estimate[2]))
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+non.zero <- complete.active.quarters %>% filter(num.days.lost >0 & active)
+y <- log(non.zero$num.days.lost)
+mu <- mean(y)
+sig <- sd(y)
+qqplot(qnorm(ppoints(500), mean = mu, sd = sig), y,
+       xlab = "log(num.days.lost)", ylab = "quantile", main = "Normal Q-Q Plot")
+qqline(y, distribution = function(p) qnorm(p, mean = mu, sd = sig))
+```

### 2017-03-14 San Antonio/simple_lm.R

@@ -8,8 +8,39 @@ require(dplyr)
 require(survival)
 require(ggplot2)
 setwd("~/Git/violation-data-analysis")
-load("./San Antonio/output/result_select_accidents.RData")
+load("./San Antonio/output/result.RData")
 train <- complete.active.quarters %>% filter(active)
+train <- train %>% mutate(severe = num.death + num.dis > 0)
+
+hist(train$num.days.restrict)
+
+# good mine/bad mine
+# http://www.markhneedham.com/blog/2014/11/26/r-dplyr-select-random-rows-from-a-data-frame/
+set.seed(1)
+train.positive <- train %>% filter(severe)
+train.negative <- train %>% filter(!severe) %>% sample_n(nrow(train.positive))
+train.training <- rbind(train.positive, train.negative)
+train.training <- train.training %>%
+  mutate(magic.num = ifelse(last.year.viol == 0,
+                        last.year.viol / last.three.years.viol * 3,
+                        1)
+         )
+ggplot(train.training, aes(magic.num, fill = severe)) + geom_density(alpha = 0.2)
+
+#
+y <- rchisq(500, df = 3)
+## Q-Q plot for Chi^2 data against true theoretical distribution:
+# qqplot(qchisq(ppoints(500), df = 3), y)
+qqline(y, distribution = function(p) qnorm(p, mean = 3.3, sd = 1.78),
+       prob = c(0.1, 0.6), col = 2)
+
+
+# lognormal 
+training.nonzero <- complete.active.quarters %>% filter(num.days.lost >0 & active)
+qqnorm(log(training.nonzero$num.days.lost))
+qqline(log(training.nonzero$num.days.lost), distribution = function(p) qnorm(p, mean = 3.3, sd = 1.78))
+      # distribution = function(p) qnorm(p, mean = 3.3, sd = 1.78))
+
 
 train.non.zero <- train %>% filter(num.days.lost > 0)
 plot(ecdf(train.non.zero$num.days.lost))

### 2017-03-14 San Antonio/presentation_materials.Rmd
added statistics of ai_class to top ten fatal accidents

@@
 require(dplyr)
+require(ggplot2)
 knitr::opts_knit$set(root.dir = "~/Git/violation-data-analysis")
@@
 ```
-### Top 10 Causes of Death
+### Top 10 Causes of Death Since 2005
 ```{r top 10 causes of death}
@@
 fatal.accidents <- accidents %>% 
-  filter(inj_degr_desc == "FATALITY") %>%
+  filter(inj_degr_desc == "FATALITY" & cal_yr >= 2005) %>%
   group_by(ai_class_desc) %>%
@@
 # {r upper big branch, echo=FALSE, fig.show = 'hold', out.width = '49%'}
-library(ggplot2)
+if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
+fatal.accidents <- accidents %>% 
+  filter(inj_degr_desc == "FATALITY" & 
+           cal_yr == 2010 & cal_qtr == 2 & mine_id ==4608436) %>%
+  group_by(ai_class_desc) %>%
+  summarise(count = n()
+            ) %>%
+  arrange(desc(count))
+head(fatal.accidents, 10)
+
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
@@
 ```{r sago}
-library(ggplot2)
+if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
+fatal.accidents <- accidents %>% 
+  filter(inj_degr_desc == "FATALITY" & 
+           cal_yr == 2006 & cal_qtr == 1 & mine_id ==4608791) %>%
+  group_by(ai_class_desc) %>%
+  summarise(count = n()
+            ) %>%
+  arrange(desc(count))
+head(fatal.accidents, 10)
+
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
@@
 ```{r crandall}
-library(ggplot2)
+if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
+fatal.accidents <- accidents %>% 
+  filter(inj_degr_desc == "FATALITY" & 
+           cal_yr == 2007 & cal_qtr == 3 & mine_id ==4201715) %>%
+  group_by(ai_class_desc) %>%
+  summarise(count = n()
+            ) %>%
+  arrange(desc(count))
+head(fatal.accidents, 10)
+
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-
 yearly.viol <- complete.active.quarters %>% 
@@
 ```{r darby}
-library(ggplot2)
+if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
+fatal.accidents <- accidents %>% 
+  filter(inj_degr_desc == "FATALITY" & 
+           cal_yr == 2006 & cal_qtr == 2 & mine_id ==1518185) %>%
+  group_by(ai_class_desc) %>%
+  summarise(count = n()
+            ) %>%
+  arrange(desc(count))
+head(fatal.accidents, 10)
+
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
@@
 ```{r gibson}
-library(ggplot2)
+if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
+fatal.accidents <- accidents %>% 
+  filter(inj_degr_desc == "FATALITY" & 
+           cal_yr == 2007 & cal_qtr == 3 & mine_id ==1202215) %>%
+  group_by(ai_class_desc) %>%
+  summarise(count = n()
+            ) %>%
+  arrange(desc(count))
+head(fatal.accidents, 10)
+
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
@@
 ```{r Revenue Mine}
-library(ggplot2)
+if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
+fatal.accidents <- accidents %>% 
+  filter(inj_degr_desc == "FATALITY" & 
+           cal_yr == 2013 & cal_qtr == 4 & mine_id ==503528) %>%
+  group_by(ai_class_desc) %>%
+  summarise(count = n()
+            ) %>%
+  arrange(desc(count))
+head(fatal.accidents, 10)
+
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
@@
 ```{r Dotiki Mine}
-library(ggplot2)
+if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
+fatal.accidents <- accidents %>% 
+  filter(inj_degr_desc == "FATALITY" & 
+           cal_yr == 2010 & cal_qtr == 2 & mine_id ==1502132) %>%
+  group_by(ai_class_desc) %>%
+  summarise(count = n()
+            ) %>%
+  arrange(desc(count))
+head(fatal.accidents, 10)
+
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
@@
 ```{r p-3}
-library(ggplot2)
+if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
+fatal.accidents <- accidents %>% 
+  filter(inj_degr_desc == "FATALITY" & 
+           cal_yr == 2005 & cal_qtr == 3 & mine_id ==1510212) %>%
+  group_by(ai_class_desc) %>%
+  summarise(count = n()
+            ) %>%
+  arrange(desc(count))
+head(fatal.accidents, 10)
+
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
@@
 ```{r D-14}
-library(ggplot2)
+if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
+fatal.accidents <- accidents %>% 
+  filter(inj_degr_desc == "FATALITY" & 
+           cal_yr == 2005 & cal_qtr == 3 & mine_id ==1517165) %>%
+  group_by(ai_class_desc) %>%
+  summarise(count = n()
+            ) %>%
+  arrange(desc(count))
+head(fatal.accidents, 10)
+
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
@@
 ```{r equality}
-library(ggplot2)
+if(!exists("accidents")){load("./San Antonio/data/secret_ingredients.RData")}
+fatal.accidents <- accidents %>% 
+  filter(inj_degr_desc == "FATALITY" & 
+           cal_yr == 2011 & cal_qtr == 4 & mine_id ==1519344) %>%
+  group_by(ai_class_desc) %>%
+  summarise(count = n()
+            ) %>%
+  arrange(desc(count))
+head(fatal.accidents, 10)
+
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
@@
     ai_occ_desc, 
-    ai_narr
+    ai_narr,
+    inj_degr_desc
     ) 
 print(as.character(new.era.top.accidents$ai_narr[1]))
+print(as.character(new.era.top.accidents$inj_degr_desc[1]))
 ```
@@
 print(as.character(new.era.top.accidents$ai_narr[2]))
+print(as.character(new.era.top.accidents$inj_degr_desc[2]))
 ```
@@
 
-### Lognormal Distribution
+### Lognormal Distribution of Nonzero Days Lost
 

### 2017-03-17 San Antonio/markov_chains.Rmd
All markov chain complete except for weighted average

---
title: "markov_chains"
output: html_document
---

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(dplyr)
require(plyr)
require(RcppRoll)
require(markovchain)
require(reshape)
knitr::opts_knit$set(root.dir = "~/Documents/Violation_Project")
setwd("~/Documents/Violation_Project")

if(!exists("accidents") | 
   !exists("full.days_lost.accidents.date") | 
   !exists("roll_over") | 
   !exists("mines")) {
  load("./San Antonio/data/secret_ingredients.RData")
  load("./San Antonio/output/result_select_accidents.RData")
  print("raw RData loaded")
}
```

```{r}

temp <- complete.active.quarters %>% filter(active==TRUE)
set.seed(20)
clustering <- kmeans(temp[, 6:ncol(temp)], 3, nstart=20)

freq_table <- table(clustering$cluster)

```

```{r}
clustering
freq_table

clustering$centers
```

Transformed data to become in a melted format containing quarter-year, mine_id allowing for markov chain analysis
```{r}

temp$cluster <- clustering$cluster
temp$date <- paste(as.character(temp$year), "-" , as.character(temp$quarter))
temp2 <- temp %>% select(mine_id, cluster, date)
melted <- melt(temp2, id=c("mine_id", "date"))
aggregate_cluster_seq <- cast(melted, mine_id~date)
```

Markov Chain stuff
```{r}
markov <- function(x){
  x <- as.integer(x)
  cluster_seq_raw <- x[!is.na(x)]
  if(length(cluster_seq_raw) > 1){
    cluster_sequence <- cluster_seq_raw[2:length(cluster_seq_raw)]
    cur_markov_chain <- markovchainFit(data=cluster_sequence)
    return (list("mine_id" = x[1], "markov_chain" = cur_markov_chain$estimate, "num_rows" = length(cluster_seq_raw)))
  }
  return (list("mine_id" = x[1], "markov_chain" = NA, "num_rows" = 0))
}
```

```{r}
result<-alply(aggregate_cluster_seq, 1, markov)
```

```{r}
num_rows_ind <- sapply(result, "[[", 3)
weights <- num_rows_ind/sum(num_rows_ind)
markov_chains_ind <- sapply(result, "[[", 2)
aggregated_markov_chain <- weighted.mean(markov_chains_ind, weights)
```

### 2017-03-18 San Antonio/presentation_materials.Rmd
Presentation stuff

@@
 author: "Yu"
-date: "3/13/2017"
+date: "3/17/2017"
 output: 
@@
 knitr::opts_knit$set(root.dir = "~/Git/violation-data-analysis")
+```
+
+## Accident Trends
+
+### Fatality since 1983
+
+```{r fatality trend}
+# if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+# yearly.death <- complete.active.quarters %>% 
+#   filter(year >= 2000 & active) %>%
+#   group_by(year) %>%
+#   summarise(count = sum(num.death),
+#             avg = mean(num.death))
+# qplot(x = year, y = count, data = yearly.death, geom = c("point", "line"),
+#       main = "Number of Total Fatalities Due to Accidents",
+#       xlab = "Year", ylab = "Count")
+# # qplot(x = year, y = avg, data = yearly.death, geom = c("point", "line"),
+# #       main = "Number of Avgerage Fatalities Due to Accidents",
+# #       xlab = "Year", ylab = "Count")
+if(!exists("accidents")) {load("./San Antonio/data/secret_ingredients.RData")}
+yearly.death <- accidents %>% 
+  mutate(death = ifelse(inj_degr_desc == "FATALITY", 1, 0)) %>%
+  group_by(cal_yr) %>%
+  summarise(count = sum(death))
+qplot(x = cal_yr, y = count, data = yearly.death, geom = c("point", "line"),
+      main = "Number of Total Fatalities Due to Accidents",
+      xlab = "Year", ylab = "Count")
+```
+
+### Number of Days Lost since 2000
+
+```{r num.days.lost trend}
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+yearly.lost <- complete.active.quarters %>% 
+  filter(year >= 2000 & active) %>%
+  group_by(year) %>%
+  summarise(count = sum(num.days.lost),
+            avg = mean(num.days.lost))
+qplot(x = year, y = count, data = yearly.lost, geom = c("point", "line"),
+      main = "Number of Total Days Lost Due to Accidents",
+      xlab = "Year", ylab = "Count")
+# qplot(x = year, y = avg, data = yearly.lost, geom = c("point", "line"),
+#       main = "Number of Average Days Lost Due to Accidents",
+#       xlab = "Year", ylab = "Count")
+```
+
+### Number of Violations since 2000
+
+```{r viol.quantity trend}
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+yearly.viol <- complete.active.quarters %>% 
+  filter(year >= 2000 & active) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity),
+            avg = mean(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Total Violations",
+      xlab = "Year", ylab = "Count")
+# qplot(x = year, y = avg, data = yearly.lost, geom = c("point", "line"),
+#       main = "Number of Average Days Lost Due to Accidents",
+#       xlab = "Year", ylab = "Count")
 ```
@@
       main = "Number of Yearly Violations",
-      xlab = "Year", ylab = "count")
+      xlab = "Year", ylab = "Count")
 ```

### 2017-03-18 San Antonio/slides/presentation_slides.Rmd
Presentation stuff

---
title: "A Data-Driven Early Warning System for Mining Accident"
shorttitle: "Mining Accident Data Analysis"
author: |
  | Yu Luo, Ashutosh Nanda, Shiva Rajgopal, Vinay Ramesh,
  | Venkat Venkatasubramanian, Zhizun Zhang, and Catherine Zhao
shortauthor: "Luo et al."
institute: "Columbia University"
date: "3/27/2017"
output: 
  beamer_presentation:
    # incremental: true
    template: rmd_template.tex
    fig_height: 4.75
    toc: true
    slide_level: 2
    includes:
      in_header: rmd_header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
require(dplyr)
require(ggplot2)
knitr::opts_knit$set(root.dir = "~/Git/violation-data-analysis")
```

# Introduction: Mine Safety and MSHA Data

<!-- https://www.msha.gov/about/mission -->
## Mine Safety and Health Administration (MSHA)

- Formed in	1977
- Agency of the United States Department of Labor
- Mission
    - Prevent death, illness, and injury from mining
    - Promote safe and healthful workplaces for U.S. miners
    - Develop and enforce safety and health rules
    - Provide technical, educational, and other types of assistance

## Fatality Trend Since 1983

```{r fatality trend}
if(!exists("accidents")) {load("./San Antonio/data/secret_ingredients.RData")}
yearly.death <- accidents %>% 
  mutate(death = ifelse(inj_degr_desc == "FATALITY", 1, 0)) %>%
  group_by(cal_yr) %>%
  summarise(count = sum(death))
qplot(x = cal_yr, y = count, data = yearly.death, geom = c("point", "line"),
      main = "Number of Total Fatalities Due to Accidents",
      xlab = "Year", ylab = "Count")
```

## Upper Big Branch Disaster

# Methods: Data and Model Assumptions

## Primary Datasets

- **Mine accidents**: "msha_accident.csv" (681,386 obs., retrieved 1/26/2017, from <https://enforcedata.dol.gov/views/data_summary.php>)
- **MSHA assessed violations**: "AssessedViolations.csv" (2,169,804 obs., retrieved 12/10/2016, from <https://arlweb.msha.gov/OpenGovernmentData/OGIMSHA.asp>)

## Master Data Table

```{r master data}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
names(complete.active.quarters)
```

## Model

# Results and Discussion

## Number of Days Lost since 2000

```{r num.days.lost trend}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.lost <- complete.active.quarters %>% 
  filter(year >= 2000 & active) %>%
  group_by(year) %>%
  summarise(count = sum(num.days.lost),
            avg = mean(num.days.lost))
qplot(x = year, y = count, data = yearly.lost, geom = c("point", "line"),
      main = "Number of Total Days Lost Due to Accidents",
      xlab = "Year", ylab = "Count")
# qplot(x = year, y = avg, data = yearly.lost, geom = c("point", "line"),
#       main = "Number of Average Days Lost Due to Accidents",
#       xlab = "Year", ylab = "Count")
```

## Number of Violations since 2000

```{r viol.quantity trend}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
yearly.viol <- complete.active.quarters %>% 
  filter(year >= 2000 & active) %>%
  group_by(year) %>%
  summarise(count = sum(viol.quantity),
            avg = mean(viol.quantity))
qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
      main = "Number of Total Violations",
      xlab = "Year", ylab = "Count")
```

# Conclusion

<!-- USEFUL LINKS:
http://www.sr.bham.ac.uk/~ajrs/R/r-show_data.html
http://rmarkdown.rstudio.com/articles_beamer.html
https://groups.google.com/forum/#!searchin/pandoc-discuss/beamer/pandoc-discuss/XMYG6ewG7a0/Q1LKHaOFBDUJ -->

### 2017-03-18 San Antonio/markov_chains.Rmd
completed transition matrix

@@
 ---
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
 
@@
 require(reshape)
+```
+```{r setting directories}
 knitr::opts_knit$set(root.dir = "~/Documents/Violation_Project")
@@
 freq_table
-
-clustering$centers
 ```
@@
 
-Markov Chain stuff
+Markov Chain function
 ```{r}
@@
     cur_markov_chain <- markovchainFit(data=cluster_sequence)
-    return (list("mine_id" = x[1], "markov_chain" = cur_markov_chain$estimate, "num_rows" = length(cluster_seq_raw)))
+    return (list("mine_id" = x[1], "markov_chain" = cur_markov_chain$estimate@transitionMatrix, "num_rows" = length(cluster_seq_raw)))
   }
@@
 weights <- num_rows_ind/sum(num_rows_ind)
+
 markov_chains_ind <- sapply(result, "[[", 2)
-aggregated_markov_chain <- weighted.mean(markov_chains_ind, weights)
+
+weighted_transition <- mapply(function(x,y) x * y, x = weights, y = markov_chains_ind)
+
+#adding matrix of different dimentions, source: http://stackoverflow.com/questions/13571359/join-and-sum-not-compatible-matrices
+add_matrices_1 <- function(...) {
+  a <- list(...)
+  cols <- sort(unique(unlist(lapply(a, colnames))))
+  rows <- sort(unique(unlist(lapply(a, rownames))))
+  out <- array(0, dim=c(length(rows), length(cols)), dimnames=list(rows,cols))
+  for(M in a) { out[rownames(M), colnames(M)] <- out[rownames(M), colnames(M)] + M }
+  return(out)
+}
+
+final_transition<- matrix(rep(0,times= 9), ncol = 3, nrow = 3)
+colnames(final_transition) <- c(1,2,3)
+rownames(final_transition) <- c(1,2,3)
+
+for(i in 1:length(weighted_transition)){
+  final_transition <- add_matrices_1(final_transition,weighted_transition[[i]])
+}
+final_transition
+
 ```

### 2017-03-18 San Antonio/fixed_effects_luo.R
fixed effects model kinda works now

@@ -1,0 +1,43 @@
+rm(list=ls())
+
+library(dplyr)
+library(survival)
+library(caret)
+library(e1071)
+setwd("~/Git/violation-data-analysis")
+load("./San Antonio/output/result.RData")
+
+sigmoid <- function(x) {
+  1 / (1 + exp(-x))
+}
+
+active.quarters <- complete.active.quarters %>% filter(active)
+active.quarters <- active.quarters %>% mutate(
+  severe = ifelse(
+    num.death + num.dis > 0, TRUE, FALSE
+  )
+)
+fe.model <- clogit(severe ~ 
+                     last.quarter.lost+last.year.lost+last.three.years.lost
+                   +last.quarter.restrict+last.year.restrict+last.three.years.restrict
+                   +last.quarter.viol+last.year.viol+last.three.years.viol
+                   +last.quarter.death+last.year.death+last.three.years.death
+                   + strata(mine_id), data = active.quarters)
+summary(fe.model)
+
+
+
+fe.predict <- predict(object = fe.model, type = "lp") %>% sigmoid()
+test <- active.quarters %>% select(mine_id, mine.name, year, quarter, num.death, num.dis, severe)
+test$prob <- fe.predict
+test <- test %>% mutate(
+  prediction = ifelse(
+    prob > 0.5, TRUE, FALSE
+  )
+)
+head(test %>% filter(year >= 2005) %>% arrange(desc(num.death)), 20)
+
+test.performance <- confusionMatrix(data = test$prediction, reference = test$severe, positive = "TRUE")
+print(test.performance)
+
+save(fe.model, active.quarters, test, test.performance, file="./San Antonio/output/results_fixed_effects.RData")

### 2017-03-19 San Antonio/presentation_materials.Rmd
more complete fixed effects

@@
 
+### Fixed Effects Model with Combined Data (training = testing)
+
+```{r fe.cmb}
+if (!exists("fe.cmb.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+head(fe.cmb.result %>% filter(year >= 2005) %>% arrange(desc(num.death)), 20) %>% select(
+  mine.name, num.death, severe, prob, pred
+)
+print(fe.cmb.performance)
+```
+
+### Fixed Effects Model with Divided Data (half training, half testing)
+
+```{r fe.div}
+if (!exists("fe.div.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+head(fe.div.result %>% filter(year >= 2005) %>% arrange(desc(num.death)), 20) %>% select(
+  mine.name, num.death, severe, prob, pred
+)
+print(fe.div.performance)
+```
+
+### Logistic Regression Model without Fixed Effects
+
+```{r glm}
+if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+head(glm.result %>% filter(year >= 2005) %>% arrange(desc(num.death)), 20) %>% select(
+  mine.name, num.death, severe, prob, pred
+)
+print(glm.performance)
+```
+
 ### Simple Linear Model

### 2017-03-19 San Antonio/fixed_effects_luo.R

@@ -1,5 +1,6 @@
 rm(list=ls())
 
+# initialize
 library(dplyr)
 library(survival)
 library(caret)
@@ -11,33 +12,95 @@ sigmoid <- function(x) {
   1 / (1 + exp(-x))
 }
 
+# use nonzero fatality or permenant disability as severe label
 active.quarters <- complete.active.quarters %>% filter(active)
 active.quarters <- active.quarters %>% mutate(
   severe = ifelse(
     num.death + num.dis > 0, TRUE, FALSE
   )
 )
-fe.model <- clogit(severe ~ 
-                     last.quarter.lost+last.year.lost+last.three.years.lost
+
+# fixed effects model with conditional logistic regression on **COMBINED** data
+fe.cmb.model <- clogit(severe ~ last.quarter.lost+last.year.lost+last.three.years.lost
                    +last.quarter.restrict+last.year.restrict+last.three.years.restrict
                    +last.quarter.viol+last.year.viol+last.three.years.viol
                    +last.quarter.death+last.year.death+last.three.years.death
                    + strata(mine_id), data = active.quarters)
-summary(fe.model)
+summary(fe.cmb.model)
 
-
-
-fe.predict <- predict(object = fe.model, type = "lp") %>% sigmoid()
-test <- active.quarters %>% select(mine_id, mine.name, year, quarter, num.death, num.dis, severe)
-test$prob <- fe.predict
-test <- test %>% mutate(
-  prediction = ifelse(
+# test the fixed effects model (see top 20 fatal accidents since 2005)
+fe.cmb.prediction <- predict(object = fe.cmb.model, type = "lp") %>% sigmoid()
+fe.cmb.result <- active.quarters %>% select(mine_id, mine.name, year, quarter, num.death, num.dis, severe)
+fe.cmb.result$prob <- fe.cmb.prediction
+fe.cmb.result <- fe.cmb.result %>% mutate(
+  pred = ifelse(
     prob > 0.5, TRUE, FALSE
   )
 )
-head(test %>% filter(year >= 2005) %>% arrange(desc(num.death)), 20)
+head(fe.cmb.result %>% filter(year >= 2005) %>% arrange(desc(num.death)), 20)
 
-test.performance <- confusionMatrix(data = test$prediction, reference = test$severe, positive = "TRUE")
-print(test.performance)
+# confusion matrix and other performance metrics
+fe.cmb.performance <- confusionMatrix(data = fe.cmb.result$pred, reference = fe.cmb.result$severe, positive = "TRUE")
[…]

### 2017-03-20 San Antonio/presentation_materials.Rmd
slides first draft

--- 2017-03-18 San Antonio/presentation_materials.Rmd
@@
 knitr::opts_chunk$set(echo = FALSE, message=FALSE)
+options(scipen = 1, digits = 2)
 require(dplyr)
@@
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-top.deaths <- complete.active.quarters %>% filter(active & year >= 2005) %>% arrange(desc(num.death)) %>% select(mine.name, mine_id, year, quarter, num.death)
+top.deaths <- complete.active.quarters %>% filter(active & year >= 2005) %>% arrange(desc(num.death), mine.name) %>% select(mine.name, mine_id, year, quarter, num.death)
 head(top.deaths, 10)
@@
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-top.days.lost <- complete.active.quarters %>% filter(active) %>% arrange(desc(num.days.lost)) %>% select(mine.name, year, quarter, num.days.lost)
+top.days.lost <- complete.active.quarters %>% filter(active) %>% arrange(desc(num.days.lost), mine.name) %>% select(mine.name, mine_id, year, quarter, num.days.lost)
 head(top.days.lost, 10)
@@
 
-## Statistical Models
+## Statistical Analyses
+
+### Severe Accidents
+
+```{r severe}
+# options(digits = 1)
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+severe.summary <- complete.active.quarters %>%
+  filter(active) %>%
+  mutate(severe = ifelse(
+    num.death + num.dis > 0, TRUE, FALSE
+  )) %>%
+  group_by(severe) %>%
+  summarise(n = n()) %>%
+  mutate(perc = n / sum(n) * 100)
+print(severe.summary)
+```
+
+### Fixed Effects Model with Combined Data (training = testing)
+
+```{r fe.cmb}
+if (!exists("fe.cmb.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+head(fe.cmb.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 20) %>% select(
+  mine.name, num.death, severe, prob, pred
+) %>% mutate (mine.name = strtrim(mine.name, 40))
+head(fe.cmb.result %>% filter(year >= 2005) %>% arrange(desc(prob), mine.name), 20) %>% select(
+  mine.name, num.death, severe, prob, pred
+) %>% mutate (mine.name = strtrim(mine.name, 40))
+print(fe.cmb.performance)
+```
+
+### Fixed Effects Model with Divided Data (half training, half testing)
+
+```{r fe.div}
+if (!exists("fe.div.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+head(fe.div.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 20) %>% select(
+  mine.name, num.death, severe, prob, pred
+) %>% mutate (mine.name = strtrim(mine.name, 40))
+head(fe.div.result %>% filter(year >= 2005) %>% arrange(desc(prob), mine.name), 20) %>% select(
+  mine.name, num.death, severe, prob, pred
+) %>% mutate (mine.name = strtrim(mine.name, 40))
+print(fe.div.performance)
+```
+
+### Logistic Regression Model without Fixed Effects
+
+```{r glm}
+if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+head(glm.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 20) %>% select(
+  mine.name, num.death, severe, prob, pred
+) %>% mutate (mine.name = strtrim(mine.name, 40))
+head(glm.result %>% filter(year >= 2005) %>% arrange(desc(prob), mine.name), 20) %>% select(
+  mine.name, num.death, severe, prob, pred
+) %>% mutate (mine.name = strtrim(mine.name, 40))
+print(glm.performance)
+```
 
@@
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-num.deaths.lm <- lm(1*num.death+0*num.days.lost+0*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost
+num.deaths.lm <- lm(1*num.death+1*num.dis +0*num.days.lost+0*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost
            +last.quarter.restrict+last.year.restrict+last.three.years.restrict

### 2017-03-20 San Antonio/slides/presentation_slides.Rmd
slides first draft

@@
 shortauthor: "Luo et al."
-institute: "Columbia University"
+institute: |
+  | Chemical Engineering, Computer Science, and Business School
+  | Columbia University
+shortinstitute: "Columbia University"
 date: "3/27/2017"
@@
   beamer_presentation:
-    # incremental: true
+    keep_tex: false
+    incremental: false
     template: rmd_template.tex
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message=FALSE)
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, fig.align = "center")
+options(scipen = 1, digits = 2)
 require(dplyr)
@@
 
-# Introduction: Mine Safety and MSHA Data
+## Complex, Resilient, Intelligent Systems (CRIS Lab)
+
+![CRIS Lab](CrisResearch720.png)
+
+<!-- https://aiche.confex.com/aiche/s17/webprogram/Paper480429.html -->
+# Introduction: A Data Approach to Mine Safety
+
+## Systemic Risk
+
+- Systemic disasters
+    - SARS (2003)
+    - Northeast Blackout (2003)
+    - Subprime Crisis (2008)
+    - Deepwater Horizon Oil Spill (2010)
+- Emerging systemic risks
+    - Climate change
+    - Income/wealth inequality
+    - Cyber-physical security
+    - Technological singularity
+- Fast-paced and connected
+- Design complex systems
+- Analyze systemic risk
+
+<!-- 
+https://en.wikipedia.org/wiki/Upper_Big_Branch_Mine_disaster
+https://arlweb.msha.gov/MEDIA/PRESS/2011/NR111206.asp 
+-->
+## Upper Big Branch Mine Disaster (2010)
+
+- April 5, 2010, Raleigh County, West Virginia, owned by Massey Energy
+- 29 deaths, the worst mining in the United States since 1970
+- MSHA cites corporate culture as root cause of Upper Big Branch Mine disaster
+
+## Sago Mine Disaster (2006)
+
+- January 2, 2006, Sago, West Virginia, owned by Anker West Virginia Mining
+- 13 miners were trapped for nearly two days; only one survived
+- Fatality number was exceeded by the Upper Big Branch Mine disaster
+- MSHA reports prior history of safety violations and fatalities
 
@@
 
-## Upper Big Branch Disaster
+## Can We Further Improve Mine Safety?
 
-# Methods: Data and Model Assumptions
+- Process MSHA safety data
+- Understand the underlying causal relationships
+- Develop early warning systems
+
+# Methods: Data Sources and Model Preliminaries
 
@@
 
-- **Mine accidents**: "msha_accident.csv" (681,386 obs., retrieved 1/26/2017, from <https://enforcedata.dol.gov/views/data_summary.php>)
-- **MSHA assessed violations**: "AssessedViolations.csv" (2,169,804 obs., retrieved 12/10/2016, from <https://arlweb.msha.gov/OpenGovernmentData/OGIMSHA.asp>)
+- Mine accidents table: "msha_accident.csv" 
+    - 681,386 rows
+    - Retrieved 1/26/2017, from <https://enforcedata.dol.gov/views/data_summary.php>
+- MSHA assessed violations table: "AssessedViolations.csv" 
+    - 2,169,804 rows
+    - Retrieved 12/10/2016, from <https://arlweb.msha.gov/OpenGovernmentData/OGIMSHA.asp>
+
+## Master Data Table
+
+- 664,128 rows
+- 10,377 unique mines
+- From 2000 to 2015 in quarters
+- Each row represents data for a unique combination of mine, year, and quarter
+    - e.g., Upper Big Branch Mine in the second quarter of 2010
+- Each row contains both current and past information
 
@@
 
-## Model
+## Top 10 Fatal Accidents Since 2005
+
+```{r top 10 fatality}
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+top.deaths <- complete.active.quarters %>% filter(active & year >= 2005) %>% arrange(desc(num.death), mine.name) %>% select(mine.name, mine_id, year, quarter, num.death)
+head(top.deaths, 10)
+```
+
+## Violation Trend: Upper Big Branch
+
+```{r upper big branch}
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 4608436 & year < 2010) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "Count")
+```
+
+## Violation Trend: Sago Mine
+
+```{r sago}
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 4608791 & year < 2006) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "Count")
+```
+
+## Violation Trend: Crandall Canyon
+
+```{r crandall}
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 4201715 & year < 2007) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "Count")
+```
+
+## Violation Trend: Darby Mine No. 1
+
+```{r darby}
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 1518185 & year < 2006) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "Count")
+```
+
+## Violation Trend: Gibson
+
+```{r gibson}
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+yearly.viol <- complete.active.quarters %>% 
+  filter(mine_id == 1202215 & year < 2007) %>%
+  group_by(year) %>%
+  summarise(count = sum(viol.quantity))
+qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
+      main = "Number of Yearly Violations",
+      xlab = "Year", ylab = "Count")
+```
+
+## Predictive Model
+
+- Rising violation trends before disasters
+- A disaster classifier based on historical data?
+- Define a **severe** accident as one with death or permenant disability
+- Unbalanced data
+
+```{r severe all}
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+severe.summary <- complete.active.quarters %>%
+  filter() %>%
+  mutate(severe = ifelse(
+    num.death + num.dis > 0, TRUE, FALSE
+  )) %>%
+  group_by(severe) %>%
+  summarise(n = n()) %>%
+  mutate(perc = n / sum(n) * 100)
+print(severe.summary)
+```
 
@@
 
-## Number of Days Lost since 2000
+## Processing Data
 
-```{r num.days.lost trend}
+- Remove **inactive** quarters
+
+```{r severe active}
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-yearly.lost <- complete.active.quarters %>% 
-  filter(year >= 2000 & active) %>%
-  group_by(year) %>%
-  summarise(count = sum(num.days.lost),
-            avg = mean(num.days.lost))
-qplot(x = year, y = count, data = yearly.lost, geom = c("point", "line"),
-      main = "Number of Total Days Lost Due to Accidents",
-      xlab = "Year", ylab = "Count")
-# qplot(x = year, y = avg, data = yearly.lost, geom = c("point", "line"),
-#       main = "Number of Average Days Lost Due to Accidents",
-#       xlab = "Year", ylab = "Count")
+severe.summary <- complete.active.quarters %>%
+  filter(active) %>%
+  mutate(severe = ifelse(
+    num.death + num.dis > 0, TRUE, FALSE
+  )) %>%
+  group_by(severe) %>%
+  summarise(n = n()) %>%
+  mutate(perc = n / sum(n) * 100)
+print(severe.summary)
 ```
 
-## Number of Violations since 2000
+<!-- http://www.stata.com/manuals13/rclogit.pdf -->
+## Processing Data
 
-```{r viol.quantity trend}
+- Consider fixed-mine effects
+    - Biostatisticians and epidemiologists call it "conditional logistic regression" (R, survival::clogit)
+    - Suitable for panel data (e.g., our master data table)
+    - Model includes mine-specific but time-invariant variables (e.g., same slope but different intercepts for different mines)
+
+## Logistic Regression Without Fixed Effects
+
+- Train and test on all data
+
+```{r glm}
+if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+print(glm.performance$table)
+print(glm.performance$byClass)
+```
+
+## Logistic Regression Without Fixed Effects
+
+- Sort result based on number of deaths
+
+```{r glm rank death}
+if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+head(glm.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10) %>% select(
+  mine.name, num.death, severe, prob, pred
+) %>% mutate (mine.name = strtrim(mine.name, 40))
+```
+
+<!--
+## Logistic Regression Without Fixed Effects
+
+- Sort result based on model probability
+
+```{r glm rank prob}
+if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+head(glm.result %>% filter(year >= 2005) %>% arrange(desc(prob)), 10) %>% select(
+  mine.name, num.death, severe, prob, pred
+) %>% mutate (mine.name = strtrim(mine.name, 40))
+```
+-->
+
+## Logistic Regression with Fixed Effects
+
+- Randomly select half of the data to train and the other half to test
+
+```{r fe.div}
+if (!exists("fe.div.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+print(fe.div.performance$table)
+print(fe.div.performance$byClass)
+```
+
+## Logistic Regression with Fixed Effects
+
+- Sort result based on number of deaths
+
+```{r fe.div rank death}
+if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+head(fe.div.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10) %>% select(
+  mine.name, num.death, severe, prob, pred
+) %>% mutate (mine.name = strtrim(mine.name, 40))
+```
+
+## Logistic Regression with Fixed Effects
+
+- Sort result based on model probability
+
+```{r fe.div rank prob}
+if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+head(fe.div.result %>% filter(year >= 2005) %>% arrange(desc(prob)), 10) %>% select(
+  mine.name, num.death, severe, prob, pred
+) %>% mutate (mine.name = strtrim(mine.name, 40))
+```
+
+## New Era Mine
+
+- The American Coal Company New Era Mine occupies 6 slots on the top 10 list of most days lost due to accidents
+
+```{r top 10 days lost}
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+top.days.lost <- complete.active.quarters %>% filter(active) %>% arrange(desc(num.days.lost), mine.name) %>% select(mine.name, year, quarter, num.days.lost)
+head(top.days.lost, 10)
+```
+
+## New Era Mine
+
+<!--- Violation trend prior to the first quarter in 2006 where model probability is 0.99-->
+
+```{r new era viol}
 if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
 yearly.viol <- complete.active.quarters %>% 
-  filter(year >= 2000 & active) %>%
+  filter(mine_id == 1102752 & year < 2006) %>%
   group_by(year) %>%
-  summarise(count = sum(viol.quantity),
-            avg = mean(viol.quantity))
+  summarise(count = sum(viol.quantity))
 qplot(x = year, y = count, data = yearly.viol, geom = c("point", "line"),
-      main = "Number of Total Violations",
+      main = "Number of Yearly Violations",
       xlab = "Year", ylab = "Count")
@@
 
+## Simple Linear Model on Number of Days Lost
+
+- Adjusted $R^2=0.36$
+
+```{r simple lm of days lost}
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+num.days.lm <- lm(num.days.lost~last.quarter.lost+last.year.lost+last.three.years.lost
+           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
+           +last.quarter.viol+last.year.viol+last.three.years.viol
+           +last.quarter.death+last.year.death+last.three.years.death, 
+           data=complete.active.quarters %>% filter(active))
+summary(num.days.lm)$coefficients
+```
+
 # Conclusion
+
+## Conclusion
+
+<!-- - Causal relationship between accidents and historical data -->
+- TBD
 

### 2017-03-20 San Antonio/slides/systemic_risk.tex
slides first draft

\begin{frame}{Systemic Risk}\pause
\begin{columns}[c]
	\column{0.475\linewidth}
		\begin{itemize}[<+->]
		\item Systemic disasters
		\begin{itemize}
			\item SARS (2003)
			\item Northeast Blackout (2003)
			\item Subprime Crisis (2008)
			\item Deepwater Horizon
			%Oil Spill 
			(2010)
		\end{itemize}
		\item Emerging systemic risks
		\begin{itemize}
			\item Climate change
			\item Income/wealth inequality
			\item Cyber-physical security
			\item Technological singularity
		\end{itemize}
		\item Fast-paced and connected
		\item Design complex systems
		\item Analyze systemic risk
		\end{itemize}
	\column{0.525\linewidth}
	   \onslide<3->
	   \includegraphics[width = .5\linewidth]{fig/wide_SARS.png}
	   \onslide<8->
	   \includegraphics[width = .5\linewidth]{fig/wide_polarbear.png}\\%http://i.dailymail.co.uk/i/pix/2014/09/09/1410255561618_wps_15_Spitsbergen_Norway_Polar_.jpg
	   \onslide<4->
	   \includegraphics[width = .5\linewidth]{fig/wide_NEB.png}
	   \onslide<9->
	   \includegraphics[width = .5\linewidth]{fig/wide_99.png}\\%https://c1.staticflickr.com/7/6055/6247188828_305409d56a_b.jpg
	   \onslide<5->
	   \includegraphics[width = .5\linewidth]{fig/wide_FIN.png}
	   \onslide<10->
	   \includegraphics[width = .5\linewidth]{fig/wide_cyber.png}\\%https://c1.staticflickr.com/9/8604/16042227002_1d00e0771d_b.jpg
	   \onslide<6->
	   \includegraphics[width = .5\linewidth]{fig/wide_BP.png}
	   \onslide<11->
	   \includegraphics[width = .5\linewidth]{fig/wide_exmachina.png}%https://c1.staticflickr.com/1/569/20167701293_e3ed4b066f_b.jpg
\end{columns}
\end{frame}

### 2017-03-21 San Antonio/markov_chains.Rmd
Update markov_chains.Rmd

@@
 ---
+
+This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+
+When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
+You can embed an R code chunk like this:
 
@@
 require(reshape)
+
 ```
-```{r setting directories}
+
+```{r}
 knitr::opts_knit$set(root.dir = "~/Documents/Violation_Project")
@@
 aggregate_cluster_seq <- cast(melted, mine_id~date)
+
+aggregate_cluster_seq
 ```
@@
     cluster_sequence <- cluster_seq_raw[2:length(cluster_seq_raw)]
-    cur_markov_chain <- markovchainFit(data=cluster_sequence)
-    return (list("mine_id" = x[1], "markov_chain" = cur_markov_chain$estimate@transitionMatrix, "num_rows" = length(cluster_seq_raw)))
+    cur_markov_chain <- createSequenceMatrix(cluster_sequence, sanitize = FALSE)
+    return (list("mine_id" = x[1], 
+                 "markov_chain" = cur_markov_chain, 
+                 "num_rows" = length(cluster_seq_raw),
+                 "low" = 2 %in% cluster_sequence, "mid" = 1 %in% cluster_sequence, "high" = 3 %in% cluster_sequence
+                 )
+            )
   }
-  return (list("mine_id" = x[1], "markov_chain" = NA, "num_rows" = 0))
+  return (list("mine_id" = x[1], 
+               "markov_chain" = NA, 
+               "num_rows" = 0,
+               "low" = FALSE, "mid" = FALSE, "high" = FALSE
+               )
+          )
 }
+
 ```
@@
 ```{r}
-num_rows_ind <- sapply(result, "[[", 3)
-weights <- num_rows_ind/sum(num_rows_ind)
-
+#transition matrix
 markov_chains_ind <- sapply(result, "[[", 2)
-
-weighted_transition <- mapply(function(x,y) x * y, x = weights, y = markov_chains_ind)
 
@@
 for(i in 1:length(weighted_transition)){
-  final_transition <- add_matrices_1(final_transition,weighted_transition[[i]])
+  final_transition <- add_matrices_1(final_transition,markov_chains_ind[[i]])
 }
+print("Tranition matrix frequency")
 final_transition
 
+final_markov<- matrix(rep(0,times= 9), ncol = 3, nrow = 3)
+for(i in 1:3){
+  final_markov[i,] <- final_transition[i,]/sum(final_transition[i,])  
+}
+
+print("Transition matrix probability")
+final_markov
+
 ```

### 2017-03-21 San Antonio/slides/presentation_slides.Rmd
updated slides

@@
 - Understand the underlying causal relationships
-- Develop early warning systems
+- Develop early warning systems based on past behaviors
+- Credit rating/score analogy
+    - Predict default probability within 18 months
+    - Accidents: defaults a month or a year prior to application
+    - Violations: missed payments, late payments, etc.
 
@@
 
-## Primary Datasets
+## Department of Labor Enforcement Data
+
+- Link: <https://enforcedata.dol.gov/views/data_catalogs.php>
+- Updated daily or weekly
+- Enforecement Data Catalog
+    - MSHA Enforcement Data
+    - OSHA Enforcement Data
+    - Wage and Hour Compliance Action Data
+    - OFCCP Compliance Evaluation and Complaint Investigation Data
+    - EBSA Enforcement Data
+
+## MSHA Data: Sources
 
@@
 
+## MSHA Data: Advantages
+
+- Each mine has a unique mine ID, e.g., Upper Big Branch (4608436)
+- Rich details, e.g., selected attributes from the accidents table:
+
+```{r msha accidents}
+if(!exists("accidents")) {load("./San Antonio/data/secret_ingredients.RData")}
+names(accidents[c(1, 14, 5:13, 19:20, 37:59)])
+```
+
+## MSHA Data: Challenges
+
+- Missing data
+- Inactive mines
+- Most data are not numeric
+- Relatively few severe accidents ($\sim0.5\%$)
+
+## Model Preliminaries
+
+- Create a master data table with selected attributes
+- Plot violation trends for Upper Big Branch Mine and Sago Mine
+- Design a predictive model
+
 ## Master Data Table
 
-- 664,128 rows
-- 10,377 unique mines
+- Combine and summarize accidents/violations based on mines
+- 664,128 rows, 10,377 unique mines
 - From 2000 to 2015 in quarters
@@
 
+<!--
 ## Violation Trend: Crandall Canyon
@@
 ```
+-->
 
@@
 - Unbalanced data
-
-```{r severe all}
-if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-severe.summary <- complete.active.quarters %>%
-  filter() %>%
-  mutate(severe = ifelse(
-    num.death + num.dis > 0, TRUE, FALSE
-  )) %>%
-  group_by(severe) %>%
-  summarise(n = n()) %>%
-  mutate(perc = n / sum(n) * 100)
-print(severe.summary)
-```
-
-# Results and Discussion
-
-## Processing Data
-
-- Remove **inactive** quarters
 
@@
 <!-- http://www.stata.com/manuals13/rclogit.pdf -->
-## Processing Data
+## Fixed-Mine Effects
 
-- Consider fixed-mine effects
-    - Biostatisticians and epidemiologists call it "conditional logistic regression" (R, survival::clogit)
+- Biostatisticians and epidemiologists call it "conditional logistic regression" (`survival::clogit`)
 - Suitable for panel data (e.g., our master data table)
 - Model includes mine-specific but time-invariant variables (e.g., same slope but different intercepts for different mines)
+- Logistic regression (for every mine)
+
+\[\Pr(Y=1|\mathbf{X})=F(\mathbf{X}) = \frac{1}{1+e^{-(\alpha+\boldsymbol\beta\mathbf{X})}}\]
+
+- Logistic regression with fixed effects (for the $i$-th mine)
+
+\[\Pr(Y=1|\mathbf{X},i)=F(\mathbf{x},i) = \frac{1}{1+e^{-(\alpha_i+\boldsymbol\beta\mathbf{x})}}\]
+
+# Results and Discussion
 
@@
 
-- Sort result based on number of deaths
+- True positives
 
@@
 head(glm.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10) %>% select(
-  mine.name, num.death, severe, prob, pred
+  mine.name, year, quarter, severe, pred
 ) %>% mutate (mine.name = strtrim(mine.name, 40))
@@
 
-<!--
 ## Logistic Regression Without Fixed Effects
 
-- Sort result based on model probability
+- Top predicted positives (possibly false positives)
 
@@
 if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
-head(glm.result %>% filter(year >= 2005) %>% arrange(desc(prob)), 10) %>% select(
-  mine.name, num.death, severe, prob, pred
+head(glm.result %>% filter(year >= 2005) %>% arrange(desc(prob), mine.name), 10) %>% select(
+  mine.name, year, quarter, severe, pred
 ) %>% mutate (mine.name = strtrim(mine.name, 40))
 ```
--->
 
@@
 
-- Sort result based on number of deaths
+- Top true positives (deadliest accidents)
 
 ```{r fe.div rank death}
-if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+if (!exists("fe.div.result")){load("./San Antonio/output/results_fixed_effects.RData")}
 head(fe.div.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10) %>% select(
-  mine.name, num.death, severe, prob, pred
+  mine.name, year, quarter, severe, pred
 ) %>% mutate (mine.name = strtrim(mine.name, 40))
@@
 
-- Sort result based on model probability
+- Top predicted positives (possibly false positives)
 
 ```{r fe.div rank prob}
-if (!exists("glm.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+if (!exists("fe.div.result")){load("./San Antonio/output/results_fixed_effects.RData")}
 head(fe.div.result %>% filter(year >= 2005) %>% arrange(desc(prob)), 10) %>% select(
-  mine.name, num.death, severe, prob, pred
+  mine.name, year, quarter, severe, pred
 ) %>% mutate (mine.name = strtrim(mine.name, 40))
@@
 
+<!--
 ## New Era Mine
@@
 yearly.viol <- complete.active.quarters %>% 
-  filter(mine_id == 1102752 & year < 2006) %>%
+  filter(mine_id == 1102752) %>%
   group_by(year) %>%
@@
 
-## Simple Linear Model on Number of Days Lost
+## New Labels Including Days Lost
+
+- Previously defined severe accidents and days lost $>$ 300
+
+```{r fe.div2}
+if (!exists("fe.div2.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+print(fe.div2.performance$table)
+print(fe.div2.performance$byClass)
+```
+
+## New Labels Including Days Lost
+
+- Top true positives (deadliest accidents)
+
+```{r fe.div2 rank death}
+if (!exists("fe.div2.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+head(fe.div2.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10) %>% select(
+  mine.name, year, quarter, severe, pred
+) %>% mutate (mine.name = strtrim(mine.name, 40))
+```
+
+## New Labels Including Days Lost
+
+- Top predicted positives (possibly false positives)
+
+```{r fe.div2 rank prob}
+if (!exists("fe.div2.result")){load("./San Antonio/output/results_fixed_effects.RData")}
+head(fe.div2.result %>% filter(year >= 2005) %>% arrange(desc(prob)), 10) %>% select(
+  mine.name, year, quarter, severe, pred
+) %>% mutate (mine.name = strtrim(mine.name, 40))
+```
+
+# Conclusion
+
+## Conclusion
+
+<!-- - Causal relationship between accidents and historical data -->
+- Summary
+    - Two deadliest mine accidents in the last decade: Upper Big Branch & Sago
+    - MSHA
+
+## Appendix: Simple Linear Model
 
@@
 
-# Conclusion
-
-## Conclusion
-
-<!-- - Causal relationship between accidents and historical data -->
-- TBD
-
 <!-- USEFUL LINKS:

### 2017-03-21 San Antonio/fixed_effects_luo.R

@@ -1,6 +1,6 @@
 rm(list=ls())
 
-# initialize
+# initialize ----
 library(dplyr)
 library(survival)
 library(caret)
@@ -12,90 +12,131 @@ sigmoid <- function(x) {
   1 / (1 + exp(-x))
 }
 
-# use nonzero fatality or permenant disability as severe label
-active.quarters <- complete.active.quarters %>% filter(active)
-active.quarters <- active.quarters %>% mutate(
+# label data ----
+labeled.data.death <- complete.active.quarters %>% filter(active) %>% mutate(
   severe = ifelse(
     num.death + num.dis > 0, TRUE, FALSE
   )
 )
+labeled.data.lost <- complete.active.quarters %>% filter(active) %>% mutate(
+  severe = ifelse(
+    num.death + num.dis > 0 | num.days.lost > 300, TRUE, FALSE
+  )
+)
 
-# fixed effects model with conditional logistic regression on **COMBINED** data
+# fixed effects on combined data ----
+# fixed effects model
 fe.cmb.model <- clogit(severe ~ last.quarter.lost+last.year.lost+last.three.years.lost
                    +last.quarter.restrict+last.year.restrict+last.three.years.restrict
                    +last.quarter.viol+last.year.viol+last.three.years.viol
                    +last.quarter.death+last.year.death+last.three.years.death
-                   + strata(mine_id), data = active.quarters)
+                   + strata(mine_id), data = labeled.data.death)
 summary(fe.cmb.model)
 
 # test the fixed effects model (see top 20 fatal accidents since 2005)
 fe.cmb.prediction <- predict(object = fe.cmb.model, type = "lp") %>% sigmoid()
-fe.cmb.result <- active.quarters %>% select(mine_id, mine.name, year, quarter, num.death, num.dis, severe)
+fe.cmb.result <- labeled.data.death %>% select(mine_id, mine.name, year, quarter, num.death, num.dis, severe)
 fe.cmb.result$prob <- fe.cmb.prediction
 fe.cmb.result <- fe.cmb.result %>% mutate(
   pred = ifelse(
     prob > 0.5, TRUE, FALSE
   )
 )
-head(fe.cmb.result %>% filter(year >= 2005) %>% arrange(desc(num.death)), 20)
+head(fe.cmb.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10)
[…]

### 2017-03-21 San Antonio/slides/presentation_slides.Rmd
updated slides

@@
     - OSHA Enforcement Data
-    - Wage and Hour Compliance Action Data
-    - OFCCP Compliance Evaluation and Complaint Investigation Data
-    - EBSA Enforcement Data
+    <!-- - Wage and Hour Compliance Action Data -->
+    <!-- - OFCCP Compliance Evaluation and Complaint Investigation Data -->
+    <!-- - EBSA Enforcement Data -->
+    <!-- - Department of Justice -->
+    - EPA
+    - FDA
 
@@
 
-- Missing data
-- Inactive mines
+- Missing data, typos
+- Inactive mines are not labeled
 - Most data are not numeric
@@
 
-## Model Preliminaries
+<!-- ## Model Preliminaries -->
 
-- Create a master data table with selected attributes
-- Plot violation trends for Upper Big Branch Mine and Sago Mine
-- Design a predictive model
+<!-- - Create a master data table with selected attributes -->
+<!-- - Plot violation trends for Upper Big Branch Mine and Sago Mine -->
+<!-- - Conduct statistical analyses -->
 
-## Master Data Table
+## Consolidated Data
 
@@
 
-## Master Data Table
+## Consolidated Data
+
+- Number of days lost (`num.days.lost`), numer of days restricted (`num.days.restrict`), number of deaths (`num.death`), number of permenant disabilities (`num.dis`), and number of violations (`viol.quantity`)
 
@@
 - Biostatisticians and epidemiologists call it "conditional logistic regression" (`survival::clogit`)
-- Suitable for panel data (e.g., our master data table)
-- Model includes mine-specific but time-invariant variables (e.g., same slope but different intercepts for different mines)
+- Suitable for **panel data** (e.g., our consolidated data)
+- Model includes mine-specific but time-invariant variables
 - Logistic regression (for every mine)
@@
 
-- Train and test on all data
+<!-- 解释 confusion matrix和 specificity -->
+
+- In-sample model
 
@@
 
-- True positives
+- Fail to predict top 10 true positives
 
@@
 
-- Top predicted positives (possibly false positives)
+- False positive predictions
+<!-- - Top predicted positives (possibly false positives) -->
 
@@
 
-- Randomly select half of the data to train and the other half to test
+- Out-of-sample model (randomly select half of the data to train and the other half to test)
 
@@
 
-- Top true positives (deadliest accidents)
+- Successfully predict all top 10 true positives
 
@@
 
-- Top predicted positives (possibly false positives)
+- False positive predictions
 
@@
 <!--- Violation trend prior to the first quarter in 2006 where model probability is 0.99-->
+<!-- 即便没有灾难，也有趋势，警示作用，后来可能改善了 -->
 
@@
 
-- Top true positives (deadliest accidents)
+- Successfully predict 9 out of top 10 true positives
 
@@
 
-- Top predicted positives (possibly false positives)
+-
 
@@
 ```
+
+## Unsupervised Clustering
+
+## Markov Chain
 
@@
     - MSHA 
+- Application
+    - "Credit score" for mine safety
+    - Regulators, mines, stakeholders
+- Future
+    - Expand data: OSHA, EPA, etc.
+    - Models and techniques: artificial neural networks (restricted boltzmann machine), text mining, etc.
 

### 2017-03-22 San Antonio/markov_chains_luo.R
updated markov by luo

@@ -1,0 +1,94 @@
+rm(list=ls())
+
+# initialize ----
+require(dplyr)
+require(markovchain)
+require(reshape)
+setwd("~/Git/violation-data-analysis")
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+
+# clustering ----
+temp <- complete.active.quarters %>% filter(active==TRUE)
+set.seed(20)
+clustering <- kmeans(temp[, 6:ncol(temp)], 3, nstart=20)
+freq_table <- table(clustering$cluster)
+print(clustering$centers)
+print(clustering$size)
+
+# assign low, mid, high risk labels
+unsorted.risk.label <- c("low", "mid", "high")
+sorted.risk.labels <- sort.int(clustering$centers[,1], index.return = TRUE)$ix
+
+# Melted format containing quarter-year, mine_id allowing for markov chain analysis ----
+temp$cluster <- clustering$cluster
+temp$date <- paste(as.character(temp$year), "-" , as.character(temp$quarter))
+temp2 <- temp %>% select(mine_id, cluster, date)
+melted <- melt(temp2, id=c("mine_id", "date"))
+aggregate_cluster_seq <- cast(melted, mine_id~date)
+
+# Markov Chain function ----
+markov <- function(x){
+  x <- as.integer(x)
+  cluster_seq_raw <- x[!is.na(x)]
+  if(length(cluster_seq_raw) > 1){
+    cluster_sequence <- cluster_seq_raw[2:length(cluster_seq_raw)]
+    cur_markov_chain <- createSequenceMatrix(cluster_sequence, sanitize = FALSE)
+    return (list("mine_id" = x[1], 
+                 "markov_chain" = cur_markov_chain, 
+                 "num_rows" = length(cluster_seq_raw),
+                 "low" = sorted.risk.labels[1] %in% cluster_sequence, 
+                 "mid" = sorted.risk.labels[2] %in% cluster_sequence, 
+                 "high" = sorted.risk.labels[3] %in% cluster_sequence
+    )
+    )
+  }
+  return (list("mine_id" = x[1], 
+               "markov_chain" = NA, 
+               "num_rows" = 0,
+               "low" = FALSE, "mid" = FALSE, "high" = FALSE
+  )
+  )
+}
+
+# create mine-date transition matrices ----
+result<-alply(aggregate_cluster_seq, 1, markov)
+
+# create overall transition matrix ----
[…]

### 2017-03-22 San Antonio/slides/presentation_slides.Rmd
updated markov by luo
semi-final slides

@@
     fig_height: 4.75
+    # fig_width: 4
     toc: true
@@
 require(ggplot2)
+require(markovchain)
 knitr::opts_knit$set(root.dir = "~/Git/violation-data-analysis")
@@
 <!-- https://aiche.confex.com/aiche/s17/webprogram/Paper480429.html -->
-# Introduction: A Data Approach to Mine Safety
+# Mine Safety: A Data-Driven Approach
 
 ## Systemic Risk
+
+\placelogofalse
 
@@
 
-<!-- 
-https://en.wikipedia.org/wiki/Upper_Big_Branch_Mine_disaster
-https://arlweb.msha.gov/MEDIA/PRESS/2011/NR111206.asp 
--->
+<!-- https://en.wikipedia.org/wiki/Upper_Big_Branch_Mine_disaster -->
+<!-- https://arlweb.msha.gov/MEDIA/PRESS/2011/NR111206.asp  -->
 ## Upper Big Branch Mine Disaster (2010)
@@
     - Provide technical, educational, and other types of assistance
+- A constantly improving industry in terms of safety
 
@@
     - Violations: missed payments, late payments, etc.
+- Can we develop a "credit score" for mine safety?
 
@@
 - Updated daily or weekly
-- Enforecement Data Catalog
-    - MSHA Enforcement Data
-    - OSHA Enforcement Data
+- Publicly available 
+    - Department of Labor: MSHA, OSHA, etc.
+    - Other departments: EPA, FDA, DOJ, etc.
+
+    <!-- - MSHA Enforcement Data -->
+    <!-- - OSHA Enforcement Data -->
     <!-- - Wage and Hour Compliance Action Data -->
@@
     <!-- - Department of Justice -->
-    - EPA
-    - FDA
+    <!-- - EPA -->
+    <!-- - FDA -->
 
@@
 - Each mine has a unique mine ID, e.g., Upper Big Branch (4608436)
-- Rich details, e.g., selected attributes from the accidents table:
+- Rich details: e.g., classification, description, and severity
+- Selected attributes from the accidents table (omitting 42 attributes)
 
@@
 if(!exists("accidents")) {load("./San Antonio/data/secret_ingredients.RData")}
-names(accidents[c(1, 14, 5:13, 19:20, 37:59)])
+names(accidents[c(1, 14, 19:20, 5:13, 37, 39, 49:50)])
 ```
@@
 
-- Missing data, typos
-- Inactive mines are not labeled
+- Missing data, human errors
+- No information about inactive/nonoperating mines
 - Most data are not numeric
-- Relatively few severe accidents ($\sim0.5\%$)
+- Lots of zeros, few severe accidents ($\sim0.5\%$)
 
@@
 
-- Combine and summarize accidents/violations based on mines
+- Group and summarize accidents/violations by mines
 - 664,128 rows, 10,377 unique mines
-- From 2000 to 2015 in quarters
+- From 2000 to 2015
 - Each row represents data for a unique combination of mine, year, and quarter
@@
 - Each row contains both current and past information
+    - i.e., current quarter, past quarter, past year, and past three years
 
@@
 
-- Number of days lost (`num.days.lost`), numer of days restricted (`num.days.restrict`), number of deaths (`num.death`), number of permenant disabilities (`num.dis`), and number of violations (`viol.quantity`)
+<!-- - Number of days lost (`num.days.lost`), numer of days restricted (`num.days.restrict`), number of deaths (`num.death`), number of permenant disabilities (`num.dis`), and number of violations (`viol.quantity`)  -->
+
+- All 25 attributes of the consolidated data
 
@@
 
+- Query the consolidated data on the deadliest accidents
+
 ```{r top 10 fatality}
@@
 ```
+
+- Plot violation trends prior to Upper Big Branch and Sago disasters
 
@@
 - Rising violation trends before disasters
-- A disaster classifier based on historical data?
+- Develop a disaster classifier based on historical data
 - Define a **severe** accident as one with death or permenant disability
-- Unbalanced data
+<!-- - Unbalanced data -->
 
@@
 - Biostatisticians and epidemiologists call it "conditional logistic regression" (`survival::clogit`)
-- Suitable for **panel data** (e.g., our consolidated data)
-- Model includes mine-specific but time-invariant variables
-- Logistic regression (for every mine)
+- Suitable for **panel data** (e.g., longitudinal data, our consolidated data)
+- Model includes mine-specific biases
+- Logistic function (for every mine)
 
-\[\Pr(Y=1|\mathbf{X})=F(\mathbf{X}) = \frac{1}{1+e^{-(\alpha+\boldsymbol\beta\mathbf{X})}}\]
+\[\Pr(Y=1|\mathbf{X}) = \frac{1}{1+e^{-(\alpha+\boldsymbol\beta\mathbf{X})}}\]
 
-- Logistic regression with fixed effects (for the $i$-th mine)
+- Logistic function with fixed effects (for the $i$-th mine)
 
-\[\Pr(Y=1|\mathbf{X},i)=F(\mathbf{x},i) = \frac{1}{1+e^{-(\alpha_i+\boldsymbol\beta\mathbf{x})}}\]
+\[\Pr(Y=1|\mathbf{X},i) = \frac{1}{1+e^{-(\alpha_i+\boldsymbol\beta\mathbf{x})}}\]
 
@@
 ## Logistic Regression Without Fixed Effects
-
-<!-- 解释 confusion matrix和 specificity -->
 
@@
 print(glm.performance$table)
-print(glm.performance$byClass)
+print(c(glm.performance$overall["Accuracy"], glm.performance$byClass[c("Sensitivity", "Specificity", "Precision", "F1")]))
 ```
+
+- Accuracy = (TP + TN)/(P + N)
+- Sensitivity/recall = TP/P
+- Specificity = TN/N
+- Precision = TP/(TP + FP)
+- F1: harmonic mean of sensitivity and precision
 
@@
 
-- Fail to predict top 10 true positives
+- Fails to predict top 10 deadliest disasters
 
@@
 
-- False positive predictions
+- List of false positive predictions based on predicted probability
 <!-- - Top predicted positives (possibly false positives) -->
@@
 print(fe.div.performance$table)
-print(fe.div.performance$byClass)
+print(c(fe.div.performance$overall["Accuracy"], fe.div.performance$byClass[c("Sensitivity", "Specificity", "Precision", "F1")]))
 ```
@@
 
-- Successfully predict all top 10 true positives
+- Successfully predicts top 10 deadliest disasters
 
@@
 
-- False positive predictions
+- Still has a lot of false positive predictions
+- List of false positive predictions based on predicted probability
 
@@
 ```
+
+- What happened in the New Era Mine?
 
@@
 -->
+
+## New Era Mine
+
+- Among the worst mines by number of days lost due to accidents
+
+```{r new-era-num-days-lost}
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+head(
+  complete.active.quarters %>% filter(active) %>% arrange(desc(num.days.lost)) %>% select(
+    mine.name, year, quarter, num.days.lost
+  )
+, 10)
+```
+
+- Rising violation trend from 2000 to 2005
 
@@
 
-- Previously defined severe accidents and days lost $>$ 300
+- Updated severe accident label
+    - Previously defined criteria plus days lost $>$ 300
+- Redo out-of-sample model
 
@@
 print(fe.div2.performance$table)
-print(fe.div2.performance$byClass)
+print(c(fe.div2.performance$overall["Accuracy"], fe.div2.performance$byClass[c("Sensitivity", "Specificity", "Precision", "F1")]))
 ```
+
+- Worse true positive rate, improved F1 score
 
@@
 
-- Successfully predict 9 out of top 10 true positives
+- Successfully predicts 9 out of top 10 deadliest accidents
 
@@
 
--
+- Accidents of the New Era mine are now true positives
 
@@
 
+- Apply $k$-means clustering to consolidated data on all 20 features
+- 3 clusters: `low`-risk, `mid`-risk, and `high`-risk
+- Selected cluster centers (omitting 15 features)
+
+```{r k-means}
+if(!exists("clustering")){load("./San Antonio/output/results_markov.RData")}
+sorted.cluster.centers <- clustering$centers[cluster.to.risk.label,c("num.days.lost", "num.days.restrict", "num.death", "num.dis", "viol.quantity")]
+rownames(sorted.cluster.centers) <- risk.labels
+print(sorted.cluster.centers)
+```
+
+- Cluster sizes
+
+```{r k-means sizes}
+if(!exists("clustering")){load("./San Antonio/output/results_markov.RData")}
+sorted.cluster.size <- matrix(data = clustering$size[cluster.to.risk.label], nrow = 1, ncol = 3)
+colnames(sorted.cluster.size) <- risk.labels
+rownames(sorted.cluster.size) <- "size"
+print(sorted.cluster.size)
+```
+
+<!-- https://cran.r-project.org/web/packages/markovchain/vignettes/an_introduction_to_markovchain_package.pdf -->
 ## Markov Chain
+
+- Overall transition matrix
+
+```{r markov-transition}
+if(!exists("markov.mine")){load("./San Antonio/output/results_markov.RData")}
+print(round(markov.mine[], 3))
+```
+
+- Steady-state distribution
+
+```{r markov-steady-state}
+if(!exists("markov.mine")){load("./San Antonio/output/results_markov.RData")}
+print(round(steadyStates(markov.mine), 3))
+```
 
@@
     - Two deadliest mine accidents in the last decade: Upper Big Branch & Sago
-    - MSHA 
+    - Rich MSHA data that need clean-up
+    - Supervised predictive model
+    - Unsupervised clustering of risk
 - Application
@@
 - Future
+    - Improve model performance
     - Expand data: OSHA, EPA, etc.
-    - Models and techniques: artificial neural networks (restricted boltzmann machine), text mining, etc.
+    - Other techniques: artificial neural networks (restricted boltzmann machine), text mining, etc.
 

### 2017-03-22 San Antonio/fixed_effects_luo.R

@@ -1,10 +1,10 @@
 rm(list=ls())
 
 # initialize ----
-library(dplyr)
-library(survival)
-library(caret)
-library(e1071)
+require(dplyr)
+require(survival)
+require(caret)
+require(e1071)
 setwd("~/Git/violation-data-analysis")
 load("./San Antonio/output/result.RData")
 
@@ -126,13 +126,14 @@ fe.div2.result <- fe.div2.result %>% mutate(
     prob > 0.5, TRUE, FALSE
   )
 )
-head(fe.div2.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10)
-head(fe.div2.result %>% filter(year >= 2005) %>% arrange(desc(prob), mine.name), 10)
+# head(fe.div2.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10)
+# head(fe.div2.result %>% filter(year >= 2005) %>% arrange(desc(prob), mine.name), 10)
 
 
 # confusion matrix and other performance metrics
 fe.div2.performance <- confusionMatrix(data = fe.div2.result$pred, reference = fe.div2.result$severe, positive = "TRUE")
-print(fe.div2.performance)
+print(fe.div2.performance$table)
+print(c(fe.div2.performance$overall["Accuracy"], fe.div2.performance$byClass[c("Sensitivity", "Specificity", "Precision", "F1")]))
 
 # output file ----
 save(

### 2017-03-22 San Antonio/markov_chains_luo.R

@@ -2,6 +2,7 @@ rm(list=ls())
 
 # initialize ----
 require(dplyr)
+require(plyr)
 require(markovchain)
 require(reshape)
 setwd("~/Git/violation-data-analysis")
@@ -16,8 +17,8 @@ print(clustering$centers)
 print(clustering$size)
 
 # assign low, mid, high risk labels
-unsorted.risk.label <- c("low", "mid", "high")
-sorted.risk.labels <- sort.int(clustering$centers[,1], index.return = TRUE)$ix
+risk.labels <- c("low", "mid", "high")
+cluster.to.risk.label <- sort.int(clustering$centers[,1], index.return = TRUE)$ix
 
 # Melted format containing quarter-year, mine_id allowing for markov chain analysis ----
 temp$cluster <- clustering$cluster
@@ -36,9 +37,9 @@ markov <- function(x){
     return (list("mine_id" = x[1], 
                  "markov_chain" = cur_markov_chain, 
                  "num_rows" = length(cluster_seq_raw),
-                 "low" = sorted.risk.labels[1] %in% cluster_sequence, 
-                 "mid" = sorted.risk.labels[2] %in% cluster_sequence, 
-                 "high" = sorted.risk.labels[3] %in% cluster_sequence
+                 "low" = cluster.to.risk.label[1] %in% cluster_sequence, 
+                 "mid" = cluster.to.risk.label[2] %in% cluster_sequence, 
+                 "high" = cluster.to.risk.label[3] %in% cluster_sequence
     )
     )
   }
@@ -82,13 +83,15 @@ for(i in 1:3){
 }
 
 print("Transition matrix probability")
-sorted.final.markov <- final_markov[sorted.risk.labels, sorted.risk.labels]
-colnames(sorted.final.markov) <- unsorted.risk.label
-rownames(sorted.final.markov) <- unsorted.risk.label
+sorted.final.markov <- final_markov[cluster.to.risk.label, cluster.to.risk.label]
+colnames(sorted.final.markov) <- risk.labels
+rownames(sorted.final.markov) <- risk.labels
 print(sorted.final.markov)
 
 # create Markov chain object
-markov.mine <- new("markovchain", states = unsorted.risk.label, byrow = TRUE,
+markov.mine <- new("markovchain", states = risk.labels, byrow = TRUE,
                    transitionMatrix = sorted.final.markov, name = "Mine")
 
[…]

### 2017-03-24 San Antonio/slides/presentation_slides.Rmd
semi-final slides, again
final-isa draft

@@
 ---
-title: "A Data-Driven Early Warning System for Mining Accident"
+title: "A Data-Driven Early Warning System for Mining Accidents"
 shorttitle: "Mining Accident Data Analysis"
 author: |
-  | Yu Luo, Ashutosh Nanda, Shiva Rajgopal, Vinay Ramesh,
-  | Venkat Venkatasubramanian, Zhizun Zhang, and Catherine Zhao
+  | Yu Luo, Ashutosh Nanda, Shivaram Rajgopal, Vinay Ramesh,
+  | Zhizun Zhang, Catherine Zhao, and Venkat Venkatasubramanian
 shortauthor: "Luo et al."
@@
 
-\placelogofalse
-
 - Systemic disasters
@@
 - Fast-paced and connected
-- Design complex systems
-- Analyze systemic risk
+- Prevent systemic disasters
+- Analyze systemic risk: go beyond one-off accidents
 
@@
 
-- Formed in	1977
-- Agency of the United States Department of Labor
+- Formed in	1977, agency of the U.S. Department of Labor
 - Mission
@@
 
-- Plot violation trends prior to Upper Big Branch and Sago disasters
+- Plot violation trends prior to disasters
 
@@
 
-## Unsupervised Clustering
+# Conclusion
+
+## Conclusion
+
+<!-- - Causal relationship between accidents and historical data -->
+- Summary
+    - Two deadliest mine accidents in the last decade: Upper Big Branch & Sago
+    - Rich MSHA data that need clean-up
+    - Supervised predictive model
+- Application
+    - "Credit score" for mine safety
+    - Regulators, mines, stakeholders
+- Future
+    - Improve model performance
+    - Unsupervised clustering, neural nets, etc.
+    - Expand data: OSHA, EPA, etc.
+
+## Appendix: Simple Linear Model
+
+- Adjusted $R^2=0.36$
+
+```{r simple lm of days lost}
+if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
+num.days.lm <- lm(num.days.lost~last.quarter.lost+last.year.lost+last.three.years.lost
+           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
+           +last.quarter.viol+last.year.viol+last.three.years.viol
+           +last.quarter.death+last.year.death+last.three.years.death, 
+           data=complete.active.quarters %>% filter(active))
+summary(num.days.lm)$coefficients
+```
+
+## Appendix: Unsupervised Clustering
 
@@
 <!-- https://cran.r-project.org/web/packages/markovchain/vignettes/an_introduction_to_markovchain_package.pdf -->
-## Markov Chain
+## Appendix: Markov Chain
 
@@
 
-# Conclusion
-
-## Conclusion
-
-<!-- - Causal relationship between accidents and historical data -->
-- Summary
-    - Two deadliest mine accidents in the last decade: Upper Big Branch & Sago
-    - Rich MSHA data that need clean-up
-    - Supervised predictive model
-    - Unsupervised clustering of risk
-- Application
-    - "Credit score" for mine safety
-    - Regulators, mines, stakeholders
-- Future
-    - Improve model performance
-    - Expand data: OSHA, EPA, etc.
-    - Other techniques: artificial neural networks (restricted boltzmann machine), text mining, etc.
-
-## Appendix: Simple Linear Model
-
-- Adjusted $R^2=0.36$
-
-```{r simple lm of days lost}
-if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
-num.days.lm <- lm(num.days.lost~last.quarter.lost+last.year.lost+last.three.years.lost
-           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
-           +last.quarter.viol+last.year.viol+last.three.years.viol
-           +last.quarter.death+last.year.death+last.three.years.death, 
-           data=complete.active.quarters %>% filter(active))
-summary(num.days.lm)$coefficients
-```
-
 <!-- USEFUL LINKS:

### 2017-03-29 San Antonio/slides/latex_src/presentation_slides.tex
final presentation and src

% !TEX encoding = UTF-8 Unicode
\documentclass[]{beamer}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\newif\ifbibliography
\hypersetup{
            pdftitle={A Data-Driven Early Warning System for Mining Accidents},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight0.8\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

% Prevent slide breaks in the middle of a paragraph:
\widowpenalties 1 10000
\raggedbottom

\AtBeginPart{
  \let\insertpartnumber\relax
  \let\partname\relax
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \let\insertsectionnumber\relax
    \let\sectionname\relax
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \let\insertsubsectionnumber\relax
  \let\subsectionname\relax
  \frame{\subsectionpage}
}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
\usepackage{cris_style,l16cn}

% http://rmarkdown.rstudio.com/articles_beamer.html
\AtBeginPart{}
\AtBeginSection{}
\AtBeginSubsection{}
\AtBeginSubsubsection{}
\setlength{\emergencystretch}{0em}
\setlength{\parskip}{0pt}

%% http://stackoverflow.com/questions/38323331/code-chunk-font-size-in-beamer-with-knitr-and-latex
%%% change fontsize of R code
%\let\oldShaded\Shaded
%\let\endoldShaded\endShaded
%\renewenvironment{Shaded}{\footnotesize\oldShaded}{\endoldShaded}

%% change fontsize of output
\let\oldverbatim\verbatim
\let\endoldverbatim\endverbatim
\renewenvironment{verbatim}{\footnotesize\oldverbatim}{\endoldverbatim}

\title[Mining Accident Data Analysis]{A Data-Driven Early Warning System for Mining Accidents}
\author[Luo et al.]{Yu Luo, Ashutosh Nanda, Shivaram Rajgopal, Vinay Ramesh,\\
Zhizun Zhang, Catherine Zhao, and Venkat Venkatasubramanian}
\institute[Columbia University]{Chemical Engineering, Computer Science, and Business School\\
Columbia University}
\date{3/27/2017}

\begin{document}
\frame{\titlepage}

\begin{frame}
\tableofcontents[hideallsubsections]
\end{frame}

\begin{frame}{Complex, Resilient, Intelligent Systems (CRIS Lab)}

\begin{figure}[htbp]
\centering
\includegraphics{CrisResearch720.png}
\caption{CRIS Lab}
\end{figure}

\end{frame}

\section{Mine Safety: A Data-Driven
Approach}\label{mine-safety-a-data-driven-approach}

\input{sys_risk.tex}

%\begin{frame}{Systemic Risk}
%\begin{itemize}[<+->]
%\tightlist
%\item
%  Systemic disasters
%
%  \begin{itemize}[<+->]
%  \tightlist
%  \item
%    SARS (2003)
%  \item
%    Northeast Blackout (2003)
%  \item
%    Subprime Crisis (2008)
%  \item
%    Deepwater Horizon Oil Spill (2010)
%  \end{itemize}
%\item
%  Emerging systemic risks
%
%  \begin{itemize}[<+->]
%  \tightlist
%  \item
%    Climate change
%  \item
%    Income/wealth inequality
%  \item
%    Cyber-physical security
%  \item
%    Technological singularity
%  \end{itemize}
%\item
%  Fast-paced and connected
%\item
%  Prevent systemic disasters
%\item
%  Analyze systemic risk: go beyond one-off accidents
%\end{itemize}
%
%\end{frame}

\begin{frame}{Upper Big Branch Mine Disaster (2010)}

\begin{itemize}[<+->]
\tightlist
\item
  April 5, 2010, Raleigh County, West Virginia, owned by Massey Energy
\item
  29 deaths, the worst mining in the United States since 1970
\item
  MSHA cites corporate culture as root cause of Upper Big Branch Mine
  disaster
\end{itemize}

\end{frame}

\begin{frame}{Sago Mine Disaster (2006)}

\begin{itemize}[<+->]
\tightlist
\item
  January 2, 2006, Sago, West Virginia, owned by Anker West Virginia
  Mining
\item
  13 miners were trapped for nearly two days; only one survived
\item
  Fatality number was exceeded by the Upper Big Branch Mine disaster
\item
  MSHA reports prior history of safety violations and fatalities
\end{itemize}

\end{frame}

\begin{frame}{Mine Safety and Health Administration (MSHA)}

\begin{itemize}[<+->]
\tightlist
\item
  Formed in 1977, agency of the U.S. Department of Labor
\item
  Mission

  \begin{itemize}[<+->]
  \tightlist
  \item
    Prevent death, illness, and injury from mining
  \item
    Promote safe and healthful workplaces for U.S. miners
  \item
    Develop and enforce safety and health rules
  \item
    Provide technical, educational, and other types of assistance
  \end{itemize}
\item
  A constantly improving industry in terms of safety
\end{itemize}

\end{frame}

\begin{frame}{Fatality Trend Since 1983}

\begin{center}\includegraphics{presentation_slides_files/figure-beamer/fatality trend-1} \end{center}

\end{frame}

\begin{frame}{Can We Further Improve Mine Safety?}

\begin{itemize}[<+->]
\tightlist
\item
  Process MSHA safety data
\item
  Understand the underlying causal relationships
\item
  Develop early warning systems based on past behaviors
\item
  Credit rating/score analogy

  \begin{itemize}[<+->]
  \tightlist
  \item
    Predict default probability within 18 months
  \item
    Accidents: defaults a month or a year prior to application
  \item
    Violations: missed payments, late payments, etc.
  \end{itemize}
\item
  Can we develop a ``credit score'' for mine safety?
\end{itemize}

\end{frame}

\section{Methods: Data Sources and Model
Preliminaries}\label{methods-data-sources-and-model-preliminaries}

\begin{frame}{Department of Labor Enforcement Data}

\begin{itemize}[<+->]
\tightlist
\item
  Link: \url{https://enforcedata.dol.gov/views/data_catalogs.php}
\item
  Updated daily or weekly
\item
  Publicly available

  \begin{itemize}[<+->]
  \tightlist
  \item
    Department of Labor: MSHA, OSHA, etc.
  \item
    Other departments: EPA, FDA, DOJ, etc.
  \end{itemize}
\end{itemize}

\end{frame}

\begin{frame}{MSHA Data: Sources}

\begin{itemize}[<+->]
\tightlist
\item
  Mine accidents table: ``msha\_accident.csv''

  \begin{itemize}[<+->]
  \tightlist
  \item
    681,386 rows
  \item
    Retrieved 1/26/2017, from
    \url{https://enforcedata.dol.gov/views/data_summary.php}
  \end{itemize}
\item
  MSHA assessed violations table: ``AssessedViolations.csv''

  \begin{itemize}[<+->]
  \tightlist
  \item
    2,169,804 rows
  \item
    Retrieved 12/10/2016, from
    \url{https://arlweb.msha.gov/OpenGovernmentData/OGIMSHA.asp}
  \end{itemize}
\end{itemize}

\end{frame}

\begin{frame}[fragile]{MSHA Data: Advantages}

\begin{itemize}[<+->]
\tightlist
\item
  Each mine has a unique mine ID, e.g., Upper Big Branch (4608436)
\item
  Rich details: e.g., classification, description, and severity
\item
  Selected attributes from the accidents table (omitting 42 attributes)
%\end{itemize}

\begin{verbatim}
##  [1] "mine_id"          "controller_id"    "cal_yr"          
##  [4] "cal_qtr"          "ai_dt"            "inj_degr_desc"   
##  [7] "ai_class_desc"    "ai_occ_desc"      "ai_acty_desc"    
## [10] "exper_tot_calc"   "exper_mine_calc"  "exper_job_calc"  
## [13] "ai_narr"          "accident_type_cd" "no_injuries"     
## [16] "days_restrict"    "days_lost"
\end{verbatim}
\end{itemize}
\end{frame}

\begin{frame}{MSHA Data: Challenges}

\begin{itemize}[<+->]
\tightlist
\item
  Missing data, human errors
\item
  No information about inactive/nonoperating mines
\item
  Most data are not numeric
\item
  Lots of zeros, few severe accidents (\(\sim0.5\%\))
\end{itemize}

\end{frame}

\begin{frame}{Consolidated Data}

\begin{itemize}[<+->]
\tightlist
\item
  Group and summarize accidents/violations by mines
\item
  664,128 rows, 10,377 unique mines
\item
  From 2000 to 2015
\item
  Each row represents data for a unique combination of mine, year, and
  quarter

  \begin{itemize}[<+->]
  \tightlist
  \item
    e.g., Upper Big Branch Mine in the second quarter of 2010
  \end{itemize}
\item
  Each row contains both current and past information

  \begin{itemize}[<+->]
  \tightlist
  \item
    i.e., current quarter, past quarter, past year, and past three years
  \end{itemize}
\end{itemize}

\end{frame}

\begin{frame}[fragile]{Consolidated Data}

\begin{itemize}[<+->]
\tightlist
\item
  All 25 attributes of the consolidated data

\begin{verbatim}
##  [1] "mine_id"                   "mine.name"                
##  [3] "year"                      "quarter"                  
##  [5] "active"                    "num.days.lost"            
##  [7] "last.quarter.lost"         "last.year.lost"           
##  [9] "last.three.years.lost"     "num.days.restrict"        
## [11] "last.quarter.restrict"     "last.year.restrict"       
## [13] "last.three.years.restrict" "num.death"                
## [15] "last.quarter.death"        "last.year.death"          
## [17] "last.three.years.death"    "num.dis"                  
## [19] "last.quarter.dis"          "last.year.dis"            
## [21] "last.three.years.dis"      "viol.quantity"            
## [23] "last.quarter.viol"         "last.year.viol"           
## [25] "last.three.years.viol"
\end{verbatim}
\end{itemize}

\end{frame}

\begin{frame}[fragile]{Top 10 Fatal Accidents Since 2005}

\begin{itemize}[<+->]
\tightlist
\item
  Query the consolidated data on the deadliest accidents

\begin{verbatim}
##                      mine.name mine_id year quarter num.death
## 1  Upper Big Branch Mine-South 4608436 2010       2        29
## 2                    Sago Mine 4608791 2006       1        12
## 3         Crandall Canyon Mine 4201715 2007       3         9
## 4              Darby Mine No 1 1518185 2006       2         5
## 5                  Gibson Mine 1202215 2007       3         3
## 6                Affinity Mine 4608878 2013       1         2
## 7         Aracoma Alma Mine #1 4608801 2006       1         2
## 8       Black Stallion UG Mine 4609086 2014       2         2
## 9                Cucumber Mine 4609066 2007       1         2
## 10             D-14 Stillhouse 1517165 2005       3         2
\end{verbatim}

\item
  Plot violation trends prior to disasters
\end{itemize}

\end{frame}

\begin{frame}{Violation Trend: Upper Big Branch}

\begin{center}\includegraphics{presentation_slides_files/figure-beamer/upper big branch-1} \end{center}

\end{frame}

\begin{frame}{Violation Trend: Sago Mine}

\begin{center}\includegraphics{presentation_slides_files/figure-beamer/sago-1} \end{center}

\end{frame}

\begin{frame}[fragile]{Predictive Model}

\begin{itemize}[<+->]
\tightlist
\item
  Rising violation trends before disasters
\item
  Develop a disaster classifier based on historical data
\item
  Define a \textbf{severe} accident as one with death or permenant
  disability

\begin{verbatim}
## # A tibble: 2 × 3
##   severe      n  perc
##    <lgl>  <int> <dbl>
## 1  FALSE 477077 99.46
## 2   TRUE   2608  0.54
\end{verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Fixed-Mine Effects}

\begin{itemize}[<+->]
\tightlist
\item
  Biostatisticians and epidemiologists call it ``conditional logistic
  regression'' (\texttt{survival::clogit})
\item
  Suitable for \textbf{panel data} (e.g., longitudinal data, our
  consolidated data)
\item
  Model includes mine-specific biases
\item
  Logistic function (for every mine)

\[\Pr(Y=1|\mathbf{X}) = \frac{1}{1+e^{-(\alpha+\boldsymbol\beta\mathbf{X})}}\]

\item
  Logistic function with fixed effects (for the \(i\)-th mine)

\[\Pr(Y=1|\mathbf{X},i) = \frac{1}{1+e^{-(\alpha_i+\boldsymbol\beta\mathbf{x})}}\]
\end{itemize}

\end{frame}

\section{Results and Discussion}\label{results-and-discussion}

\begin{frame}[fragile]{Logistic Regression Without Fixed Effects}

\begin{itemize}[<+->]
\tightlist
\item
  In-sample model
%\end{itemize}

\begin{verbatim}
##           Reference
## Prediction  FALSE   TRUE
##      FALSE 477011   2600
##      TRUE      66      8
\end{verbatim}

\begin{verbatim}
##    Accuracy Sensitivity Specificity   Precision          F1 
##      0.9944      0.0031      0.9999      0.1081      0.0060
\end{verbatim}

%\begin{itemize}[<+->]
\tightlist
\item
  Accuracy = (TP + TN)/(P + N)
\item
  Sensitivity/recall = TP/P
\item
  Specificity = TN/N
\item
  Precision = TP/(TP + FP)
\item
  F1: harmonic mean of sensitivity and precision
\end{itemize}

\end{frame}

\begin{frame}[fragile]{Logistic Regression Without Fixed Effects}

\begin{itemize}[<+->]
\tightlist
\item
  Fails to predict top 10 deadliest disasters

\begin{verbatim}
##                      mine.name year quarter severe  pred
## 1  Upper Big Branch Mine-South 2010       2   TRUE FALSE
## 2                    Sago Mine 2006       1   TRUE FALSE
## 3         Crandall Canyon Mine 2007       3   TRUE FALSE
## 4              Darby Mine No 1 2006       2   TRUE FALSE
## 5                  Gibson Mine 2007       3   TRUE FALSE
## 6                Affinity Mine 2013       1   TRUE FALSE
## 7         Aracoma Alma Mine #1 2006       1   TRUE FALSE
## 8       Black Stallion UG Mine 2014       2   TRUE FALSE
## 9                Cucumber Mine 2007       1   TRUE FALSE
## 10             D-14 Stillhouse 2005       3   TRUE FALSE
\end{verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Logistic Regression Without Fixed Effects}

\begin{itemize}[<+->]
\tightlist
\item
  List of false positive predictions based on predicted probability

\begin{verbatim}
##                                 mine.name year quarter severe pred
## 1  The American Coal Company New Era Mine 2008       3  FALSE TRUE
## 2  The American Coal Company New Era Mine 2008       2  FALSE TRUE
## 3  The American Coal Company New Era Mine 2007       4  FALSE TRUE
## 4  The American Coal Company New Era Mine 2008       4  FALSE TRUE
## 5  The American Coal Company New Era Mine 2008       1  FALSE TRUE
## 6  The American Coal Company New Era Mine 2009       1   TRUE TRUE
## 7  The American Coal Company New Era Mine 2007       3  FALSE TRUE
## 8  The American Coal Company New Era Mine 2006       1  FALSE TRUE
## 9  The American Coal Company New Era Mine 2005       4  FALSE TRUE
## 10 The American Coal Company New Era Mine 2006       2   TRUE TRUE
\end{verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Logistic Regression with Fixed Effects}

\begin{itemize}[<+->]
\tightlist
\item
  Out-of-sample model (randomly select half of the data to train and the
  other half to test)

\begin{verbatim}
##           Reference
## Prediction  FALSE   TRUE
##      FALSE 141332    483
##      TRUE   97167    852
\end{verbatim}

\begin{verbatim}
##    Accuracy Sensitivity Specificity   Precision          F1 
##      0.5928      0.6382      0.5926      0.0087      0.0172
\end{verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Logistic Regression with Fixed Effects}

\begin{itemize}[<+->]
\tightlist
\item
  Successfully predicts top 10 deadliest disasters

\begin{verbatim}
##                        mine.name year quarter severe pred
## 1                      Sago Mine 2006       1   TRUE TRUE
## 2           Crandall Canyon Mine 2007       3   TRUE TRUE
## 3                Darby Mine No 1 2006       2   TRUE TRUE
## 4                  Cucumber Mine 2007       1   TRUE TRUE
## 5                    Dotiki Mine 2010       2   TRUE TRUE
## 6                       Equality 2011       4   TRUE TRUE
## 7                    Meikle Mine 2010       3   TRUE TRUE
## 8            Nanuuq Gold Project 2007       3   TRUE TRUE
## 9  4 J's Gravel Crushing Plant 2 2011       3   TRUE TRUE
## 10                         Adams 2006       3   TRUE TRUE
\end{verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Logistic Regression with Fixed Effects}

\begin{itemize}[<+->]
\tightlist
\item
  Still has a lot of false positive predictions
\item
  List of false positive predictions based on predicted probability

\begin{verbatim}
##                                 mine.name year quarter severe pred
## 1  The American Coal Company New Era Mine 2006       1  FALSE TRUE
## 2             Upper Big Branch Mine-South 2009       3  FALSE TRUE
## 3             Upper Big Branch Mine-South 2009       1  FALSE TRUE
## 4             Upper Big Branch Mine-South 2006       4  FALSE TRUE
## 5             Upper Big Branch Mine-South 2005       1  FALSE TRUE
## 6  The American Coal Company New Era Mine 2005       3  FALSE TRUE
## 7  The American Coal Company New Era Mine 2008       1  FALSE TRUE
## 8  The American Coal Company New Era Mine 2007       4  FALSE TRUE
## 9             Upper Big Branch Mine-South 2006       1  FALSE TRUE
## 10            Upper Big Branch Mine-South 2006       3  FALSE TRUE
\end{verbatim}

%\begin{itemize}[<+->]
%\tightlist
\item
  What happened in the New Era Mine?
\end{itemize}

\end{frame}

\begin{frame}[fragile]{New Era Mine}

\begin{itemize}[<+->]
\tightlist
\item
  Among the worst mines by number of days lost due to accidents
%\end{itemize}

\begin{verbatim}
##                                 mine.name year quarter days.lost
## 1  The American Coal Company New Era Mine 2005       2      2940
## 2  The American Coal Company New Era Mine 2003       2      2914
## 3  The American Coal Company New Era Mine 2005       3      2874
## 4                                 Mathies 2002       1      2840
## 5  The American Coal Company New Era Mine 2004       3      2613
## 6  The American Coal Company New Era Mine 2004       1      2591
## 7                  Monongalia County Mine 2013       3      2563
## 8  The American Coal Company New Era Mine 2005       4      2487
## 9                     Powhatan No. 6 Mine 2013       1      2409
## 10                            Maple Creek 2001       1      2030
\end{verbatim}

%\begin{itemize}[<+->]
%\tightlist
\item
  Rising violation trend from 2000 to 2005
\end{itemize}

\end{frame}

\begin{frame}{New Era Mine}

\begin{center}\includegraphics{presentation_slides_files/figure-beamer/new era viol-1} \end{center}

\end{frame}

\begin{frame}[fragile]{New Labels Including Days Lost}

\begin{itemize}[<+->]
\tightlist
\item
  Updated severe accident label

  \begin{itemize}[<+->]
  \tightlist
  \item
    Previously defined criteria plus days lost \(>\) 300
  \end{itemize}
\item
  Redo out-of-sample model
%\end{itemize}

\begin{verbatim}
##           Reference
## Prediction  FALSE   TRUE
##      FALSE 148496   1267
##      TRUE   88426   1645
\end{verbatim}

\begin{verbatim}
##    Accuracy Sensitivity Specificity   Precision          F1 
##       0.626       0.565       0.627       0.018       0.035
\end{verbatim}

%\begin{itemize}[<+->]
%\tightlist
\item
  Worse true positive rate, improved F1 score
\end{itemize}

\end{frame}

\begin{frame}[fragile]{New Labels Including Days Lost}

\begin{itemize}[<+->]
\tightlist
\item
  Successfully predicts 9 out of top 10 deadliest accidents

\begin{verbatim}
##                        mine.name year quarter severe  pred
## 1                      Sago Mine 2006       1   TRUE  TRUE
## 2           Crandall Canyon Mine 2007       3   TRUE  TRUE
## 3                Darby Mine No 1 2006       2   TRUE  TRUE
## 4                  Cucumber Mine 2007       1   TRUE  TRUE
## 5                    Dotiki Mine 2010       2   TRUE  TRUE
## 6                       Equality 2011       4   TRUE  TRUE
## 7                    Meikle Mine 2010       3   TRUE  TRUE
## 8            Nanuuq Gold Project 2007       3   TRUE  TRUE
## 9  4 J's Gravel Crushing Plant 2 2011       3   TRUE  TRUE
## 10                         Adams 2006       3   TRUE FALSE
\end{verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{New Labels Including Days Lost}

\begin{itemize}[<+->]
\tightlist
\item
  Accidents of the New Era mine are now true positives

\begin{verbatim}
##                                 mine.name year quarter severe pred
## 1  The American Coal Company New Era Mine 2006       1   TRUE TRUE
## 2  The American Coal Company New Era Mine 2005       3   TRUE TRUE
## 3  The American Coal Company New Era Mine 2005       1   TRUE TRUE
## 4                  Monongalia County Mine 2014       3   TRUE TRUE
## 5                     Powhatan No. 6 Mine 2013       3   TRUE TRUE
## 6                     Powhatan No. 6 Mine 2013       4   TRUE TRUE
## 7                    Marshall County Mine 2015       4   TRUE TRUE
## 8  The American Coal Company New Era Mine 2008       1   TRUE TRUE
## 9                      Willow Lake Portal 2008       2   TRUE TRUE
## 10                    Powhatan No. 6 Mine 2013       1   TRUE TRUE
\end{verbatim}
\end{itemize}
\end{frame}

\section{Conclusion}\label{conclusion}

\begin{frame}{Conclusion}

\begin{itemize}[<+->]
\tightlist
\item
  Summary

  \begin{itemize}[<+->]
  \tightlist
  \item
    Two deadliest mine accidents in the last decade: Upper Big Branch \&
    Sago
  \item
    Rich MSHA data that need clean-up
  \item
    Supervised predictive model
  \end{itemize}
\item
  Application

  \begin{itemize}[<+->]
  \tightlist
  \item
    ``Credit score'' for mine safety
  \item
    Regulators, mines, stakeholders
  \end{itemize}
\item
  Future

  \begin{itemize}[<+->]
  \tightlist
  \item
    Improve model performance
  \item
    Unsupervised clustering, neural nets, etc.
  \item
    Expand data: OSHA, EPA, etc.
  \end{itemize}
\end{itemize}

\end{frame}

\begin{frame}[fragile]{Appendix: Simple Linear Model}

\begin{itemize}[<+->]
\tightlist
\item
  Adjusted \(R^2=0.36\)

\begin{verbatim}
##                           Estimate Std. Error t value Pr(>|t|)
## (Intercept)                 0.5243    0.06725     7.8  6.4e-15
## last.quarter.lost           0.0566    0.00179    31.6 2.9e-218
## last.year.lost              0.0724    0.00093    77.8  0.0e+00
## last.three.years.lost       0.0338    0.00032   105.6  0.0e+00
## last.quarter.restrict      -0.0173    0.00461    -3.8  1.7e-04
## last.year.restrict         -0.0123    0.00243    -5.1  3.9e-07
## last.three.years.restrict   0.0072    0.00085     8.4  3.8e-17
## last.quarter.viol           0.3083    0.01095    28.1 3.5e-174
## last.year.viol              0.1352    0.00490    27.6 2.1e-167
## last.three.years.viol      -0.0346    0.00141   -24.7 4.2e-134
## last.quarter.death         -5.7149    1.09783    -5.2  1.9e-07
## last.year.death            -3.6943    0.64330    -5.7  9.3e-09
## last.three.years.death     -0.5155    0.33261    -1.5  1.2e-01
\end{verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Appendix: Unsupervised Clustering}

\begin{itemize}[<+->]
\tightlist
\item
  Apply \(k\)-means clustering to consolidated data on all 20 features
\item
  3 clusters: \texttt{low}-risk, \texttt{mid}-risk, and
  \texttt{high}-risk
\item
  Selected cluster centers (omitting 17 features)

\begin{verbatim}
##      num.days.lost num.days.restrict num.death
## low            5.3               2.1    0.0013
## mid          100.5              18.6    0.0164
## high         508.4              32.7    0.0431
\end{verbatim}

%\begin{itemize}[<+->]
%\tightlist
\item
  Cluster sizes
%\end{itemize}

\begin{verbatim}
##         low   mid high
## size 465203 13299 1183
\end{verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Appendix: Markov Chain}

\begin{itemize}[<+->]
\tightlist
\item
  Overall transition matrix
%\end{itemize}

\begin{verbatim}
##        low   mid  high
## low  0.997 0.003 0.000
## mid  0.087 0.906 0.006
## high 0.000 0.072 0.928
\end{verbatim}

%\begin{itemize}[<+->]
%\tightlist
\item
  Steady-state distribution

\begin{verbatim}
##       low   mid  high
## [1,] 0.97 0.028 0.003
\end{verbatim}
\end{itemize}
\end{frame}

\end{document}

### 2017-03-29 San Antonio/slides/presentation_slides.Rmd
final presentation and src

@@
   beamer_presentation:
-    keep_tex: false
+    keep_tex: true
     incremental: false
@@
     mine.name, year, quarter, num.days.lost
-  )
+  ) %>% rename(days.lost = num.days.lost)
 , 10)
@@
 - 3 clusters: `low`-risk, `mid`-risk, and `high`-risk
-- Selected cluster centers (omitting 15 features)
+- Selected cluster centers (omitting 17 features)
 
@@
 if(!exists("clustering")){load("./San Antonio/output/results_markov.RData")}
-sorted.cluster.centers <- clustering$centers[cluster.to.risk.label,c("num.days.lost", "num.days.restrict", "num.death", "num.dis", "viol.quantity")]
+sorted.cluster.centers <- clustering$centers[cluster.to.risk.label,c("num.days.lost", "num.days.restrict", "num.death")]#, "num.dis", "viol.quantity")]
 rownames(sorted.cluster.centers) <- risk.labels

### 2017-03-29 San Antonio/slides/presentation_slides.tex
final presentation and src

--- 2017-03-29 San Antonio/slides/latex_src/presentation_slides.tex
@@
 % !TEX encoding = UTF-8 Unicode
-\documentclass[]{beamer}
+\documentclass[
+%ignorenonframetext,
+]{beamer}
 \setbeamertemplate{caption}[numbered]
@@
 
-\input{sys_risk.tex}
+\begin{frame}{Systemic Risk}
 
-%\begin{frame}{Systemic Risk}
-%\begin{itemize}[<+->]
-%\tightlist
-%\item
-%  Systemic disasters
-%
-%  \begin{itemize}[<+->]
-%  \tightlist
-%  \item
-%    SARS (2003)
-%  \item
-%    Northeast Blackout (2003)
-%  \item
-%    Subprime Crisis (2008)
-%  \item
-%    Deepwater Horizon Oil Spill (2010)
-%  \end{itemize}
-%\item
-%  Emerging systemic risks
-%
-%  \begin{itemize}[<+->]
-%  \tightlist
-%  \item
-%    Climate change
-%  \item
-%    Income/wealth inequality
-%  \item
-%    Cyber-physical security
-%  \item
-%    Technological singularity
-%  \end{itemize}
-%\item
-%  Fast-paced and connected
-%\item
-%  Prevent systemic disasters
-%\item
-%  Analyze systemic risk: go beyond one-off accidents
-%\end{itemize}
-%
-%\end{frame}
+\begin{itemize}
+\tightlist
+\item
+  Systemic disasters
+
+  \begin{itemize}
+  \tightlist
+  \item
+    SARS (2003)
+  \item
+    Northeast Blackout (2003)
+  \item
+    Subprime Crisis (2008)
+  \item
+    Deepwater Horizon Oil Spill (2010)
+  \end{itemize}
+\item
+  Emerging systemic risks
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Climate change
+  \item
+    Income/wealth inequality
+  \item
+    Cyber-physical security
+  \item
+    Technological singularity
+  \end{itemize}
+\item
+  Fast-paced and connected
+\item
+  Prevent systemic disasters
+\item
+  Analyze systemic risk: go beyond one-off accidents
+\end{itemize}
+
+\end{frame}
 
@@
 
-\begin{itemize}[<+->]
+\begin{itemize}
 \tightlist
@@
 
-  \begin{itemize}[<+->]
+  \begin{itemize}
   \tightlist
@@
   Selected attributes from the accidents table (omitting 42 attributes)
-%\end{itemize}
+\end{itemize}
 
@@
 \end{verbatim}
-\end{itemize}
+
 \end{frame}
@@
   All 25 attributes of the consolidated data
+\end{itemize}
 
@@
 \end{verbatim}
-\end{itemize}
 
@@
   Query the consolidated data on the deadliest accidents
+\end{itemize}
 
@@
 
+\begin{itemize}
+\tightlist
 \item
@@
   disability 
+\end{itemize}
 
@@
   Logistic function (for every mine)
+\end{itemize}
 
@@
 
+\begin{itemize}
+\tightlist
 \item
   Logistic function with fixed effects (for the \(i\)-th mine)
+\end{itemize}
 
 \[\Pr(Y=1|\mathbf{X},i) = \frac{1}{1+e^{-(\alpha_i+\boldsymbol\beta\mathbf{x})}}\]
-\end{itemize}
 
@@
   In-sample model
-%\end{itemize}
+\end{itemize}
 
@@
 
-%\begin{itemize}[<+->]
+\begin{itemize}
 \tightlist
@@
   Fails to predict top 10 deadliest disasters
+\end{itemize}
 
@@
   List of false positive predictions based on predicted probability 
+\end{itemize}
 
@@
   other half to test)
+\end{itemize}
 
@@
   Successfully predicts top 10 deadliest disasters
+\end{itemize}
 
@@
   List of false positive predictions based on predicted probability
+\end{itemize}
 
@@
 
-%\begin{itemize}[<+->]
-%\tightlist
+\begin{itemize}
+\tightlist
 \item
@@
   Among the worst mines by number of days lost due to accidents
-%\end{itemize}
+\end{itemize}
 
@@
   Redo out-of-sample model
-%\end{itemize}
+\end{itemize}
 
@@
   Successfully predicts 9 out of top 10 deadliest accidents
+\end{itemize}
 
@@
   Accidents of the New Era mine are now true positives
+\end{itemize}
 
@@
   Adjusted \(R^2=0.36\)
+\end{itemize}
 
@@
 \item
-  Selected cluster centers (omitting 17 features)
+  Selected cluster centers (omitting 15 features)
+\end{itemize}
 
 \begin{verbatim}
-##      num.days.lost num.days.restrict num.death
-## low            5.3               2.1    0.0013
-## mid          100.5              18.6    0.0164
-## high         508.4              32.7    0.0431
+##      num.days.lost num.days.restrict num.death num.dis viol.quantity
+## low            5.3               2.1    0.0013  0.0029           2.6
+## mid          100.5              18.6    0.0164  0.0313          34.3
+## high         508.4              32.7    0.0431  0.0871          98.9
 \end{verbatim}
 
-%\begin{itemize}[<+->]
-%\tightlist
+\begin{itemize}
+\tightlist
 \item
   Cluster sizes
-%\end{itemize}
+\end{itemize}
 
@@
   Overall transition matrix
-%\end{itemize}
+\end{itemize}
 
@@
 
-%\begin{itemize}[<+->]
-%\tightlist
+\begin{itemize}
+\tightlist
 \item
   Steady-state distribution
+\end{itemize}
 

### 2017-08-07 manuscript/abstract.tex
start writing manuscript

% !TEX root = journal_main.tex

Often, safety condition of a mine deteriorates before an accident happens. For example, before the 2010 Upper Big Branch Mine disaster, one of the largest in the U.S. history, the mine displayed an alarming rising trend of safety violations. Similar building-up behaviors apply to other major mining accidents in the U.S. According to the Mine Safety and Health Administration (MSHA), on average, there are more than a hundred thousand citations and orders issued yearly. The MSHA accident and inspection databases are an untapped resource for safety analysis. Our model analyzes semi-structured data from MSHA, attempting to uncover a company's safety culture from its public regulatory records. This work has implications for developing a data-driven early warning system. In the future, we plan to extend this research to inspection data from the Occupational Safety and Health Administration (OSHA) and the Environmental Protection Agency (EPA).

### 2017-08-07 manuscript/journal_main.tex
start writing manuscript

\documentclass[journal]{IEEEtran}
\usepackage[]{l16cn}
%\renewcommand\hl[1]{#1} % activate to remove highlights
\makenoidxglossaries
\loadglsentries{definitions}

\begin{document}
\title{\huge A Data-Driven Early Warning System for\\Mining Accidents}
\author{
  Yu Luo, Ashutosh Nanda, Shivaram Rajgopal, Vinay Ramesh, Zhizun Zhang, Catherine Zhao, and 
  Venkat Venkatasubramanian$^{\ast}$%
  \thanks{\bf Last updated on~\today{}}
  \thanks{Yu Luo, 
    Columbia University, Department of Chemical Engineering, yl2750@columbia.edu}
  \thanks{Ashutosh Nanda, 
    Columbia University, Department of Computer Science, an2655@columbia.edu}
  \thanks{Shivaram Rajgopal, 
    Columbia Business School, sr3269@columbia.edu}    
  \thanks{Vinay Ramesh, 
    Columbia University, Department of Computer Science, vrr2112@columbia.edu}    
  \thanks{Zhizun Zhang, 
    Columbia University, Department of Chemical Engineering, zz2216@columbia.edu}
  \thanks{Catherine Zhao, 
    Barnard College, Department of Computer Science, caz2114@barnard.edu}    
  \thanks{Venkat Venkatasubramanian$^{\ast}$
    (to whom correspondence should be addressed),
    Columbia University, Department of Chemical Engineering, venkat@columbia.edu}
}

\maketitle

\begin{abstract}
	\input{abstract}
\end{abstract}

\IEEEpeerreviewmaketitle

\input{text}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}

### 2018-05-05 Houston/src/manuscript_preparation.Rmd
Create new branch for the manuscript

---
title: "Manuscript Preparation"
author: "Albert"
date: "4/13/2018"
output: 
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
options(scipen = 1, digits = 2)
require(dplyr)
require(ggplot2)
require(RcppRoll)
require(markovchain)
require(reshape)
require(survival)
require(caret)
require(e1071)
source("./Houston/src/prepare_violation.R")
source("./Houston/src/roll_over.R")

knitr::opts_knit$set(root.dir = "~/documents/projects/violation-data-analysis") # !! CHANGE PATH TO MATCH YOUR OWN
```

## Data Preparation

### Load data

Loading raw data.

```{r load data, include=FALSE}
if(!exists("Accidents") | 
   !exists("AssessedViolations") | 
   !exists("Mines")) {
  load("./Houston/data/Accidents.RData")
  load("./Houston/data/AssessedViolations.RData")
  load("./Houston/data/Mines.RData")
  print("raw RData loaded")
}
```

Let's examine the loaded raw datasets.

```{r data preview, include=FALSE}
# Attributes in Accidents Table
head(Accidents)

# Attributes in Violations Table
head(AssessedViolations)

# Attributes in Mines Table
head(Mines)
```

The `result.RData` from San Antonio folder is the same as the `Consolidated.RData` in Houston folder.

```{r load result, include=FALSE}
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
print("load previous consolidated data")

colnames(complete.active.quarters)
```

### Data Visualization

If we define severe as death plus permenant disability, then severe composition is shown as following.

```{r severe1}
# options(digits = 1)
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
severe.summary <- complete.active.quarters %>%
  filter(active) %>%
  mutate(severe = ifelse(
    num.death + num.dis > 0, TRUE, FALSE
  )) %>%
  group_by(severe) %>%
  summarise(n = n()) %>%
  mutate(perc = n / sum(n) * 100)
print(severe.summary)
```

If we redefine severe as death or permenant diability greater than 0, or more than 300 days lost or days retrict, then the severe composition is following.
```{r severe2}
# options(digits = 1)
if(!exists("complete.active.quarters")){load("./San Antonio/output/result.RData")}
severe.summary <- complete.active.quarters %>%
  filter(active) %>%
  mutate(severe = ifelse(
    (num.death + num.dis > 0) | (num.days.lost + num.days.restrict > 300), TRUE, FALSE
  )) %>%
  group_by(severe) %>%
  summarise(n = n()) %>%
  mutate(perc = n / sum(n) * 100)
print(severe.summary)
```

## Model Development

### Linear Regression

Read these two links first: [How to interpret regression p-values](http://blog.minitab.com/blog/adventures-in-statistics-2/how-to-interpret-regression-analysis-results-p-values-and-coefficients) and
[How to interpret the output of the summary() method for an lm object](http://stats.stackexchange.com/questions/59250/how-to-interpret-the-output-of-the-summary-method-for-an-lm-object-in-r)

We first examine the linearity between "number of days lost" and "last three years' number of violations".

```{r check linearity, include=FALSE}
ggplot(data=complete.active.quarters, mapping = aes(x=last.three.years.viol,y=num.days.lost))+geom_point()
```
It turns out that these variables do not show linear correlation, which violates the linear regression assumption.

Linear regression on num of days lost or restricted:

```{r, include=FALSE}
if(!exists("complete_active_quarters")){load("./Houston/output/Consolidated.RData")}
colnames(complete_active_quarters)

num.days.lm <- lm(0*num.death+1*num.days.lost+1*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost
           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
           +last.quarter.viol+last.year.viol+last.three.years.viol
           +last.quarter.death+last.year.death+last.three.years.death, 
           data=complete.active.quarters %>% filter(active))
summary(num.days.lm)
```
Linear regression on num of deaths:

```{r echo=FALSE}
summary(lm(num.death~last.three.years.lost+last.quarter.viol+last.year.viol, data=complete.active.quarters %>% filter(active)))
```

### Logistic Regression

```{r logistic reg, include=FALSE}
# label data ----
labeled.data.death <- complete.active.quarters %>% filter(active) %>% mutate(
  severe = ifelse(
    num.death + num.dis > 0, TRUE, FALSE
  )
)
labeled.data.lost <- complete.active.quarters %>% filter(active) %>% mutate(
  severe = ifelse(
    num.death + num.dis > 0 | num.days.lost > 300, TRUE, FALSE
  )
)

# logistic regression without fixed effects ----
# regular logistic regression model
glm.model <- glm(severe ~ 
                   last.quarter.lost+last.year.lost+last.three.years.lost
                 +last.quarter.restrict+last.year.restrict+last.three.years.restrict
                 +last.quarter.viol+last.year.viol+last.three.years.viol
                 +last.quarter.death+last.year.death+last.three.years.death,
                 data = labeled.data.death, family = binomial(link='logit'))
summary(glm.model)

glm.prediction <- predict(object = glm.model, type = "response")
glm.result <- labeled.data.death %>% select(mine_id, mine.name, year, quarter, num.death, num.dis, severe)
glm.result$prob <- glm.prediction
glm.result <- glm.result %>% mutate(
  pred = ifelse(
    prob > 0.5, TRUE, FALSE
  )
)
head(glm.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10)
head(glm.result %>% filter(year >= 2005) %>% arrange(desc(prob), mine.name), 10)

glm.performance <- confusionMatrix(data = glm.result$pred, reference = glm.result$severe, positive = "TRUE")
print(glm.performance)	
```

### Conditional Logistic Regression

```{r fixed effect, include=FALSE}
# fixed effects on training and testing data ----
# train and test
set.seed(1)
data.len <- nrow(labeled.data.death)
train.indices <- sample(seq_len(nrow(labeled.data.death)), size = floor(data.len / 2))

labeled.data.death.train <- labeled.data.death[train.indices, ]
labeled.data.death.result <- labeled.data.death[-train.indices, ]
labeled.data.death.result <- labeled.data.death.result %>% filter(is.element(mine_id, unique(labeled.data.death.train$mine_id)))

fe.div.model <- clogit(severe ~ last.quarter.lost+last.year.lost+last.three.years.lost
                   +last.quarter.restrict+last.year.restrict+last.three.years.restrict
                   +last.quarter.viol+last.year.viol+last.three.years.viol
                   +last.quarter.death+last.year.death+last.three.years.death
                   + strata(mine_id), data = labeled.data.death.train)
summary(fe.div.model)

fe.div.prediction <- predict(object = fe.div.model, newdata = labeled.data.death.result,type = "lp") %>% sigmoid()
fe.div.result <- labeled.data.death.result %>% select(mine_id, mine.name, year, quarter, num.death, num.dis, severe)
fe.div.result$prob <- fe.div.prediction
fe.div.result <- fe.div.result %>% mutate(
  pred = ifelse(
    prob > 0.5, TRUE, FALSE
  )
)
head(fe.div.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10)
head(fe.div.result %>% filter(year >= 2005) %>% arrange(desc(prob), mine.name), 10)

# confusion matrix and other performance metrics
#fe.div.performance <- confusionMatrix(data = fe.div.result$pred, reference = fe.div.result$severe, positive = "TRUE")
fe.div.performance <- table(fe.div.result$pred, fe.div.result$severe)
print(fe.div.performance)

# fixed effects on training and testing data (using labeled.data.lost) ----
# fe.div + num.days.lost: train and test
labeled.data.lost.train <- labeled.data.lost[train.indices, ]
labeled.data.lost.result <- labeled.data.lost[-train.indices, ]
labeled.data.lost.result <- labeled.data.lost.result %>% filter(is.element(mine_id, unique(labeled.data.lost.train$mine_id)))

fe.div2.model <- clogit(severe ~ last.quarter.lost+last.year.lost+last.three.years.lost
                       +last.quarter.restrict+last.year.restrict+last.three.years.restrict
                       +last.quarter.viol+last.year.viol+last.three.years.viol
                       +last.quarter.death+last.year.death+last.three.years.death
                       + strata(mine_id), data = labeled.data.lost.train)
summary(fe.div2.model)

fe.div2.prediction <- predict(object = fe.div2.model, newdata = labeled.data.lost.result,type = "lp") %>% sigmoid()
fe.div2.result <- labeled.data.lost.result %>% select(mine_id, mine.name, year, quarter, num.death, num.dis, severe)
fe.div2.result$prob <- fe.div2.prediction
fe.div2.result <- fe.div2.result %>% mutate(
  pred = ifelse(
    prob > 0.5, TRUE, FALSE
  )
)
# head(fe.div2.result %>% filter(year >= 2005) %>% arrange(desc(num.death), mine.name), 10)
# head(fe.div2.result %>% filter(year >= 2005) %>% arrange(desc(prob), mine.name), 10)

# confusion matrix and other performance metrics
fe.div2.performance <- confusionMatrix(data = fe.div2.result$pred, reference = fe.div2.result$severe, positive = "TRUE")
print(fe.div2.performance$table)
print(c(fe.div2.performance$overall["Accuracy"], fe.div2.performance$byClass[c("Sensitivity", "Specificity", "Precision", "F1")]))   
```

### Clustering

First, we cluster the data using K-means.

```{r clustering, include=FALSE}
# clustering ----
temp <- complete.active.quarters %>% filter(active==TRUE)
set.seed(20)
clustering <- kmeans(temp[, 6:ncol(temp)], 3, nstart=20)
freq_table <- table(clustering$cluster)

print(clustering$centers)
print(clustering$size)
```

Then, let's visualize the clusters.

```{r cluster visualization, include=FALSE}
clustering$cluster <- as.factor(clustering$cluster)
ggplot(temp, aes(num.death, viol.quantity, color = clustering$cluster)) + geom_point()
```

### Markov Chain

```{r markov, include=FALSE}
# assign low, mid, high risk labels
risk.labels <- c("low", "mid", "high")
cluster.to.risk.label <- sort.int(clustering$centers[,1], index.return = TRUE)$ix

# Melted format containing quarter-year, mine_id allowing for markov chain analysis ----
temp$cluster <- clustering$cluster
temp$date <- paste(as.character(temp$year), "-" , as.character(temp$quarter))
temp2 <- temp %>% select(mine_id, cluster, date)
melted <- melt(temp2, id=c("mine_id", "date"))
aggregate_cluster_seq <- cast(melted, mine_id~date)

# Markov Chain function ----
markov <- function(x){
  x <- as.integer(x)
  cluster_seq_raw <- x[!is.na(x)]
  if(length(cluster_seq_raw) > 1){
    cluster_sequence <- cluster_seq_raw[2:length(cluster_seq_raw)]
    cur_markov_chain <- createSequenceMatrix(cluster_sequence, sanitize = FALSE)
    return (list("mine_id" = x[1], 
                 "markov_chain" = cur_markov_chain, 
                 "num_rows" = length(cluster_seq_raw),
                 "low" = cluster.to.risk.label[1] %in% cluster_sequence, 
                 "mid" = cluster.to.risk.label[2] %in% cluster_sequence, 
                 "high" = cluster.to.risk.label[3] %in% cluster_sequence
    )
    )
  }
  return (list("mine_id" = x[1], 
               "markov_chain" = NA, 
               "num_rows" = 0,
               "low" = FALSE, "mid" = FALSE, "high" = FALSE
  )
  )
}

# create mine-date transition matrices ----
result<-apply(aggregate_cluster_seq, 1, markov)

# Create Overall transition matrix ----
markov_chains_ind <- sapply(result, "[[", 2)

#adding matrix of different dimentions, source: http://stackoverflow.com/questions/13571359/join-and-sum-not-compatible-matrices
add_matrices_1 <- function(...) {
  a <- list(...)
  cols <- sort(unique(unlist(lapply(a, colnames))))
  rows <- sort(unique(unlist(lapply(a, rownames))))
  out <- array(0, dim=c(length(rows), length(cols)), dimnames=list(rows,cols))
  for(M in a) { out[rownames(M), colnames(M)] <- out[rownames(M), colnames(M)] + M }
  return(out)
}

final_transition<- matrix(rep(0,times= 9), ncol = 3, nrow = 3)
colnames(final_transition) <- c(1,2,3)
rownames(final_transition) <- c(1,2,3)

for(i in 1:length(markov_chains_ind)){
  final_transition <- add_matrices_1(final_transition,markov_chains_ind[[i]])
}
print("Tranition matrix frequency")
final_transition

final_markov<- matrix(rep(0,times= 9), ncol = 3, nrow = 3)
for(i in 1:3){
  final_markov[i,] <- final_transition[i,]/sum(final_transition[i,])  
}

print("Transition matrix probability")
sorted.final.markov <- final_markov[cluster.to.risk.label, cluster.to.risk.label]
colnames(sorted.final.markov) <- risk.labels
rownames(sorted.final.markov) <- risk.labels
print(sorted.final.markov)
```

### SVM

We try to classify risky mines using SVM.

```{r svm, include=FALSE}
# check train and test data
head(labeled.data.lost.train)
head(labeled.data.lost.result)

# convert label from `bool` to `int`
labeled.data.lost.train <- labeled.data.lost.train %>% mutate(severe = severe*1)
labeled.data.lost.result <- labeled.data.lost.result %>% mutate(severe = severe*1)

svm.model <- svm(severe ~ 
                   last.quarter.lost+last.year.lost+last.three.years.lost
                 +last.quarter.restrict+last.year.restrict+last.three.years.restrict
                 +last.quarter.viol+last.year.viol+last.three.years.viol
                 +last.quarter.death+last.year.death+last.three.years.death,
                 data=labeled.data.lost.train, method="C-classification", kernel="linear")
summary(svm.model)

pred.train <- predit(svm.model, labeled.data.lost.train)
mean(pred.train==labeled.data.lost.train$severe)
```

## Evaluation

### 2018-05-24 Houston/src/manuscript_preparation.Rmd
asiv21: restructured manuscript_preparation.Rmd

@@
 require(e1071)
-source("./Houston/src/prepare_violation.R")
-source("./Houston/src/roll_over.R")
+knitr::opts_knit$set(root.dir = "~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/") # !! CHANGE PATH TO MATCH YOUR OWN
 
-knitr::opts_knit$set(root.dir = "~/documents/projects/violation-data-analysis") # !! CHANGE PATH TO MATCH YOUR OWN
+# source("~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/Houston/src/prepare_violation.R")
+# source("~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/Houston/src/roll_over.R")
+
+# knitr::opts_knit$set(root.dir = "~/documents/projects/violation-data-analysis") # !! CHANGE PATH TO MATCH YOUR OWN
+
 ```
@@
 ```{r load data, include=FALSE}
+source("./Houston/src/prepare_violation.R")
+source("./Houston/src/roll_over.R")
 if(!exists("Accidents") | 
@@
 
-glm.performance <- confusionMatrix(data = glm.result$pred, reference = glm.result$severe, positive = "TRUE")
+glm.performance <- confusionMatrix(data = factor(glm.result$pred), reference = factor(glm.result$severe), positive = "TRUE")
 print(glm.performance)	

### 2018-06-06 Houston/src/manuscript_preparation.Rmd
SVM kernel radial basis

@@
 require(RcppRoll)
-require(markovchain)
+require(markovchain) # FIX: markovchain seems not compatible with R3.5.0
 require(reshape)
@@
 require(e1071)
-knitr::opts_knit$set(root.dir = "~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/") # !! CHANGE PATH TO MATCH YOUR OWN
-
+# knitr::opts_knit$set(root.dir = "~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/") # !! CHANGE PATH TO MATCH YOUR OWN
 # source("~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/Houston/src/prepare_violation.R")
 # source("~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/Houston/src/roll_over.R")
-
-# knitr::opts_knit$set(root.dir = "~/documents/projects/violation-data-analysis") # !! CHANGE PATH TO MATCH YOUR OWN
+#knitr::opts_knit$set(root.dir = "~/documents/projects/violation-data-analysis") # !! CHANGE PATH TO MATCH YOUR OWN
 
@@
                  +last.quarter.death+last.year.death+last.three.years.death,
-                 data=labeled.data.lost.train, method="C-classification", kernel="linear")
+                 data=labeled.data.lost.train, method="C-classification", kernel="radial")
 summary(svm.model)
 
-pred.train <- predit(svm.model, labeled.data.lost.train)
-mean(pred.train==labeled.data.lost.train$severe)
+# save the computationally costly svm model
+#save(svm.model, file="./Houston/data/svm_radial.RData")
+
+plot(svm.model, labeled.data.lost.train)
+
+compareTable <- table (labeled.data.lost.train$severe, predict(svm.model))  # tabulate
+mean(labeled.data.lost.train$severe != predict(svm)) # false rate
+
+pred.train <- predict(svm.model, labeled.data.lost.train)
+mean(pred.train==labeled.data.lost.train$severe) # true rate
 ```

### 2018-06-18 Houston/src/manuscript_preparation.Rmd
Random forest and tuning

@@
 require(e1071)
+require(randomForest)
 # knitr::opts_knit$set(root.dir = "~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/") # !! CHANGE PATH TO MATCH YOUR OWN
@@
 ```
-Linear regression on num of deaths:
+Linear Regression on num of deaths:
 
@@
 
+#### K-means
+
 First, we cluster the data using K-means.
@@
 ```
+
+#### Spectral Clustering
+
+Next, let's try out spectral clustering.  (TODO)
 
@@
 #save(svm.model, file="./Houston/data/svm_radial.RData")
+load("./Houston/output/svm_radial.RData")
 
-plot(svm.model, labeled.data.lost.train)
+tune_out <- tune.svm(severe~.,
+                     data=labeled.data.lost.train,
+                     gamma=10^(-3:3),
+                     cost=c(0.01,0.1,1,10,100,1000) ) # e1071 tune functions have internal iteration limits, in this case, it is better to use another package
 
 compareTable <- table (labeled.data.lost.train$severe, predict(svm.model))  # tabulate
-mean(labeled.data.lost.train$severe != predict(svm)) # false rate
+mean(labeled.data.lost.train$severe != predict(svm)) # misclassification rate
 
 pred.train <- predict(svm.model, labeled.data.lost.train)
-mean(pred.train==labeled.data.lost.train$severe) # true rate
+mean(pred.train==labeled.data.lost.train$severe) # classification rate
 ```
 
-## Evaluation
+It turns out SVM is not a good option for classifying highly skewed data.
+`tune.svm` does not work for this dataset due to its large size.
+Current SVM model with RBF kernel gives 100% misclassification rate.
+
+### Random Forest
+
+Random forest is self-bagging algorithm, which never overfits no matter what number of trees given. It is fast and performs well in most cases. We try to use random forests to classify the data.
+
+```{r randomForest, include=FALSE}
+# reformat labeled.data.lost.train$severe to factors so that `randomForest`
+# can do a classification
+labeled.data.lost.train$severe <- labeled.data.lost.train$severe %>% factor
+labeled.data.lost.result$severe <- labeled.data.lost.result$severe %>% factor
+
+# Note: nrow of `labled.data.lost.train` and `labeled.data.lost.result` are not equal
+# so, we need to modify the longer one of them in order to run `predict` function.
+# Let's do it on `labeled.data.lost.train`.
+labeled.data.lost.train.trimmed <- labeled.data.lost.train[9:nrow(labeled.data.lost.train), ]
+
+ptm <- proc.time() # start clock to time the process
+rf.model <- randomForest(severe ~ last.quarter.lost+last.year.lost+last.three.years.lost
+                 +last.quarter.restrict+last.year.restrict+last.three.years.restrict
+                 +last.quarter.viol+last.year.viol+last.three.years.viol
+                 +last.quarter.death+last.year.death+last.three.years.death,
+                 data = labeled.data.lost.train.trimmed,
+                 strata = mind_id,
+                 nodesize = 240, # tuning parameter; does not affect much; may improve speed a little
+                 ntree=200, # tuning parameter; no effect
+                 mtry = 2 # tuning parameter; improves OOB error a little; see `rf.tune.mtry`
+                 )
+proc.time() - ptm # stop the clock
+
+#save(rf.model, file="rf_model.RData") # base model, without any tuning
+#save(rf.model, file="rf_model_tuned.RData") # tuned model
+
+load("./Houston/output/rf_model.RData")
+load("./Houston/output/rf_model_tuned.RData")
+
+rf.tune.mtry <- tuneRF(labeled.data.lost.train[, c('last.quarter.lost','last.year.lost',
+                                                'last.three.years.lost','last.quarter.restrict',
+                                                'last.year.restrict', 'last.three.years.restrict',
+                                                'last.quarter.viol','last.year.viol',
+                                                'last.three.years.viol', 'last.quarter.death',
+                                                'last.year.death', 'last.year.death',
+                                                'last.three.years.death')],
+                       labeled.data.lost.train[, 'severe']
+                       )
+
+#save(rf.model, file="rf_tune_mtry.RData") # mtry tuning results
+
+pred <- predict(rf.model, data=labeled.data.lost.result)
+freq <- table(pred, labeled.data.lost.result$severe)
+sum(diag(freq))/sum(freq) # report accuracy
+
+plot(rf.model) # plot the error rate vs number of trees
+
+```
+
+It turns out that random forest can classify the problem in a reasonable time, however, the false predication of true events is 0.88 after tuning.
+The OOB error is 1.1%.
+tuning does not improve the performance significantly.
+
+Now, let's try an ensemble learning method, which is a combination of random forest and SVM.
+
+### Ensemble Learning

### 2018-06-20 Houston/src/manuscript_preparation.Rmd
Add ensemble learning

@@
 require(randomForest)
+require(kernlab)
+require(SuperLearner)
 # knitr::opts_knit$set(root.dir = "~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/") # !! CHANGE PATH TO MATCH YOUR OWN
@@
 
+#save(clustering, file="kmeans_clustering.RData")
+
+load("./Houston/output/kmeans_clustering.RData")
+
 print(clustering$centers)
@@
 
-Next, let's try out spectral clustering.  (TODO)
+Next, let's try out spectral clustering, which is based on the connectivity between data points.
+
+```{r sprectral clustering, include=FALSE}
+temp <- complete.active.quarters %>% filter(active==TRUE)
+set.seed(20)
+
+sc.model <- specc(temp[, 6:ncol(temp)], centers=3)
+#save(sc.model, file="sc_model.RData")
+
+load("./Houston/output/sc_model.RData")
+
+#plot(temp[, 6:ncol(temp)], col=sc.model, pch=4) # too many features, very slow
+plot(temp[, 6], col=sc.model, pch=4) # plot num.days.lost cluster
+
+```
+
+The dataset is highly skewed, so K-means does not perform well as expected since it uses means to estimate centers.
+On the other hand, spectral clustering uses affinity between data points to cluster.
+However, the performance is poor as well.
 
@@
 labeled.data.lost.result$severe <- labeled.data.lost.result$severe %>% factor
+set.seed(20)
 
@@
 
-Now, let's try an ensemble learning method, which is a combination of random forest and SVM.
+### Ensemble Learning
 
-### Ensemble Learning
+Now, let's try an ensemble learning method, which is a combination of single learning algorithms.
+In this case, let's try svm and random forest and see if the ensemble learning could improve the performance.
+
+Read [this post](https://www.datacamp.com/community/tutorials/ensemble-r-machine-learning) first.
+
+```{r ensemble, include=FALSE}
+# encode response variable for superlearner
+y <- as.numeric(labeled.data.lost.train$severe)-1
+ytest <- as.numeric(labeled.data.lost.result$severe)-1
+
+# separate predictors
+x <- data.frame(labeled.data.lost.train[, c('last.quarter.lost','last.year.lost',
+                                            'last.three.years.lost','last.quarter.restrict',
+                                            'last.year.restrict', 'last.three.years.restrict',
+                                            'last.quarter.viol','last.year.viol',
+                                            'last.three.years.viol', 'last.quarter.death',
+                                            'last.year.death', 'last.year.death',
+                                            'last.three.years.death')])
+xtest <- data.frame(labeled.data.lost.result[, c('last.quarter.lost','last.year.lost',
+                                            'last.three.years.lost','last.quarter.restrict',
+                                            'last.year.restrict', 'last.three.years.restrict',
+                                            'last.quarter.viol','last.year.viol',
+                                            'last.three.years.viol', 'last.quarter.death',
+                                            'last.year.death', 'last.year.death',
+                                            'last.three.years.death')])
+
+# TODO: 1. try default params first
+#       2. then try tuned params with ksvm and randomForest
+#       3. try some other learners such as glmnet
+ensemble.model <- SuperLearner(y,
+                               x,
+                               family=binomial(),
+                               SL.library=list("SL.randomForest", # TODO: use ranger, SL.randomForest will be removed from SuperLearner
+                                               "SL.ksvm") 
+                               )
+
+SL.randomForest.tune <- function(...){
+      SL.randomForest(..., num.trees=200, mtry=2, nodeside=240)
+    }
+
+ensemble.model.tune <- SuperLearner(y,
+                                    x,
+                                    family=binomial(),
+                                    SL.library=list("SL.ranger", # TODO: do `install.packages("ranger")` first if using `SL.ranger`
+                                                    "SL.ksvm",
+                                                    "SL.randomForest.tune") 
+                                    )
+
+ensemble.model.more <- SuperLearner(y,
+                                    x,
+                                    family=binomial(),
+                                    SL.library=list("SL.ranger", # TODO: need pacakge `ranger`, please do `install.packages("ranger")`
+                                                    "SL.ksvm",
+                                                    "SL.randomForest.tune",
+                                                    "SL.glmnet") # TDOD: need package `glmnet`
+                                    )
+
+```

### 2018-06-29 Houston/src/manuscript_preparation.Rmd
Minor change to manuscript_prep.Rmd

@@
 compareTable <- table (labeled.data.lost.train$severe, predict(svm.model))  # tabulate
-mean(labeled.data.lost.train$severe != predict(svm)) # misclassification rate
+mean(labeled.data.lost.train$severe != predict(svm.model)) # misclassification rate
 

### 2018-07-27 Houston/src/manuscript_preparation.Rmd
Modifications to manuscript_preparation.Rmd

@@
 require(SuperLearner)
+
+## Abhishek Macbook
 # knitr::opts_knit$set(root.dir = "~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/") # !! CHANGE PATH TO MATCH YOUR OWN
@@
 # source("~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/Houston/src/roll_over.R")
-#knitr::opts_knit$set(root.dir = "~/documents/projects/violation-data-analysis") # !! CHANGE PATH TO MATCH YOUR OWN
+
+##Abhishek Desktop
+knitr::opts_knit$set(root.dir = "D:/Github/CRIS/violation-data-analysis/") # !! CHANGE PATH TO MATCH YOUR OWN
+source("D:/Github/CRIS/violation-data-analysis/Houston/src/prepare_violation.R")
+source("D:/Github/CRIS/violation-data-analysis/Houston/src/roll_over.R")
+
+##Albert Macbook
+# knitr::opts_knit$set(root.dir = "~/documents/projects/violation-data-analysis") # !! CHANGE PATH TO MATCH YOUR OWN
 
@@
 # confusion matrix and other performance metrics
-fe.div2.performance <- confusionMatrix(data = fe.div2.result$pred, reference = fe.div2.result$severe, positive = "TRUE")
+fe.div2.performance <- confusionMatrix(data = factor(fe.div2.result$pred), reference = factor(fe.div2.result$severe), positive = "TRUE")
 print(fe.div2.performance$table)
@@
 
-pred.train <- predict(svm.model, labeled.data.lost.train)
+pred.train <- predict(svm.model, labeled.data.lost.train) # Returns value from 0 to 1. Needs to be thresholded to give 0 or 1
 mean(pred.train==labeled.data.lost.train$severe) # classification rate
@@
                                             'last.three.years.lost','last.quarter.restrict',
+
                                             'last.year.restrict', 'last.three.years.restrict',

### 2018-07-28 Houston/src/conditional_logistic.R
changed train test data split to keep single mine data together

@@ -29,10 +29,17 @@ labeled_data_death_dis_only <- complete_active_quarters %>% filter(ACTIVE) %>% m
 labeled_all_data <- labeled_data_death_dis_only
 set.seed(1)
 data_len <- nrow(labeled_data_death_dis_only)
-train_indices <- sample(seq_len(nrow(labeled_data_death_dis_only)), size = floor(data_len * 0.5))
-labeled_train <- labeled_data_death_dis_only[train_indices, ]
-labeled_test <- labeled_data_death_dis_only[-train_indices, ]
-labeled_test <- labeled_test %>% filter(is.element(MINE_ID, unique(labeled_train$MINE_ID)))
+
+groups <-
+	labeled_all_data %>%
+	select(MINE_ID) %>%
+	distinct(MINE_ID) %>%
+	rowwise() %>%
+	mutate(group=sample(c("train","test"),1,replace=TRUE,prob=c(.5,.5)))
+
+labeled_all_data <- labeled_all_data %>% left_join(groups)
+labeled_test <- filter(labeled_all_data, group == "test")
+labeled_train <- filter(labeled_all_data, group == "train")
 
 # fixed effects on combined data ----
 # fixed effects model
@@ -41,7 +48,7 @@ in_sample_model <- clogit(SEVERE ~
                          LAST_QUARTER_DAYS_RESTRICT + LAST_YEAR_DAYS_RESTRICT + LAST_THREE_YEARS_DAYS_RESTRICT +
                          LAST_QUARTER_DEATH + LAST_YEAR_DEATH + LAST_THREE_YEARS_DEATH +
                          LAST_QUARTER_DIS + LAST_YEAR_DIS + LAST_THREE_YEARS_DIS +
-                         LAST_QUARTER_VIOLATION + LAST_YEAR_VIOLATION + LAST_THREE_YEARS_VIOLATION +
+                         LAST_YEAR_VIOLATION + LAST_THREE_YEARS_VIOLATION +
                          LAST_QUARTER_PENALTY + LAST_YEAR_PENALTY + LAST_THREE_YEARS_PENALTY
                          # + strata(MINE_ID), # use this, we get 82% (TP) + 56% (TN)
                          + strata(CURRENT_MINE_TYPE,COAL_METAL_IND), # use this, we get 58% (TP) + 77% (TN)
@@ -77,7 +84,7 @@ out_sample_model <- clogit(SEVERE ~
[…]

### 2018-07-28 Houston/src/consolidate_data.R
organized output columns

@@ -155,8 +155,12 @@ temp <- merge(temp, mines, by = "MINE_ID", all.x=TRUE)
 #rearranging columns
 temp <- temp[, c(1, 29:ncol(temp), 2:3, 28, 4:27)]
 
+
+q_order <- temp[order(temp[,1], temp[,6], temp[,5]),]
+
+
 # result
-complete_active_quarters <- temp
+complete_active_quarters <- q_order
 save(complete_active_quarters, file="./Houston/output/Consolidated.RData")
 
 # see simple_lm.R for simple lm

### 2018-07-28 Houston/output/resetCorrection.py
added new code for reset correction and false positive analysis

@@ -1,0 +1,106 @@
+import rpy2
+from rpy2.robjects import r, pandas2ri
+import pandas as pd
+from sklearn.metrics import confusion_matrix
+from sklearn.metrics import accuracy_score
+import math
+import numpy as np
+from operator import mul
+
+def exp(s):
+    e = math.e
+    b = (1-e**-11)/(e-e*s)
+    a = s + 1/(e*b)
+    return a,b
+
+def reset_correction(file_path):
+    pandas2ri.activate()
+    rf=r['load'](file_path)
+    
+    df = pd.DataFrame(data=r['in_sample_result'])
+    df_bad = df.loc[df['SEVERE'] == 1]
+    bad_mines = set(df_bad['MINE_ID'])
+    all_mines = set(df['MINE_ID'])
+    good_mines = all_mines - bad_mines
+    
+    dfs = []
+    for mine in set(df_bad['MINE_ID']):
+        df_mine = df.loc[df['MINE_ID'] == mine]
+        
+        real = list(df_mine['SEVERE'])
+        pred = list(df_mine['PREDICTION'])
+        prob = list(df_mine['PROBABILITY'])
+        
+        counts = [real[0]]
+        for i in range(1,len(pred)):
+            c = real[i] + counts[i-1]
+            counts.append(c)
+
+        time = [0]
+        begin = 1
+        for i in range(1,len(counts)):
+            if counts[i] == 0:
+                time.append(0)
+            else:
+                if counts[i-1] < counts[i]:
+                    time.append(0)
+                    begin = 1
+                else:
+                    if time[i-1] + 1 > 12:
+                        time.append(0)
+                        begin = 0
+                    else:
+                        if begin == 1:
+                            time.append(time[i-1]+1)
+                        else:
+                            time.append(0)
+
+        e = math.e
+        adj = [1]
+        i = 1
+        while i < len(counts):
+            if counts[i] == 0:
+                adj.append(1)
+            else:
+                if time[i] == 0:
+                    if counts[i] > counts[i-1] and counts[i] > 1:
+                        if adj[i-1] == 1:
+                            adj.append(1)
+                        else:
+                            adj.append(a - e**(-1*(time[i-1]+1 + counts[i-1] -2))/b)        
+                    else: 
+                        adj.append(1)
+
[…]

### 2018-07-28 Houston/output/resetCorrection.py
added comments and main function

@@ -7,6 +7,7 @@ import math
 import numpy as np
 from operator import mul
 
+#exponential discount factor
 def exp(s):
     e = math.e
     b = (1-e**-11)/(e-e*s)
@@ -36,6 +37,7 @@ def reset_correction(file_path):
             c = real[i] + counts[i-1]
             counts.append(c)
 
+        #calculate time since accident
         time = [0]
         begin = 1
         for i in range(1,len(counts)):
@@ -55,6 +57,7 @@ def reset_correction(file_path):
                         else:
                             time.append(0)
 
+        #calculate discount factor. becomes less generous with more accidents, and decays over 3 years
         e = math.e
         adj = [1]
         i = 1
@@ -103,4 +106,13 @@ def reset_correction(file_path):
                      'NEW_PROB',
                      'NEW_PRED']]
     df_out = df_tmp.rename(index=str, columns={'NEW_PRED':'PREDICTION','NEW_PROB':'PROBABILITY'})
+    
+    #output new prediction results
     df_out.to_csv('outputResults.csv')
+    
+
+def main():
+    reset_correction('./Result_clogit.RData')
+    
+    
+main()

### 2018-07-28 Houston/src/conditional_logistic.R
Reverted EJ's modifications on .R and .RData

@@ -29,17 +29,10 @@ labeled_data_death_dis_only <- complete_active_quarters %>% filter(ACTIVE) %>% m
 labeled_all_data <- labeled_data_death_dis_only
 set.seed(1)
 data_len <- nrow(labeled_data_death_dis_only)
-
-groups <-
-	labeled_all_data %>%
-	select(MINE_ID) %>%
-	distinct(MINE_ID) %>%
-	rowwise() %>%
-	mutate(group=sample(c("train","test"),1,replace=TRUE,prob=c(.5,.5)))
-
-labeled_all_data <- labeled_all_data %>% left_join(groups)
-labeled_test <- filter(labeled_all_data, group == "test")
-labeled_train <- filter(labeled_all_data, group == "train")
+train_indices <- sample(seq_len(nrow(labeled_data_death_dis_only)), size = floor(data_len * 0.5))
+labeled_train <- labeled_data_death_dis_only[train_indices, ]
+labeled_test <- labeled_data_death_dis_only[-train_indices, ]
+labeled_test <- labeled_test %>% filter(is.element(MINE_ID, unique(labeled_train$MINE_ID)))
 
 # fixed effects on combined data ----
 # fixed effects model
@@ -48,7 +41,7 @@ in_sample_model <- clogit(SEVERE ~
                          LAST_QUARTER_DAYS_RESTRICT + LAST_YEAR_DAYS_RESTRICT + LAST_THREE_YEARS_DAYS_RESTRICT +
                          LAST_QUARTER_DEATH + LAST_YEAR_DEATH + LAST_THREE_YEARS_DEATH +
                          LAST_QUARTER_DIS + LAST_YEAR_DIS + LAST_THREE_YEARS_DIS +
-                         LAST_YEAR_VIOLATION + LAST_THREE_YEARS_VIOLATION +
+                         LAST_QUARTER_VIOLATION + LAST_YEAR_VIOLATION + LAST_THREE_YEARS_VIOLATION +
                          LAST_QUARTER_PENALTY + LAST_YEAR_PENALTY + LAST_THREE_YEARS_PENALTY
                          # + strata(MINE_ID), # use this, we get 82% (TP) + 56% (TN)
                          + strata(CURRENT_MINE_TYPE,COAL_METAL_IND), # use this, we get 58% (TP) + 77% (TN)
@@ -84,7 +77,7 @@ out_sample_model <- clogit(SEVERE ~
[…]

### 2018-07-28 Houston/src/consolidate_data.R

@@ -155,12 +155,8 @@ temp <- merge(temp, mines, by = "MINE_ID", all.x=TRUE)
 #rearranging columns
 temp <- temp[, c(1, 29:ncol(temp), 2:3, 28, 4:27)]
 
-
-q_order <- temp[order(temp[,1], temp[,6], temp[,5]),]
-
-
 # result
-complete_active_quarters <- q_order
+complete_active_quarters <- temp
 save(complete_active_quarters, file="./Houston/output/Consolidated.RData")
 
 # see simple_lm.R for simple lm

### 2018-12-06 Houston/src/manuscript_preparation.Rmd
Update manuscript_prep.Rmd

@@
 require(RcppRoll)
-require(markovchain) # FIX: markovchain seems not compatible with R3.5.0
+#require(markovchain) # FIX: markovchain seems not compatible with R3.5.0
 require(reshape)
@@
 require(SuperLearner)
+require(ranger)
 
@@
 ##Abhishek Desktop
-knitr::opts_knit$set(root.dir = "D:/Github/CRIS/violation-data-analysis/") # !! CHANGE PATH TO MATCH YOUR OWN
-source("D:/Github/CRIS/violation-data-analysis/Houston/src/prepare_violation.R")
-source("D:/Github/CRIS/violation-data-analysis/Houston/src/roll_over.R")
+# knitr::opts_knit$set(root.dir = "D:/Github/CRIS/violation-data-analysis/") # !! CHANGE PATH TO MATCH YOUR OWN
+# source("D:/Github/CRIS/violation-data-analysis/Houston/src/prepare_violation.R")
+# source("D:/Github/CRIS/violation-data-analysis/Houston/src/roll_over.R")
 
 ##Albert Macbook
-# knitr::opts_knit$set(root.dir = "~/documents/projects/violation-data-analysis") # !! CHANGE PATH TO MATCH YOUR OWN
+knitr::opts_knit$set(root.dir = "~/documents/projects/violation-data-analysis") # !! CHANGE PATH TO MATCH YOUR OWN
 
@@
                                             'last.three.years.lost','last.quarter.restrict',
-
                                             'last.year.restrict', 'last.three.years.restrict',
